### 2025-03-26 docs/R2.qmd

---
title: "Caculating Bayesian R2"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r echo=FALSE}
library(targets)
library(tidyverse)
source("R/functions.R")
```

# Read in data

# Power-law

## H
```{r}
stan_data_lr_h <- tar_read(stan_data_lr_h)
str(stan_data_lr_h)

# Without wood density
pl_nou_h <- tar_read(fit_lr2_draws_power_law1_nou_h)
pl_nou_h

gamma1_draws_pl_nou_h <- as.matrix(pl_nou_h[, "gamma[1]"])
gamma2_draws_pl_nou_h <- as.matrix(pl_nou_h[, "gamma[2]"])
sigma_draws_pl_nou_h <- as.matrix(pl_nou_h[, "sigma"])

num_sam_pl_nou_h <- nrow(gamma1_draws_pl_nou_h)
num_obs_pl_nou_h <- nrow(stan_data_lr_h$log_x)

log_y_pred_draws_pl_nou_h <- matrix(NA, nrow = num_sam_pl_nou_h, ncol = num_obs_pl_nou_h)

for (i in 1:num_sam_pl_nou_h) {
  log_y_pred_draws_pl_nou_h[i, ] <- gamma1_draws_pl_nou_h[i] + gamma2_draws_pl_nou_h[i] * stan_data_lr_h$log_x[, 2]
}

var_log_y_pred_pl_nou_h <- apply(log_y_pred_draws_pl_nou_h, 1, var)

var_res_pl_nou_h <- sigma_draws_pl_nou_h^2

bayes_R2_draws_pl_nou_h <- var_log_y_pred_pl_nou_h / (var_log_y_pred_pl_nou_h + var_res_pl_nou_h)

summary_bayes_R2_pl_nou_h <- summary(bayes_R2_draws_pl_nou_h)

summary_bayes_R2_pl_nou_h

# With wood density
pl_h <- tar_read(fit_lr2_draws_power_law1_h)
pl_h

gamma1_draws_pl_h <- as.matrix(pl_h[, "gamma[1,1]"])
gamma2_draws_pl_h <- as.matrix(pl_h[, "gamma[2,1]"])
sigma_draws_pl_h <- as.matrix(pl_h[, "sigma"])

num_sam_pl_h <- nrow(gamma1_draws_pl_h)
num_obs_pl_h <- nrow(stan_data_lr_h$log_x)

log_y_pred_draws_pl_h <- matrix(NA, nrow = num_sam_pl_h, ncol = num_obs_pl_h)

for (i in 1:num_sam_pl_h) {
  log_y_pred_draws_pl_h[i, ] <- gamma1_draws_pl_h[i] + gamma2_draws_pl_h[i] * stan_data_lr_h$log_x[, 2]
}

var_log_y_pred_pl_h <- apply(log_y_pred_draws_pl_h, 1, var)

var_res_pl_h <- sigma_draws_pl_h^2

bayes_R2_draws_pl_h <- var_log_y_pred_pl_h / (var_log_y_pred_pl_h + var_res_pl_h)

summary_bayes_R2_pl_h <- summary(bayes_R2_draws_pl_h)
summary_bayes_R2_pl_h
```

## CR

```{r}
stan_data_lr_cr <- tar_read(stan_data_lr_cr)

# Without wood density
pl_nou_cr <- tar_read(fit_lr2_draws_power_law1_nou_cr)

gamma1_draws_pl_nou_cr <- as.matrix(pl_nou_cr[, "gamma[1]"])
gamma2_draws_pl_nou_cr <- as.matrix(pl_nou_cr[, "gamma[2]"])
sigma_draws_pl_nou_cr <- as.matrix(pl_nou_cr[, "sigma"])

num_sam_pl_nou_cr <- nrow(gamma1_draws_pl_nou_cr)
num_obs_pl_nou_cr <- nrow(stan_data_lr_cr$log_x)

log_y_pred_draws_pl_nou_cr <- matrix(NA, nrow = num_sam_pl_nou_cr, ncol = num_obs_pl_nou_cr)

for (i in 1:num_sam_pl_nou_cr) {
  log_y_pred_draws_pl_nou_cr[i, ] <- gamma1_draws_pl_nou_cr[i] + 
                                     gamma2_draws_pl_nou_cr[i] * stan_data_lr_cr$log_x[, 2]
}

var_log_y_pred_pl_nou_cr <- apply(log_y_pred_draws_pl_nou_cr, 1, var)

var_res_pl_nou_cr <- sigma_draws_pl_nou_cr^2

bayes_R2_draws_pl_nou_cr <- var_log_y_pred_pl_nou_cr / (var_log_y_pred_pl_nou_cr + var_res_pl_nou_cr)

summary_bayes_R2_pl_nou_cr <- summary(bayes_R2_draws_pl_nou_cr)

summary_bayes_R2_pl_nou_cr

# With wood density
stan_data_lr_cr <- tar_read(stan_data_lr_cr)  
pl_cr <- tar_read(fit_lr2_draws_power_law1_cr)

gamma1_draws_pl_cr <- as.matrix(pl_cr[, "gamma[1,1]"])
gamma2_draws_pl_cr <- as.matrix(pl_cr[, "gamma[2,1]"])
sigma_draws_pl_cr <- as.matrix(pl_cr[, "sigma"])

num_sam_pl_cr <- nrow(gamma1_draws_cr)
num_obs_pl_cr <- nrow(stan_data_lr_cr$log_x)

log_y_pred_draws_pl_cr <- matrix(NA, nrow = num_sam_pl_cr, ncol = num_obs_pl_cr)

for (i in 1:num_sam_cr) {
  log_y_pred_draws_pl_cr[i, ] <- gamma1_draws_pl_cr[i] + gamma2_draws_pl_cr[i] * stan_data_lr_cr$log_x[, 2]
}

var_log_y_pred_pl_cr <- apply(log_y_pred_draws_pl_cr, 1, var)

var_res_pl_cr <- sigma_draws_pl_cr^2

bayes_R2_draws_pl_cr <- var_log_y_pred_pl_cr / (var_log_y_pred_pl_cr + var_res_pl_cr)

summary_bayes_R2_pl_cr <- summary(bayes_R2_draws_pl_cr)
summary_bayes_R2__pl_cr
```

# DBH

## DBH0

```{r}
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)

# Without wood density
pl_nou_dbh <- tar_read(fit_lr2_draws_power_law1_nou_dbh)
pl_nou_dbh

gamma1_draws_nou_dbh <- as.matrix(pl_nou_dbh[ , "gamma[1]"])
gamma2_draws_nou_dbh <- as.matrix(pl_nou_dbh[ , "gamma[2]"])
gamma3_draws_nou_dbh <- as.matrix(pl_nou_dbh[ , "gamma[3]"])
sigma_draws_nou_dbh <- as.matrix(pl_nou_dbh[ , "sigma"])

num_sam_nou_dbh <- nrow(gamma1_draws_nou_dbh)
num_obs_nou_dbh <- nrow(stan_data_lr_dbh$log_x)

log_y_pred_draws_nou_dbh <- matrix(NA, nrow = num_sam_nou_dbh, ncol = num_obs_nou_dbh)

for (i in 1:num_sam_nou_dbh) {
  log_y_pred_draws_nou_dbh[i, ] <- gamma1_draws_nou_dbh[i] + 
                               gamma2_draws_nou_dbh[i] * stan_data_lr_dbh$log_x[, 2] + 
                               gamma3_draws_nou_dbh[i] * stan_data_lr_dbh$log_x[, 3]
}

var_log_y_pred_nou_dbh <- apply(log_y_pred_draws_nou_dbh, 1, var)

var_res_nou_dbh <- sigma_draws_nou_dbh^2

bayes_R2_draws_nou_dbh <- var_log_y_pred_nou_dbh / (var_log_y_pred_nou_dbh + var_res_nou_dbh)

summary_bayes_R2_nou_dbh <- summary(bayes_R2_draws_nou_dbh)

summary_bayes_R2_nou_dbh

# With wood density
pl_dbh <- tar_read(fit_lr2_draws_power_law1_dbh)  
pl_dbh

gamma1_draws_dbh <- as.matrix(pl_dbh[, "gamma[1,1]"])
gamma2_draws_dbh <- as.matrix(pl_dbh[, "gamma[2,1]"])
gamma3_draws_dbh <- as.matrix(pl_dbh[, "gamma[3,1]"])
sigma_draws_dbh <- as.matrix(pl_dbh[, "sigma"])

num_sam_dbh <- nrow(gamma1_draws_dbh)
num_obs_dbh <- nrow(stan_data_lr_dbh$log_x)

log_y_pred_draws_dbh <- matrix(NA, nrow = num_sam_dbh, ncol = num_obs_dbh)

for (i in 1:num_sam_dbh) {
  log_y_pred_draws_dbh[i, ] <- gamma1_draws_dbh[i] + 
                               gamma2_draws_dbh[i] * stan_data_lr_dbh$log_x[, 2] + 
                               gamma3_draws_dbh[i] * stan_data_lr_dbh$log_x[, 3]
}

var_log_y_pred_dbh <- apply(log_y_pred_draws_dbh, 1, var)

var_res_dbh <- sigma_draws_dbh^2

bayes_R2_draws_dbh <- var_log_y_pred_dbh / (var_log_y_pred_dbh + var_res_dbh)

summary_bayes_R2_dbh <- summary(bayes_R2_draws_dbh)

summary_bayes_R2_dbh
```

## DBH1

```{r}
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_data_lr_dbh1)

# Without wood density
pl_nou_dbh1 <- tar_read(fit_lr2_draws_power_law1_nou_dbh1)
pl_nou_dbh1

gamma1_draws_nou_dbh1 <- as.matrix(pl_nou_dbh1[, "gamma[1]"])
gamma2_draws_nou_dbh1 <- as.matrix(pl_nou_dbh1[, "gamma[2]"])
sigma_draws_nou_dbh1 <- as.matrix(pl_nou_dbh1[, "sigma"])

num_sam_nou_dbh1 <- nrow(gamma1_draws_nou_dbh1)
num_obs_nou_dbh1 <- nrow(stan_data_lr_dbh1$log_x)

log_y_pred_draws_nou_dbh1 <- matrix(NA, nrow = num_sam_nou_dbh1, ncol = num_obs_nou_dbh1)

for (i in 1:num_sam_nou_dbh1) {
  log_y_pred_draws_nou_dbh1[i, ] <- gamma1_draws_nou_dbh1[i] + 
                                    gamma2_draws_nou_dbh1[i] * stan_data_lr_dbh1$log_x[, 2]
}

var_log_y_pred_nou_dbh1 <- apply(log_y_pred_draws_nou_dbh1, 1, var)

var_res_nou_dbh1 <- sigma_draws_nou_dbh1^2

bayes_R2_draws_nou_dbh1 <- var_log_y_pred_nou_dbh1 / (var_log_y_pred_nou_dbh1 + var_res_nou_dbh1)

summary_bayes_R2_nou_dbh1 <- summary(bayes_R2_draws_nou_dbh1)

summary_bayes_R2_nou_dbh1

# With wood density
pl_dbh1 <- tar_read(fit_lr2_draws_power_law1_dbh1)  
pl_dbh1

gamma1_draws_dbh1 <- as.matrix(pl_dbh1[, "gamma[1,1]"])
gamma2_draws_dbh1 <- as.matrix(pl_dbh1[, "gamma[2,1]"])
sigma_draws_dbh1 <- as.matrix(pl_dbh1[, "sigma"])

# Number of posterior samples and observations
num_sam_dbh1 <- nrow(gamma1_draws_dbh1)
num_obs_dbh1 <- nrow(stan_data_lr_dbh1$log_x)

# Calculate predicted log values across all posterior samples using two predictors
log_y_pred_draws_dbh1 <- matrix(NA, nrow = num_sam_dbh1, ncol = num_obs_dbh1)

for (i in 1:num_sam_dbh1) {
  log_y_pred_draws_dbh1[i, ] <- gamma1_draws_dbh1[i] + 
                                gamma2_draws_dbh1[i] * stan_data_lr_dbh1$log_x[, 2]
}

var_log_y_pred_dbh1 <- apply(log_y_pred_draws_dbh1, 1, var)

var_res_dbh1 <- sigma_draws_dbh1^2

bayes_R2_draws_dbh1 <- var_log_y_pred_dbh1 / (var_log_y_pred_dbh1 + var_res_dbh1)

summary_bayes_R2_dbh1 <- summary(bayes_R2_draws_dbh1)

summary_bayes_R2_dbh1
```

## DBH2

```{r}
stan_data_lr_dbh2 <- tar_read(stan_data_lr_dbh2)
str(stan_data_lr_dbh2)

# Without wood desnity
pl_nou_dbh2 <- tar_read(fit_lr2_draws_power_law1_nou_dbh2)
pl_nou_dbh2

gamma1_draws_nou_dbh2 <- as.matrix(pl_nou_dbh2[, "gamma[1]"])
gamma2_draws_nou_dbh2 <- as.matrix(pl_nou_dbh2[, "gamma[2]"])
sigma_draws_nou_dbh2 <- as.matrix(pl_nou_dbh2[, "sigma"])

num_sam_nou_dbh2 <- nrow(gamma1_draws_nou_dbh2)
num_obs_nou_dbh2 <- nrow(stan_data_lr_dbh2$log_x)

log_y_pred_draws_nou_dbh2 <- matrix(NA, nrow = num_sam_nou_dbh2, ncol = num_obs_nou_dbh2)

for (i in 1:num_sam_nou_dbh2) {
  log_y_pred_draws_nou_dbh2[i, ] <- gamma1_draws_nou_dbh2[i] + 
                                    gamma2_draws_nou_dbh2[i] * stan_data_lr_dbh2$log_x[, 2]
}

var_log_y_pred_nou_dbh2 <- apply(log_y_pred_draws_nou_dbh2, 1, var)

var_res_nou_dbh2 <- sigma_draws_nou_dbh2^2

bayes_R2_draws_nou_dbh2 <- var_log_y_pred_nou_dbh2 / (var_log_y_pred_nou_dbh2 + var_res_nou_dbh2)

summary_bayes_R2_nou_dbh2 <- summary(bayes_R2_draws_nou_dbh2)

summary_bayes_R2_nou_dbh2

# With wood density
pl_dbh2 <- tar_read(fit_lr2_draws_power_law1_dbh2)  
pl_dbh2

gamma1_draws_dbh2 <- as.matrix(pl_dbh2[, "gamma[1,1]"])
gamma2_draws_dbh2 <- as.matrix(pl_dbh2[, "gamma[2,1]"])
sigma_draws_dbh2 <- as.matrix(pl_dbh2[, "sigma"])

num_sam_dbh2 <- nrow(gamma1_draws_dbh2)
num_obs_dbh2 <- nrow(stan_data_lr_dbh2$log_x)

log_y_pred_draws_dbh2 <- matrix(NA, nrow = num_sam_dbh2, ncol = num_obs_dbh2)

for (i in 1:num_sam_dbh2) {
  log_y_pred_draws_dbh2[i, ] <- gamma1_draws_dbh2[i] + 
                                gamma2_draws_dbh2[i] * stan_data_lr_dbh2$log_x[, 2]
}

var_log_y_pred_dbh2 <- apply(log_y_pred_draws_dbh2, 1, var)

var_res_dbh2 <- sigma_draws_dbh2^2

bayes_R2_draws_dbh2 <- var_log_y_pred_dbh2 / (var_log_y_pred_dbh2 + var_res_dbh2)

summary_bayes_R2_dbh2 <- summary(bayes_R2_draws_dbh2)

summary_bayes_R2_dbh2
```

# DBH3

```{r}
stan_data_lr_dbh3 <- tar_read(stan_data_lr_dbh3)
str(stan_data_lr_dbh3)

# Without wood density
pl_nou_dbh3 <- tar_read(fit_lr2_draws_power_law1_nou_dbh3)
pl_nou_dbh3

gamma1_draws_nou_dbh3 <- as.matrix(pl_nou_dbh3[, "gamma[1]"])
gamma2_draws_nou_dbh3 <- as.matrix(pl_nou_dbh3[, "gamma[2]"])
sigma_draws_nou_dbh3 <- as.matrix(pl_nou_dbh3[, "sigma"])

num_sam_nou_dbh3 <- nrow(gamma1_draws_nou_dbh3)
num_obs_nou_dbh3 <- nrow(stan_data_lr_dbh3$log_x)

log_y_pred_draws_nou_dbh3 <- matrix(NA, nrow = num_sam_nou_dbh3, ncol = num_obs_nou_dbh3)

for (i in 1:num_sam_nou_dbh3) {
  log_y_pred_draws_nou_dbh3[i, ] <- gamma1_draws_nou_dbh3[i] + 
                                    gamma2_draws_nou_dbh3[i] * stan_data_lr_dbh3$log_x[, 2]
}

var_log_y_pred_nou_dbh3 <- apply(log_y_pred_draws_nou_dbh3, 1, var)

var_res_nou_dbh3 <- sigma_draws_nou_dbh3^2

bayes_R2_draws_nou_dbh3 <- var_log_y_pred_nou_dbh3 / (var_log_y_pred_nou_dbh3 + var_res_nou_dbh3)

summary_bayes_R2_nou_dbh3 <- summary(bayes_R2_draws_nou_dbh3)

summary_bayes_R2_nou_dbh3

# With wood density
pl_dbh3 <- tar_read(fit_lr2_draws_power_law1_dbh3)  
pl_dbh3

gamma1_draws_dbh3 <- as.matrix(pl_dbh3[, "gamma[1,1]"])
gamma2_draws_dbh3 <- as.matrix(pl_dbh3[, "gamma[2,1]"])
sigma_draws_dbh3 <- as.matrix(pl_dbh3[, "sigma"])

num_sam_dbh3 <- nrow(gamma1_draws_dbh3)
num_obs_dbh3 <- nrow(stan_data_lr_dbh3$log_x)

log_y_pred_draws_dbh3 <- matrix(NA, nrow = num_sam_dbh3, ncol = num_obs_dbh3)

for (i in 1:num_sam_dbh3) {
  log_y_pred_draws_dbh3[i, ] <- gamma1_draws_dbh3[i] + 
                                gamma2_draws_dbh3[i] * stan_data_lr_dbh3$log_x[, 2]
}

var_log_y_pred_dbh3 <- apply(log_y_pred_draws_dbh3, 1, var)

var_res_dbh3 <- sigma_draws_dbh3^2

bayes_R2_draws_dbh3 <- var_log_y_pred_dbh3 / (var_log_y_pred_dbh3 + var_res_dbh3)

summary_bayes_R2_dbh3 <- summary(bayes_R2_draws_dbh3)

summary_bayes_R2_dbh3
```

# gMM

## H
```{r}
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

# Without wood density
gmm_nou_h <- tar_read(fit_nlr_draws_gmm1_nou_h)
gmm_nou_h

gamma1_draws_gmm_nou_h <- as.matrix(gmm_nou_h[, "gamma[1]"])
gamma2_draws_gmm_nou_h <- as.matrix(gmm_nou_h[, "gamma[2]"])
gamma3_draws_gmm_nou_h <- as.matrix(gmm_nou_h[, "gamma[3]"])

sigma_draws_gmm_nou_h <- as.matrix(gmm_nou_h[, "sigma"])

num_sam_gmm_nou_h <- nrow(gamma1_draws_gmm_nou_h)
num_obs_gmm_nou_h <- length(stan_data_nlr_h$x)

log_y_pred_draws_gmm_nou_h <- matrix(NA, nrow = num_sam_gmm_nou_h, ncol = num_obs_gmm_nou_h)

for (i in 1:num_sam_gmm_nou_h) {
  log_y_pred_draws_gmm_nou_h[i, ] <- exp(gamma1_draws_gmm_nou_h[i]) + 
                                     exp(gamma2_draws_gmm_nou_h[i]) * log(stan_data_nlr_h$x) - 
                                     log(exp(gamma3_draws_gmm_nou_h[i]) + (stan_data_nlr_h$x)^(exp(gamma2_draws_gmm_nou_h[i])))
}

var_log_y_pred_gmm_nou_h <- apply(log_y_pred_draws_gmm_nou_h, 1, var)

var_res_gmm_nou_h <- sigma_draws_gmm_nou_h^2

bayes_R2_draws_gmm_nou_h <- var_log_y_pred_gmm_nou_h / (var_log_y_pred_gmm_nou_h + var_res_gmm_nou_h)

summary_bayes_R2_gmm_nou_h <- summary(bayes_R2_draws_gmm_nou_h)
summary_bayes_R2_gmm_nou_h

# With wood density
gmm_h <- tar_read(fit_nlr_draws_gmm1_h)
gmm_h

gamma1_draws_gmm_h <- as.matrix(gmm_h[, "gamma[1,1]"])
gamma2_draws_gmm_h <- as.matrix(gmm_h[, "gamma[1,2]"])
gamma3_draws_gmm_h <- as.matrix(gmm_h[, "gamma[1,3]"])
sigma_draws_gmm_h <- as.matrix(gmm_h[, "sigma"])

num_sam_gmm_h <- nrow(gamma1_draws_gmm_h)
num_obs_gmm_h <- length(stan_data_nlr_h$x)

log_y_pred_draws_gmm_h <- matrix(NA, nrow = num_sam_gmm_h, ncol = num_obs_gmm_h)

for (i in 1:num_sam_gmm_h) {
  log_y_pred_draws_gmm_h[i, ] <- exp(gamma1_draws_gmm_h[i]) + 
                                 exp(gamma2_draws_gmm_h[i]) * log(stan_data_nlr_h$x) - 
                                 log(exp(gamma3_draws_gmm_h[i]) + (stan_data_nlr_h$x)^(exp(gamma2_draws_gmm_h[i])))
}

var_log_y_pred_gmm_h <- apply(log_y_pred_draws_gmm_h, 1, var)

var_res_gmm_h <- sigma_draws_gmm_h^2

bayes_R2_draws_gmm_h <- var_log_y_pred_gmm_h / (var_log_y_pred_gmm_h + var_res_gmm_h)

summary_bayes_R2_gmm_h <- summary(bayes_R2_draws_gmm_h)
summary_bayes_R2_gmm_h
```

## CR

```{r}
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

# Without wood density
gmm_nou_cr <- tar_read(fit_nlr_draws_gmm1_nou_cr)
gmm_nou_cr

gamma1_draws_gmm_nou_cr <- as.matrix(gmm_nou_cr[, "gamma[1]"])
gamma2_draws_gmm_nou_cr <- as.matrix(gmm_nou_cr[, "gamma[2]"])
gamma3_draws_gmm_nou_cr <- as.matrix(gmm_nou_cr[, "gamma[3]"])

sigma_draws_gmm_nou_cr <- as.matrix(gmm_nou_cr[, "sigma"])

num_sam_gmm_nou_cr <- nrow(gamma1_draws_gmm_nou_cr)
num_obs_gmm_nou_cr <- length(stan_data_nlr_cr$x)

log_y_pred_draws_gmm_nou_cr <- matrix(NA, nrow = num_sam_gmm_nou_cr, ncol = num_obs_gmm_nou_cr)

for (i in 1:num_sam_gmm_nou_cr) {
  log_y_pred_draws_gmm_nou_cr[i, ] <- exp(gamma1_draws_gmm_nou_cr[i]) + 
                                      exp(gamma2_draws_gmm_nou_cr[i]) * log(stan_data_nlr_cr$x) - 
                                      log(exp(gamma3_draws_gmm_nou_cr[i]) + (stan_data_nlr_cr$x)^(exp(gamma2_draws_gmm_nou_cr[i])))
}

var_log_y_pred_gmm_nou_cr <- apply(log_y_pred_draws_gmm_nou_cr, 1, var)

var_res_gmm_nou_cr <- sigma_draws_gmm_nou_cr^2

bayes_R2_draws_gmm_nou_cr <- var_log_y_pred_gmm_nou_cr / (var_log_y_pred_gmm_nou_cr + var_res_gmm_nou_cr)

summary_bayes_R2_gmm_nou_cr <- summary(bayes_R2_draws_gmm_nou_cr)
summary_bayes_R2_gmm_nou_cr

# With wood density
gmm_cr <- tar_read(fit_nlr_draws_gmm1_cr)
gmm_cr

gamma1_draws_gmm_cr <- as.matrix(gmm_cr[, "gamma[1,1]"])
gamma2_draws_gmm_cr <- as.matrix(gmm_cr[, "gamma[1,2]"])
gamma3_draws_gmm_cr <- as.matrix(gmm_cr[, "gamma[1,3]"])
sigma_draws_gmm_cr <- as.matrix(gmm_cr[, "sigma"])

num_sam_gmm_cr <- nrow(gamma1_draws_gmm_cr)
num_obs_gmm_cr <- length(stan_data_nlr_cr$x)

log_y_pred_draws_gmm_cr <- matrix(NA, nrow = num_sam_gmm_cr, ncol = num_obs_gmm_cr)

for (i in 1:num_sam_gmm_cr) {
  log_y_pred_draws_gmm_cr[i, ] <- exp(gamma1_draws_gmm_cr[i]) + 
                                  exp(gamma2_draws_gmm_cr[i]) * log(stan_data_nlr_cr$x) - 
                                  log(exp(gamma3_draws_gmm_cr[i]) + (stan_data_nlr_cr$x)^(exp(gamma2_draws_gmm_cr[i])))
}

var_log_y_pred_gmm_cr <- apply(log_y_pred_draws_gmm_cr, 1, var)

var_res_gmm_cr <- sigma_draws_gmm_cr^2

bayes_R2_draws_gmm_cr <- var_log_y_pred_gmm_cr / (var_log_y_pred_gmm_cr + var_res_gmm_cr)

summary_bayes_R2_gmm_cr <- summary(bayes_R2_draws_gmm_cr)
summary_bayes_R2_gmm_cr
```

# Weibull

## H

```{r}
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
str(stan_data_nlr_h)

# Without wood density
wb_nou_h <- tar_read(fit_nlr_draws_weibull1_nou_h)
wb_nou_h

gamma1_draws_wb_nou_h <- as.matrix(wb_nou_h[, "gamma[1]"])
gamma2_draws_wb_nou_h <- as.matrix(wb_nou_h[, "gamma[2]"])
gamma3_draws_wb_nou_h <- as.matrix(wb_nou_h[, "gamma[3]"])

sigma_draws_wb_nou_h <- as.matrix(wb_nou_h[, "sigma"])

num_sam_wb_nou_h <- nrow(gamma1_draws_wb_nou_h)
num_obs_wb_nou_h <- length(stan_data_nlr_h$x)

log_y_pred_draws_wb_nou_h <- matrix(NA, nrow = num_sam_wb_nou_h, ncol = num_obs_wb_nou_h)

for (i in 1:num_sam_wb_nou_h) {
  log_y_pred_draws_wb_nou_h[i, ] <- exp(gamma1_draws_wb_nou_h[i]) + 
                                    log(1- exp(-(exp(gamma2_draws_wb_nou_h[i])) * (stan_data_nlr_h$x)^(exp(gamma3_draws_wb_nou_h[i]))))
}

var_log_y_pred_wb_nou_h <- apply(log_y_pred_draws_wb_nou_h, 1, var)

var_res_wb_nou_h <- sigma_draws_wb_nou_h^2

bayes_R2_draws_wb_nou_h <- var_log_y_pred_wb_nou_h / (var_log_y_pred_wb_nou_h + var_res_wb_nou_h)

summary_bayes_R2_wb_nou_h <- summary(bayes_R2_draws_wb_nou_h)
summary_bayes_R2_wb_nou_h

# With wood density
wb_h <- tar_read(fit_nlr_draws_weibull1_h)
wb_h

# Extract the gamma and sigma draws
gamma1_draws_wb_h <- as.matrix(wb_h[, "gamma[1,1]"])
gamma2_draws_wb_h <- as.matrix(wb_h[, "gamma[1,2]"])
gamma3_draws_wb_h <- as.matrix(wb_h[, "gamma[1,3]"])
sigma_draws_wb_h <- as.matrix(wb_h[, "sigma"])

num_sam_wb_h <- nrow(gamma1_draws_wb_h)
num_obs_wb_h <- length(stan_data_nlr_h$x)

log_y_pred_draws_wb_h <- matrix(NA, nrow = num_sam_wb_h, ncol = num_obs_wb_h)

for (i in 1:num_sam_wb_h) {
  log_y_pred_draws_wb_h[i, ] <- exp(gamma1_draws_wb_h[i]) + 
                                log(1 - exp(-(exp(gamma2_draws_wb_h[i])) * 
                                (stan_data_nlr_h$x)^(exp(gamma3_draws_wb_h[i]))))
}

var_log_y_pred_wb_h <- apply(log_y_pred_draws_wb_h, 1, var)

var_res_wb_h <- sigma_draws_wb_h^2

bayes_R2_draws_wb_h <- var_log_y_pred_wb_h / (var_log_y_pred_wb_h + var_res_wb_h)

summary_bayes_R2_wb_h <- summary(bayes_R2_draws_wb_h)

summary_bayes_R2_wb_h
```

## CR

```{r}
wb_nou_cr <- tar_read(fit_nlr_draws_weibull1_nou_cr)
wb_nou_cr

gamma1_draws_wb_nou_cr <- as.matrix(wb_nou_cr[, "gamma[1]"])
gamma2_draws_wb_nou_cr <- as.matrix(wb_nou_cr[, "gamma[2]"])
gamma3_draws_wb_nou_cr <- as.matrix(wb_nou_cr[, "gamma[3]"])
sigma_draws_wb_nou_cr <- as.matrix(wb_nou_cr[, "sigma"])

num_sam_wb_nou_cr <- nrow(gamma1_draws_wb_nou_cr)
num_obs_wb_nou_cr <- length(stan_data_nlr_cr$x)

log_y_pred_draws_wb_nou_cr <- matrix(NA, nrow = num_sam_wb_nou_cr, ncol = num_obs_wb_nou_cr)

for (i in 1:num_sam_wb_nou_cr) {
  log_y_pred_draws_wb_nou_cr[i, ] <- exp(gamma1_draws_wb_nou_cr[i]) + 
                                     log(1 - exp(-(exp(gamma2_draws_wb_nou_cr[i])) * 
                                     (stan_data_nlr_cr$x)^(exp(gamma3_draws_wb_nou_cr[i]))))
}

var_log_y_pred_wb_nou_cr <- apply(log_y_pred_draws_wb_nou_cr, 1, var)

var_res_wb_nou_cr <- sigma_draws_wb_nou_cr^2

bayes_R2_draws_wb_nou_cr <- var_log_y_pred_wb_nou_cr / (var_log_y_pred_wb_nou_cr + var_res_wb_nou_cr)

summary_bayes_R2_wb_nou_cr <- summary(bayes_R2_draws_wb_nou_cr)

summary_bayes_R2_wb_nou_cr

# With wood density 
wb_cr <- tar_read(fit_nlr_draws_weibull1_cr)
wb_cr

gamma1_draws_wb_cr <- as.matrix(wb_cr[, "gamma[1,1]"])
gamma2_draws_wb_cr <- as.matrix(wb_cr[, "gamma[1,2]"])
gamma3_draws_wb_cr <- as.matrix(wb_cr[, "gamma[1,3]"])
sigma_draws_wb_cr <- as.matrix(wb_cr[, "sigma"])

num_sam_wb_cr <- nrow(gamma1_draws_wb_cr)
num_obs_wb_cr <- length(stan_data_nlr_cr$x)

log_y_pred_draws_wb_cr <- matrix(NA, nrow = num_sam_wb_cr, ncol = num_obs_wb_cr)

for (i in 1:num_sam_wb_cr) {
  log_y_pred_draws_wb_cr[i, ] <- exp(gamma1_draws_wb_cr[i]) + 
                                 log(1 - exp(-(exp(gamma2_draws_wb_cr[i])) * 
                                 (stan_data_nlr_cr$x)^(exp(gamma3_draws_wb_cr[i]))))
}

var_log_y_pred_wb_cr <- apply(log_y_pred_draws_wb_cr, 1, var)

var_res_wb_cr <- sigma_draws_wb_cr^2

bayes_R2_draws_wb_cr <- var_log_y_pred_wb_cr / (var_log_y_pred_wb_cr + var_res_wb_cr)

summary_bayes_R2_wb_cr <- summary(bayes_R2_draws_wb_cr)

summary_bayes_R2_wb_cr
```

```{r}
# H (Height) data
stan_data_lr_h <- tar_read(stan_data_lr_h)
pl_nou_h <- tar_read(fit_lr2_draws_power_law1_nou_h)
pl_h <- tar_read(fit_lr2_draws_power_law1_h)

# CR (Crown Radius) data
stan_data_lr_cr <- tar_read(stan_data_lr_cr)
pl_nou_cr <- tar_read(fit_lr2_draws_power_law1_nou_cr)
pl_cr <- tar_read(fit_lr2_draws_power_law1_cr)

# DBH0 (Diameter at Breast Height) data
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
pl_nou_dbh <- tar_read(fit_lr2_draws_power_law1_nou_dbh)
pl_dbh <- tar_read(fit_lr2_draws_power_law1_dbh)

# DBH1 data
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
pl_nou_dbh1 <- tar_read(fit_lr2_draws_power_law1_nou_dbh1)
pl_dbh1 <- tar_read(fit_lr2_draws_power_law1_dbh1)

# DBH2 data
stan_data_lr_dbh2 <- tar_read(stan_data_lr_dbh2)
pl_nou_dbh2 <- tar_read(fit_lr2_draws_power_law1_nou_dbh2)
pl_dbh2 <- tar_read(fit_lr2_draws_power_law1_dbh2)

# DBH3 data
stan_data_lr_dbh3 <- tar_read(stan_data_lr_dbh3)
pl_nou_dbh3 <- tar_read(fit_lr2_draws_power_law1_nou_dbh3)
pl_dbh3 <- tar_read(fit_lr2_draws_power_law1_dbh3)

# gMM models for H
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
gmm_nou_h <- tar_read(fit_nlr_draws_gmm1_nou_h)
gmm_h <- tar_read(fit_nlr_draws_gmm1_h)

# gMM models for CR
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
gmm_nou_cr <- tar_read(fit_nlr_draws_gmm1_nou_cr)
gmm_cr <- tar_read(fit_nlr_draws_gmm1_cr)

# Weibull models for H
wb_nou_h <- tar_read(fit_nlr_draws_weibull1_nou_h)
wb_h <- tar_read(fit_nlr_draws_weibull1_h)

# Weibull models for CR
wb_nou_cr <- tar_read(fit_nlr_draws_weibull1_nou_cr)
wb_cr <- tar_read(fit_nlr_draws_weibull1_cr)
```

# Usage

```{r}
source("R/functions.R")

# lr
bayes_R2_pl_nou_h <- calculate_bayes_R2_lr(stan_data_lr_h, pl_nou_h, c("gamma[1]", "gamma[2]"))
bayes_R2_pl_h <- calculate_bayes_R2_lr(stan_data_lr_h, pl_h, c("gamma[1,1]", "gamma[2,1]"))

bayes_R2_pl_nou_cr <- calculate_bayes_R2_lr(stan_data_lr_cr, pl_nou_cr, c("gamma[1]", "gamma[2]"))
bayes_R2_pl_cr <- calculate_bayes_R2_lr(stan_data_lr_cr, pl_cr, c("gamma[1,1]", "gamma[2,1]"))

bayes_R2_pl_nou_dbh <- calculate_bayes_R2_lr(stan_data_lr_dbh, pl_nou_dbh, c("gamma[1]", "gamma[2]", "gamma[3]"))
bayes_R2_pl_dbh <- calculate_bayes_R2_lr(stan_data_lr_dbh, pl_dbh, c("gamma[1,1]", "gamma[2,1]", "gamma[3,1]"))

bayes_R2_pl_nou_dbh1 <- calculate_bayes_R2_lr(stan_data_lr_dbh1, pl_nou_dbh1, c("gamma[1]", "gamma[2]"))
bayes_R2_pl_dbh1 <- calculate_bayes_R2_lr(stan_data_lr_dbh1, pl_dbh1, c("gamma[1,1]", "gamma[2,1]"))

bayes_R2_pl_nou_dbh2 <- calculate_bayes_R2_lr(stan_data_lr_dbh2, pl_nou_dbh2, c("gamma[1]", "gamma[2]"))
bayes_R2_pl_dbh2 <- calculate_bayes_R2_lr(stan_data_lr_dbh2, pl_dbh2, c("gamma[1,1]", "gamma[2,1]"))

bayes_R2_pl_nou_dbh3 <- calculate_bayes_R2_lr(stan_data_lr_dbh3, pl_nou_dbh3, c("gamma[1]", "gamma[2]"))
bayes_R2_pl_dbh3 <- calculate_bayes_R2_lr(stan_data_lr_dbh3, pl_dbh3, c("gamma[1,1]", "gamma[2,1]"))

bayes_R2_gmm_nou_h <- calculate_bayes_R2_nlr(stan_data_nlr_h, gmm_nou_h, c("gamma[1]", "gamma[2]", "gamma[3]"), "gmm")
bayes_R2_gmm_h <- calculate_bayes_R2_nlr(stan_data_nlr_h, gmm_h, c("gamma[11,]", "gamma[1,2]", "gamma[1,3]"), "gmm")

bayes_R2_gmm_nou_cr <- calculate_bayes_R2_nlr(stan_data_nlr_cr, gmm_nou_cr, c("gamma[1]", "gamma[2]", "gamma[3]"), "gmm")
bayes_R2_gmm_cr <- calculate_bayes_R2_nlr(stan_data_nlr_cr, gmm_cr, c("gamma[1,1]", "gamma[1,2]", "gamma[1,3]"), "gmm")

bayes_R2_wb_nou_h <- calculate_bayes_R2_nlr(stan_data_nlr_h, wb_nou_h, c("gamma[1]", "gamma[2]", "gamma[3]"), "gmm")
bayes_R2_wb_h <- calculate_bayes_R2_nlr(stan_data_nlr_h, wb_h, c("gamma[1,1]", "gamma[1,2]", "gamma[1,3]"), "gmm")

bayes_R2_wb_nou_cr <- calculate_bayes_R2_nlr(stan_data_nlr_cr, wb_nou_cr, c("gamma[1]", "gamma[2]", "gamma[3]"), "weibull")
bayes_R2_wb_cr <- calculate_bayes_R2_nlr(stan_data_nlr_cr, wb_cr, c("gamma[1,1]", "gamma[1,2]", "gamma[1,3]"), "weibull")

```

```{r}
source("R/functions.R")
r2_df <- tar_read(R2_df)
r2_df
r2_tbl <- generate_r2_table(r2_df)
r2_tbl

loo_tbl <- tar_read(loo_tbl)
loo_tbl
com_tbl <- generate_comparison_table(loo_tbl)
com_tbl

loo_r2_tbl <- loo_r2_table(r2_tbl, com_tbl)
loo_r2_tbl
```