### 2025-03-26 docs/allo_relationships.qmd

```{r}
library(readr)
library(tidyverse)
library(here)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(patchwork)
```

# data preparation
 ```{r}
allo <- read_csv("data/overlap.csv")
head(allo, 6)
allo2 <- head(allo, 300000)
```

- checking data
```{r}
hist_h <- ggplot(allo2, aes(x = h)) +
  geom_histogram(fill = "lightblue", color = "black") +
  labs(title = "Histogram of Tree Height", x = "Height (m)", y = "Frequency")

hist_dbh <- ggplot(allo2, aes(x = dbh)) +
  geom_histogram(fill = "lightgreen", color = "black") +
  labs(title = "Histogram of DBH", x = "DBH (cm)", y = "Frequency")

hist_cr <- ggplot(allo2, aes(x = cr)) +
  geom_histogram(fill = "lightpink", color = "black") +
  labs(title = "Histogram of CR", x = "CR", y = "Frequency")

combined_plot <- grid.arrange(hist_h, hist_dbh, hist_cr, ncol = 3)
ggsave("figs/combined_histograms.png", plot = combined_plot, width = 12, height = 6)
```

- tranformation of data
```{r}
allo2$dbh.ln <- log(allo2$dbh)
allo2$h.ln <- log(allo2$h)
allo2$cr.ln <- log(allo2$cr)
```

- checking again
```{r}
hist_h_ln <- ggplot(allo2, aes(x = h.ln)) +
  geom_histogram(fill = "lightblue", color = "black") +
  labs(title = "Histogram of Log(Height)", x = "Log(Height)", y = "Frequency")

hist_dbh_ln <- ggplot(allo2, aes(x = dbh.ln)) +
  geom_histogram(fill = "lightgreen", color = "black") +
  labs(title = "Histogram of Log(DBH)", x = "Log(DBH)", y = "Frequency")

hist_cr_ln <- ggplot(allo2, aes(x = cr.ln)) +
  geom_histogram(fill = "lightpink", color = "black") +
  labs(title = "Histogram of Log(CR)", x = "Log(CR)", y = "Frequency")

combined_plot_ln <- grid.arrange(hist_h_ln, hist_dbh_ln, hist_cr_ln, ncol = 3)
ggsave("figs/combined_histograms_ln.png", plot = combined_plot_ln, width = 12, height = 6)
```

- negative values in cr => transfer
```{r}
allo2$cr.ln <- log(allo2$cr + 1)
allo3 <- allo2
summary(allo3)
```

```{r}
hist_h_ln <- ggplot(allo2, aes(x = h.ln)) +
  geom_histogram(fill = "lightblue", color = "black") +
  labs(title = "Histogram of Log(Height)", x = "Log(Height)", y = "Frequency")

hist_dbh_ln <- ggplot(allo2, aes(x = dbh.ln)) +
  geom_histogram(fill = "lightgreen", color = "black") +
  labs(title = "Histogram of Log(DBH)", x = "Log(DBH)", y = "Frequency")

hist_cr_ln <- ggplot(allo2, aes(x = cr.ln)) +
  geom_histogram(fill = "lightpink", color = "black") +
  labs(title = "Histogram of Log(CR)", x = "Log(CR)", y = "Frequency")

combined_plot_ln <- grid.arrange(hist_h_ln, hist_dbh_ln, hist_cr_ln, ncol = 3)
ggsave("figs/combined_histograms_ln_plus_1.png", plot = combined_plot_ln, width = 12, height = 6)
```

# fitting power law

```{r}
dhmod1 <- lm((dbh.ln ~ h.ln), data = allo3) 
dcmod1 <- lm((dbh.ln ~ cr.ln), data = allo3)
hdmod1 <- lm((h.ln ~ dbh.ln), data = allo3)
cdmod1 <- lm((cr.ln ~ dbh.ln), data = allo3)
dhcmod1 <- lm((dbh.ln ~ h.ln + cr.ln), data = allo3)
```

 - plotting models

```{r}
plot_dhmod1 <- ggplot(allo2, aes(x = h.ln, y = dbh.ln)) +
  geom_point() + # Add data points
  labs(title = "Scatter Plot: dbh.ln vs. h.ln", x = "h.ln", y = "dbh.ln")

plot_dcmod1 <- ggplot(allo2, aes(x = cr.ln, y = dbh.ln)) +
  geom_point() +
  labs(title = "Scatter Plot: dbh.ln vs. cr.ln", x = "cr.ln", y = "dbh.ln")

plot_hdmod1 <- ggplot(allo2, aes(x = dbh.ln, y = h.ln)) +
  geom_point() +
  labs(title = "Scatter Plot: h.ln vs. dbh.ln", x = "dbh.ln", y = "h.ln")

plot_cdmod1 <- ggplot(allo2, aes(x = dbh.ln, y = cr.ln)) +
  geom_point() +
  labs(title = "Scatter Plot: cr.ln vs. dbh.ln", x = "dbh.ln", y = "cr.ln")

plot_dhcmod1 <- ggplot(allo2, aes(x = h.ln + cr.ln, y = dbh.ln)) +
  geom_point() +
  labs(title = "Scatter Plot: dbh.ln vs. h.ln + cr.ln", x = "h.ln + cr.ln", y = "dbh.ln")

# Arrange plots in one figure
combined_plots <- grid.arrange(plot_dhmod1, plot_dcmod1, plot_hdmod1, plot_cdmod1, plot_dhcmod1,
                               nrow = 2, ncol = 3)

# Save the combined plot
ggsave("figs/combined_plots.png", plot = combined_plots, width = 12, height = 8)             
```

```{r}
summary(dhmod1)
summary(dcmod1)
summary(hdmod1)
summary(cdmod1)
summary(dhcmod1)
```

```{r}
line_plot_allo2 <- ggplot(allo2) +
  geom_line(aes(x = h.ln, y = dbh.ln, group = sp)) +
  labs(title = "Log(DBH) vs. Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")
# Save the plot
ggsave("figs/line_plot_allo2.png", plot = line_plot_allo2, width = 12, height = 8)
```

```{r}
# Define the plot with points and linear regression lines
line_plot_allo2 <- ggplot(allo2, aes(x = h.ln, y = dbh.ln, group = sp)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", se = FALSE, color = "black") +  # Add linear regression line
  labs(title = "Log(DBH) vs. Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")

# Save the plot
ggsave("figs/_scatter_plot_allo2_with_lm.png", plot = line_plot_allo2, width = 12, height = 8)
```

```{r}
plot <- ggplot(data = allo2, aes(y = h.ln, x = dbh.ln, group = sp, group = sp, shape = sp)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1) +
  theme_minimal() +
  labs(title = "Relationship between DBH and Height by Species",
       subtitle = "Regression lines for each species, colored and shaped by species",
       group = "sp", shape = "sp")
ggsave("figs/_scatter_plot_allo2_with_lm.png", plot = plot, width = 12, height = 8)
```

work
```{r}
scatter_plot <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", se = FALSE, color = "green") +  # Add linear regression line
  labs(title = "Scatter Plot of Log(DBH) vs. Log(Height)",
       x = "Log(Height)", y = "Log(DBH)")

# Save the plot
ggsave("figs/scatter_plot_dbh_vs_height.png", plot = scatter_plot, width = 12, height = 8)
```

- does not work 
#```{r}
scatter_plot <- ggplot(allo3, aes(x = h.ln, y = dbh.ln, color = sp)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", se = FALSE) +  # Add separate linear regression lines for each species
  labs(title = "Scatter Plot of Log(DBH) vs. Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")

ggsave("figs/scatter_plot_dbh_vs_height_by_species.png", plot = scatter_plot, width = 12, height = 8)

```

- does not work
```{r}
scatter_plot <- ggplot(allo3, aes(x = h.ln, y = dbh.ln, color = sp)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", se = FALSE, aes(group = sp)) +  # Add separate linear regression lines for each species
  labs(title = "Scatter Plot of Log(DBH) vs. Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")

# Save the plot
ggsave("figs/scatter_plot_dbh_vs_height_by_species.png", plot = scatter_plot, width = 12, height = 8)
```

```{r}
# Scatter plot with separate linear regression lines for each species
scatter_plot <- ggplot(allo3, aes(x = h.ln, y = dbh.ln, color = sp)) +
  geom_point() +  # Add points
  geom_smooth(method = "lm", se = FALSE, aes(group = sp), color = "black") +  # Add separate linear regression lines for each species
  labs(title = "Scatter Plot of Log(DBH) vs. Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")

# Save the plot
ggsave("figs/scatter_plot_dbh_vs_height_by_species.png", plot = scatter_plot, width = 12, height = 8)
```

#```{r}
plot_allo3 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln, group = sp, color = sp)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression lines
  labs(title = "Relationship between Log(DBH) and Log(Height) by Species",
       x = "Log(Height)", y = "Log(DBH)")

# Save the plot
ggsave("figs/relationship_dbh_height_by_species.png", plot = plot_allo3, width = 12, height = 8)
#```

```{r}
plot_dhmod1 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() + # Add data points
  geom_smooth(method = "lm", se = FALSE, aes(group = sp)) + # Add linear regression lines for each species
  labs(title = "Scatter Plot: dbh.ln vs. h.ln by Species", x = "h.ln", y = "dbh.ln")

# Save the plot
ggsave("figs/scatter_plot_species_lm_lines.png", plot = plot_dhmod1, width = 12, height = 8)
```

#```{r}
plot_dhmod1 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() + # Add data points
  geom_smooth(method = "lm", se = FALSE, aes(group = sp, colour = sp)) + # Add linear regression lines for each species
  labs(title = "Scatter Plot: dbh.ln vs. h.ln by Species", x = "h.ln", y = "dbh.ln")

# Save the plot
ggsave("figs/scatter_plot_species_lm_color.png", plot = plot_dhmod1, width = 12, height = 8)
```

```{r}
plot_dhmod1 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() + # Add data points
  geom_smooth(method = "lm", se = FALSE, aes(colour = sp)) + # Add linear regression lines for each species
  labs(title = "Scatter Plot: dbh.ln vs. h.ln by Species", x = "h.ln", y = "dbh.ln")

# Save the plot
ggsave("figs/scatter_plot_species_lm_colour.png", plot = plot_dhmod1, width = 12, height = 8)
```

```{r}
scatter_plot <- ggplot( allo3, aes(x = h.ln, y = dbh.ln))+
 geom_point()+
 geom_smooth(method = "lm", se = FALSE, color ="blue") +
 labs( title = "Plot of log(dbh) vs. log(h)"; x = "log(h)", y = "log(dbh)")
ggsave("figs/scatter_plot_dbh_vs_h.png", plot = scatter_plot, width = 12, height = 8)
```

```{r}
plot_dcmod1 <- ggplot(allo3, aes(x = cr.ln, y = dbh.ln)) +
  geom_point() + # Add data points
  geom_smooth(method = "lm", se = FALSE, aes(group = sp)) + # Add linear regression lines for each species
  labs(title = "Scatter Plot: dbh.ln vs. cr.ln by Species", x = "cr.ln", y = "dbh.ln")

# Save the plot
ggsave("figs/dcmod1_scatter_plot_species_lm_lines.png", plot = plot_dcmod1, width = 12, height = 8)
```

```{r}
# Scatter plot for dhmod1
plot_dhmod1 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot: dbh.ln vs. h.ln", x = "h.ln", y = "dbh.ln")

# Scatter plot for dcmod1
plot_dcmod1 <- ggplot(allo3, aes(x = cr.ln, y = dbh.ln)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot: dbh.ln vs. cr.ln", x = "cr.ln", y = "dbh.ln")

# Scatter plot for hdmod1
plot_hdmod1 <- ggplot(allo3, aes(x = dbh.ln, y = h.ln)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot: h.ln vs. dbh.ln", x = "dbh.ln", y = "h.ln")

# Scatter plot for cdmod1
plot_cdmod1 <- ggplot(allo3, aes(x = dbh.ln, y = cr.ln)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter Plot: cr.ln vs. dbh.ln", x = "dbh.ln", y = "cr.ln")

# Scatter plot for dhcmod1
plot_dhcmod1 <- ggplot(allo3, aes(x = h.ln, y = dbh.ln)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, aes(color = cr.ln)) +
  labs(title = "Scatter Plot: dbh.ln vs. h.ln", x = "h.ln", y = "dbh.ln")

# Combine all plots into one figure
combined_plots <- grid.arrange(plot_dhmod1, plot_dcmod1, plot_hdmod1, plot_cdmod1, plot_dhcmod1,
                               nrow = 2, ncol = 3, 
                               top = textGrob("Scatter Plots with Fitted Regression Lines", 
                                              gp = gpar(fontsize = 16, fontface = "bold")))

# Save figure 
ggsave("figs/power_combined_plots.png", combined_plots, width = 15, height = 10)
```

# fitting gMM (drm())
```{r}
library(drc)
library(minpack.lm)

```

- Remove NaNs

```{r}
allo4 <- allo3 %>% 
  mutate(hcr = h.ln + cr.ln)
head(allo4)
```

```{r}
sum(is.na(allo4))
allo4 <- na.omit(allo4)
summary(allo4)
```

```{r}
michaelis_menten <- function(x, a, b, k) {
  a * x^b / (k + x^b)
}
dhmod2 <- nlsLM(dbh.ln ~ michaelis_menten(h.ln, a, b, k), data = allo4,
                 start = list(a = 1, b = 5, k = 0.1), 
                 control = nls.lm.control(maxiter = 100, maxfev = 500))

dcmod2 <- nlsLM(dbh.ln ~ michaelis_menten(cr.ln, a, b, k), data = allo4,
                start = list(a = 1, b = 1, k = 1))

hdmod2 <- nlsLM(h.ln ~ michaelis_menten(dbh.ln, a, b, k), data = allo4,
                start = list(a = 1, b = 1, k = 1))

cdmod2 <- nlsLM(cr.ln ~ michaelis_menten(dbh.ln, a, b, k), 
                data = allo4,
                start = list(a = 1, b = 1, k = 1),
                control = nls.lm.control(maxiter = 500, maxfev = 1000))

dhcmod2 <- nlsLM(dbh.ln ~ michaelis_menten(hcr, a, b, k), data = allo4,
                 start = list(a = 1, b = 1, k = 1),
                 control = nls.lm.control(maxiter = 500, maxfev = 500))

summary(dhmod2)
summary(dcmod2)
summary(hdmod2)
summary(cdmod2)
summary(dhcmod2)
```