### 2025-03-26 docs/bays_check1.qmd

---
title: "Diagnostics for MCMC"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab #readable #sandstone #spacelab #flatly
    highlight: pygments #tango #kate
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    toc_float:
     collapsed: TRUE
     smooth_scroll: TRUE
     toc: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(dplyr)
```

# Models

```{r}
loo_tbl <- tar_read(loo_tbl)
loo_tbl
print(loo_tbl , n = Inf)

#gMM1, PL1, and Weibull1 models
loo_tbl <- loo_tbl |>
  filter(grepl("gmm1|power_law1|weibull1", model))
print(loo_tbl)
```

8 models to predict DBH:

- DBH: DBH = a.(CR^b).(H^c)

- DBH1: DBH = a.(CR.H)^b

- DBH2: DBH = a.CR^b

- DBH3: DBH = a.H^b

6 models to predic CR:

- CR ~ DBH (pl1, gMM1, Weibull1)

6 models to predic H:

- H ~ DBH (pl1, gMM1, Weibull1)

Species-level Predictor: With and Without Wood density

# Convergence

I consider the ESS < 300 | Rhat >1.1 to indicate the convergence issue.

```{r}
options(width = 150)
pl1_cr <- tar_read(fit_lr2_summary_power_law1_cr)
pl1_h <- tar_read(fit_lr2_summary_power_law1_h)
pl1_dbh <- tar_read(fit_lr2_summary_power_law1_dbh)
pl1_dbh1 <- tar_read(fit_lr2_summary_power_law1_dbh1)
pl1_dbh2 <- tar_read(fit_lr2_summary_power_law1_dbh2)
pl1_dbh3 <- tar_read(fit_lr2_summary_power_law1_dbh3)
pl1_nou_cr <- tar_read(fit_lr2_summary_power_law1_nou_cr)
pl1_nou_h <- tar_read(fit_lr2_summary_power_law1_nou_h)
pl1_nou_dbh <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
pl1_nou_dbh1 <- tar_read(fit_lr2_summary_power_law1_nou_dbh1)
pl1_nou_dbh2 <- tar_read(fit_lr2_summary_power_law1_nou_dbh2)
pl1_nou_dbh3 <- tar_read(fit_lr2_summary_power_law1_nou_dbh3)

gmm1_cr <- tar_read(fit_nlr_summary_gmm1_cr)
gmm1_h <- tar_read(fit_nlr_summary_gmm1_h)
gmm1_nou_cr <- tar_read(fit_nlr_summary_gmm1_nou_cr)
gmm1_nou_h <- tar_read(fit_nlr_summary_gmm1_nou_h)
weibull1_cr <- tar_read(fit_nlr_summary_weibull1_cr)
weibull1_h <- tar_read(fit_nlr_summary_weibull1_h)
weibull1_nou_cr <- tar_read(fit_nlr_summary_weibull1_nou_cr)
weibull1_nou_h <- tar_read(fit_nlr_summary_weibull1_nou_h)
```
```{r}
pl1_cr
pl1_h
pl1_dbh
pl1_dbh1
pl1_dbh2
pl1_dbh3
pl1_nou_cr
pl1_nou_h
pl1_nou_dbh
pl1_nou_dbh1
pl1_nou_dbh2
pl1_nou_dbh3
gmm1_cr
gmm1_h
gmm1_nou_cr
gmm1_nou_h
weibull1_cr
weibull1_h
weibull1_nou_cr
weibull1_nou_h
```

```{r}
check_div <- function(model_data) {
  model_data |>
    filter((grepl("^beta", variable) | grepl("^gamma", variable)) & (rhat > 1.1 | ess_bulk < 300))
}

div_pl1_cr <- check_div(pl1_cr)
div_pl1_h <- check_div(pl1_h)
div_pl1_dbh <- check_div(pl1_dbh)
div_pl1_dbh1 <- check_div(pl1_dbh1)
div_pl1_dbh2 <- check_div(pl1_dbh2)
div_pl1_dbh3 <- check_div(pl1_dbh3)
div_pl1_nou_cr <- check_div(pl1_nou_cr)
div_pl1_nou_h <- check_div(pl1_nou_h)
div_pl1_nou_dbh <- check_div(pl1_nou_dbh)
div_pl1_nou_dbh1 <- check_div(pl1_nou_dbh1)
div_pl1_nou_dbh2 <- check_div(pl1_nou_dbh2)
div_pl1_nou_dbh3 <- check_div(pl1_nou_dbh3)
div_gmm1_cr <- check_div(gmm1_cr)
div_gmm1_h <- check_div(gmm1_h)
div_gmm1_nou_cr <- check_div(gmm1_nou_cr)
div_gmm1_nou_h <- check_div(gmm1_nou_h)
div_weibull1_cr <- check_div(weibull1_cr)
div_weibull1_h <- check_div(weibull1_h)
div_weibull1_nou_cr <- check_div(weibull1_nou_cr)
div_weibull1_nou_h <- check_div(weibull1_nou_h)
```

```{r}
div_pl1_cr
div_pl1_h
div_pl1_dbh
div_pl1_dbh1
div_pl1_dbh2
div_pl1_dbh3
div_pl1_nou_cr
div_pl1_nou_h
div_pl1_nou_dbh
div_pl1_nou_dbh1
div_pl1_nou_dbh2
div_pl1_nou_dbh3
div_gmm1_cr
div_gmm1_h
div_gmm1_nou_cr
div_gmm1_nou_h
div_weibull1_cr
div_weibull1_h
div_weibull1_nou_cr
div_weibull1_nou_h
```

## Checking the problematic species
```{r}
sp_ids <- function(div_data) {
  div_data |>
    filter(grepl("^beta_hat\\[", variable)) |>
    mutate(sp_id = as.numeric(sub("beta_hat\\[([0-9]+),.*\\]", "\\1", variable))) |>
    pull(sp_id) |>
    unique()
}

sp_ids_gmm1_cr <- sp_ids(div_gmm1_cr)
sp_ids_gmm1_h <- sp_ids(div_gmm1_h)
sp_ids_gmm1_nou_cr <- sp_ids(div_gmm1_nou_cr)
sp_ids_gmm1_nou_h <- sp_ids(div_gmm1_nou_h)
sp_ids_weibull1_cr <- sp_ids(div_weibull1_cr)
sp_ids_weibull1_h <- sp_ids(div_weibull1_h)
sp_ids_weibull1_nou_cr <- sp_ids(div_weibull1_nou_cr)
sp_ids_weibull1_nou_h <- sp_ids(div_weibull1_nou_h)

## Print the unique sp_ids for each model
print(sp_ids_gmm1_cr)
print(sp_ids_gmm1_h)
print(sp_ids_gmm1_nou_cr)
print(sp_ids_gmm1_nou_h)
print(sp_ids_weibull1_cr)
print(sp_ids_weibull1_h)
print(sp_ids_weibull1_nou_cr)
print(sp_ids_weibull1_nou_h)

```

Rhat are good, mostly close to 1 (but sp_id 593 is another case).
Model 1-12: mostly have convergent problems in gamma parameter with low ESS.
Non-linear models, in general, have many problematic species with low ESS.
We should increase MCMC iterations, adapt_delta.

# Comparing models
```{r}
loo_tbl

# CR models
cr_models <- loo_tbl |>
  filter(grepl("_cr", model)) |>
  arrange(looic)

print(cr_models)

# H models
h_models <- loo_tbl |>
  filter(grepl("_h", model)) |>
  arrange(looic)

print(h_models)

# DBH models
dbh_models <- loo_tbl |>
  filter(grepl("dbh|dbh1|dbh2|dbh3", model)) |>
  arrange(looic)

# Print the filtered and sorted models
print(dbh_models)
```

- CR models: Weibull incorporating WD is the best model 
- H models: Weibull incorporating WD is the best model 
- DBH: Incorporating CR and H improved DBH estimation accuracy.
- WD slightly effects tree architecture differentiation.

<!-- 
# Table
```{r}
calculate_delta_looic <- function(df) {
  df <- df |>
    arrange(looic) |>
    mutate(ΔLOOIC = looic - min(looic))
  return(df)
}

# Calculate ΔLOOIC for CR models
cr_models_tbl <- loo_tbl |>
  filter(grepl("_cr", model)) |>
  mutate(Wood_Density = ifelse(grepl("nou", model), "Without", "With")) |>
  select(model, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

# Print the CR models table with ΔLOOIC
print(cr_models_tbl)

# Calculate ΔLOOIC for H models
h_models_tbl <- loo_tbl |>
  filter(grepl("_h", model)) |>
  mutate(Wood_Density = ifelse(grepl("nou", model), "Without", "With")) |>
  select(model, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

# Print the H models table with ΔLOOIC
print(h_models_tbl)

# Calculate ΔLOOIC for DBH models
dbh_models_tbl <- loo_tbl |>
  filter(grepl("dbh|dbh1|dbh2|dbh3", model)) |>
  mutate(Wood_Density = ifelse(grepl("nou", model), "Without", "With")) |>
  select(model, Wood_Density, elpd, p_loo, looic) |>
  calculate_delta_looic()

# Print the DBH models table with ΔLOOIC
print(dbh_models_tbl)
```

```{r}
# Define the custom format function for tables
format_table <- function(df, title) {
  formatted_df <- df |>
    rename(
      "Model" = model,
      "Wood Density" = Wood_Density,
      "ELPD" = elpd,
      "pLOO" = p_loo,
      "LOOIC" = looic,
      "ΔLOOIC" = ΔLOOIC
    ) |>
    mutate(across(c(ELPD, pLOO, LOOIC, `ΔLOOIC`), \(x) round(x, 2))) |>
    select(Model, `Wood Density`, ELPD, pLOO, LOOIC, `ΔLOOIC`) |>
    as.data.frame()

  # Print the table with a title
  print(paste0(title, "\n"))
  print(formatted_df)

  # Return the formatted dataframe
  return(formatted_df)
}

# Applying the function to CR, H, and DBH models
cr_models_tbl_formatted <- format_table(cr_models_tbl, "Crown Radius Models")
h_models_tbl_formatted <- format_table(h_models_tbl, "Height Models")
dbh_models_tbl_formatted <- format_table(dbh_models_tbl, "DBH Models")
```
 -->