### 2025-03-26 docs/check_H.qmd

---
title: "Checking H-DBH allometry for angiosperms"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r, include=FALSE}
library(targets)
library(here)
library(tidyverse)
library(scales)
options(width=150)
source("R/figs.R")
```

# Species ID

```{r}
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_h
ang_data_h <- tallo_reduced_lr_df_ang_h |> 
  filter(division == "Angiosperm") |>
  filter(!is.na(h)) |>

ang_data_h

wd_df <- ang_data_h |>
  group_by(sp) |>
  summarize(wd = mean(wd)) |>
  mutate(sp_id = 1:n()) |>
  mutate(wd_s = scale(wd) |> as.numeric()) |>
  dplyr::select(sp_id, sp, wd, wd_s)
head(wd_df)
str(wd_df)
summary(wd_df)

stan_data_nlr_ang_h <- tar_read(stan_data_nlr_ang_h)
str(stan_data_nlr_ang_h)

h_stan_ang <- data.frame(
  sp_id = as.factor(stan_data_nlr_ang_h$jj),
  # dbh = stan_data_nlr_ang_h$x,
  wd_s = stan_data_nlr_ang_h$u[stan_data_nlr_ang_h$jj, 2])

head(h_stan_ang)
# id_66 <- wd_df |> filter(wd_df$sp_id == "66")
# id_66

h_stan_ang <- h_stan_ang |>
  group_by(sp_id, wd_s) |>
  summarise(tree_count = n(), .groups = "drop")

head(h_stan_ang)
str(h_stan_ang)
# id_66_stan <- h_stan_ang |> filter(h_stan_ang$sp_id == "66")
# id_66_stan

h_ang <- h_stan_ang |>
  mutate(sp_id = as.integer(as.character(sp_id))) |>
  left_join(wd_df |> mutate(sp_id = as.integer(sp_id)) |> select(sp_id, sp), by = "sp_id")
head(h_ang)
str(h_ang)

```

# Extracting sp-specific parameter

```{r}
fit_nlr_nou_summary_weibull_ang_h <- tar_read(fit_nlr_nou_summary_weibull_ang_h)
fit_nlr_nou_summary_weibull_ang_h

create_species_model_df <- function(beta_data, dependent_variable, functional_form, sp_list) {
  if (functional_form == "Weibull" || functional_form == "gMM") {
    a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
    a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q97.5))
    a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q2.5))

    b_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
    b_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q97.5)
    b_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q2.5)

    k_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
    k_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q97.5)
    k_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q2.5)

    ci_95 <- data.frame(
      sp = sp_list,
      a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
      b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
      k = paste0(round(k_median, 2), " (", round(k_lower, 2), ", ", round(k_upper, 2), ")"),
      Dependent_variable = dependent_variable,
      stringsAsFactors = FALSE
    )
  } else if (functional_form == "Power-Law") {
    a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50))
    a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q97.5))
    a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q2.5))

    b_median <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
    b_upper <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q97.5)
    b_lower <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q2.5)

    c_median <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)
    c_upper <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q97.5)
    c_lower <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q2.5)

    ci_95 <- data.frame(
      sp = sp_list,
      a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
      b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
      c = paste0(round(c_median, 2), " (", round(c_lower, 2), ", ", round(c_upper, 2), ")"),
      Dependent_variable = dependent_variable,
      stringsAsFactors = FALSE
    )
  }

  return(ci_95)
}

h_posterior_df <- create_species_model_df(fit_nlr_nou_summary_weibull_ang_h, "Tree Height", "Weibull", h_ang)

head(h_posterior_df)

h_posterior_df <- h_posterior_df |>
  rename(sp = sp.sp)

head(h_posterior_df)

head(h_posterior_df)
ang_data_h <- tallo_reduced_lr_df_ang_h |> filter(division == "Angiosperm")
ang_data_h
summary(ang_data_h)

large_tree_species <- ang_data_h |>
  group_by(sp) |>
  filter(any(dbh > 150)) |>
  ungroup()

large_tree_summary <- large_tree_species |>
  group_by(sp) |>
  summarize(
    min_dbh = min(dbh, na.rm = TRUE),
    max_dbh = max(dbh, na.rm = TRUE),
    mean_dbh = mean(dbh, na.rm = TRUE),
    count = n()
  ) |>
  arrange(desc(max_dbh))

top_30_species <- large_tree_summary |>
  slice(1:30)

top_30_species

top_30_data <- ang_data_h |>
  filter(sp %in% top_30_species$sp) |>
  select(tree_id, sp, dbh, h)

sub_datasets <- top_30_data |>
  group_split(sp)

names(sub_datasets) <- unique(top_30_data$sp)
```

# Function
```{r}
generate_plot <- function(sub_datasets, h_posterior_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Get posterior parameters for the species
    species_post <- h_posterior_df |>
      filter(sp == species_name)

    if (nrow(species_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Extract parameters
    a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
    log_a <- log(a)
    b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))
    k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

    dbh <- species_data$dbh

    # Compute the fitted height values
    fitted_log_y <- log_a + log(1 - exp(-b * (dbh ^ k)))
    fitted_height <- exp(fitted_log_y)

    # Create a dataframe for the fitted curve
    fitted_curve <- data.frame(
      DBH = dbh,
      H = fitted_height,
      sp = species_name,
      Source = "Fitted"
    )

    # Combine observed and fitted data
    combined_species_data <- species_data |>
      mutate(sp = species_name, Source = "Observed") |>
      select(DBH = dbh, H = h, sp, Source) |>
      bind_rows(fitted_curve)

    # Append to the combined dataset
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  p <- ggplot(combined_data, aes(x = DBH, y = H, color = Source)) +
    geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#72b6e3", linewidth = 0.6) +
    facet_wrap(~ sp, scales = "free", ncol = 4) +
    labs(
      x = "DBH (cm)",
      y = "Tree Height (m)"
    ) +
    my_theme()

  return(p)
}

generate_plot_log <- function(sub_datasets, h_posterior_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Get posterior parameters for the species
    species_post <- h_posterior_df |>
      filter(sp == species_name)

    if (nrow(species_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Extract parameters
    a <- as.numeric(sub(" \\(.*\\)", "", species_post$a))
    log_a <- log(a)
    b <- as.numeric(sub(" \\(.*\\)", "", species_post$b))
    k <- as.numeric(sub(" \\(.*\\)", "", species_post$k))

    dbh <- species_data$dbh

    # Compute the fitted height values
    fitted_log_y <- log_a + log(1 - exp(-b * (dbh ^ k)))
    fitted_height <- exp(fitted_log_y)

    # Create a dataframe for the fitted curve
    fitted_curve <- data.frame(
      DBH = dbh,
      H = fitted_height,
      sp = species_name,
      Source = "Fitted"
    )

    # Combine observed and fitted data
    combined_species_data <- species_data |>
      mutate(sp = species_name, Source = "Observed") |>
       dplyr::select(DBH = dbh, H = h, sp, Source) |>
      bind_rows(fitted_curve)

    # Append to the combined dataset
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  p <- ggplot(combined_data, aes(x = DBH, y = H, color = Source)) +
    geom_point(data = combined_data |> filter(Source == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Source == "Fitted"), aes(group = sp), color = "#72b6e3", linewidth = 0.6) +
    facet_wrap(~ sp, scales = "free", ncol = 4) +
    labs(
      x = "DBH (cm)",
      y = "Tree Height (m)"
    ) +
    scale_x_log10() +
    scale_y_log10() +
    my_theme()

  return(p)
}
```

# Plotting

```{r, fig.width=14, fig.height=20}
p <- generate_plot(sub_datasets, h_posterior_df)
p

p_log <- generate_plot_log(sub_datasets, h_posterior_df)
p_log
```

# SPECIES POSTERIOR DATAFRAME

```{r}
fit_nlr_nou_summary_weibull_ang_h <- tar_read(fit_nlr_nou_summary_weibull_ang_h)
fit_nlr_nou_summary_weibull_gym_h <- tar_read(fit_nlr_nou_summary_weibull_gym_h)
fit_lr_nou_summary_pl_ang_cr <- tar_read(fit_lr_nou_summary_pl_ang_cr)
fit_nlr_nou_summary_gmm_gym_cr <- tar_read(fit_nlr_nou_summary_gmm_gym_cr)
fit_lr_nou_summary_pl_ang_dbh <- tar_read(fit_lr_nou_summary_pl_ang_dbh)
fit_lr_nou_summary_pl_gym_dbh <- tar_read(fit_lr_nou_summary_pl_gym_dbh)
tallo_reduced_lr_df_ang_h <- tar_read(tallo_reduced_lr_df_ang_h)
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_dbh <- tar_read(tallo_reduced_lr_df_ang_dbh)
stan_data_nlr_ang_h <- tar_read(stan_data_nlr_ang_h)
stan_data_nlr_gym_h <- tar_read(stan_data_nlr_gym_h)
stan_data_lr_ang_cr <- tar_read(stan_data_lr_ang_cr)
stan_data_nlr_gym_cr <- tar_read(stan_data_nlr_gym_cr)
stan_data_lr_ang_dbh <- tar_read(stan_data_lr_ang_dbh)
stan_data_lr_gym_dbh <- tar_read(stan_data_lr_gym_dbh)
```

```{r}
str(stan_data_nlr_ang_h)
str(stan_data_nlr_gym_h)
str(stan_data_lr_ang_cr)
str(stan_data_nlr_gym_cr)
str(stan_data_lr_ang_dbh)
str(stan_data_lr_gym_dbh)

  filter_species <- function(dataset, division) {
    dataset |> 
      filter(division == division, !is.na(dbh), !is.na(cr), !is.na(h)) |> 
      group_by(sp) |> 
      summarise(count = n()) |> 
      filter(count >= 20) |> 
      pull(sp)
  }
  create_species_model_df <- function(beta_data, dependent_variable, functional_form, sp_list) {
    if (functional_form == "Weibull" || functional_form == "gMM") {
      a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
      a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q97.5))
      a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q2.5))

      b_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
      b_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q97.5)
      b_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q2.5)

      k_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
      k_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q97.5)
      k_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q2.5)

      ci_95 <- data.frame(
        sp = sp_list,
        a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
        b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
        k = paste0(round(k_median, 2), " (", round(k_lower, 2), ", ", round(k_upper, 2), ")"),
        Dependent_variable = dependent_variable,
        stringsAsFactors = FALSE
      )
    } else if (functional_form == "Power-Law") {
      a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50))
      a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q97.5))
      a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q2.5))

      b_median <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
      b_upper <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q97.5)
      b_lower <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q2.5)

      c_median <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)
      c_upper <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q97.5)
      c_lower <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q2.5)

      ci_95 <- data.frame(
        sp = sp_list,
        a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
        b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
        c = paste0(round(c_median, 2), " (", round(c_lower, 2), ", ", round(c_upper, 2), ")"),
        Dependent_variable = dependent_variable,
        stringsAsFactors = FALSE
      )
    }

    return(ci_95)
  }
  # Function to process datasets and align species IDs with Stan data
 process_dataset <- function(dataset, stan_data, division_filter) {
    # Filter the dataset by division
    dataset <- dataset |> filter(division == division_filter)

    # Prepare species data
    sp_data <- dataset |>
      group_by(sp) |>
      summarise(wd = mean(wd, na.rm = TRUE)) |>
      arrange(sp) |>
      mutate(sp_id = row_number()) |>
      mutate(wd_s = scale(wd) |> as.numeric()) |>
      select(sp_id, sp, wd, wd_s)

    # Verify Stan data structure
    if (!all(c("jj", "u") %in% names(stan_data))) {
      stop("Stan data must contain 'jj' and 'u'.")
    }

    # Handle `nlr` structure: `u` has rows indexed by `jj`
    if (nrow(stan_data$u) >= max(stan_data$jj, na.rm = TRUE)) {
      stan_processed <- data.frame(
        sp_id = as.factor(stan_data$jj),
        wd_s = stan_data$u[stan_data$jj, 2]  # Extract `wd_s` for species indices
      )
    } 
    # Handle `lr` structure: `u` has two rows (variables) and columns for species
    else if (nrow(stan_data$u) == 2) {
      stan_processed <- data.frame(
        sp_id = 1:ncol(stan_data$u),
        wd_s = stan_data$u[2, ]  # Use the second row for `wd_s`
      )
    } 
    # Error handling for unexpected structures
    else {
      stop("Unexpected structure for 'u' in Stan data.")
    }

    # Add species names and calculate tree counts
    stan_processed <- stan_processed |>
      group_by(sp_id, wd_s) |>
      summarise(tree_count = n(), .groups = "drop") |>
      mutate(sp_id = as.integer(as.character(sp_id))) |>
      left_join(sp_data, by = "sp_id")

    return(list(sp_data = sp_data, stan_processed = stan_processed))
  }

```

```{r}
h_ang <- process_dataset(tallo_reduced_lr_df_ang_h, stan_data_nlr_ang_h, "Angiosperm")

h_ang

h_gym <- process_dataset(tallo_reduced_lr_df_ang_h, stan_data_nlr_gym_h, "Gymnosperm")
h_gym

cr_ang <- process_dataset(tallo_reduced_lr_df_ang_cr, stan_data_lr_ang_cr, "Angiosperm")

cr_ang

cr_gym <- process_dataset(tallo_reduced_lr_df_ang_cr, stan_data_nlr_gym_cr, "Gymnosperm")
cr_gym

# dbh_ang <- process_dataset(tallo_reduced_lr_df_ang_dbh, stan_data_lr_ang_dbh, "Angiosperm")
# dbh_ang

# dbh_gym <- process_dataset(tallo_reduced_lr_df_ang_dbh, stan_data_lr_gym_dbh, "Gymnosperm")
# dbh_gym
ang_dbh_species <- filter_species(tallo_reduced_lr_df_ang_dbh, "Angiosperm")
gym_dbh_species <- filter_species(tallo_reduced_lr_df_ang_dbh, "Gymnosperm")

dbh_ang <- process_dataset(tallo_reduced_lr_df_ang_dbh |> filter(sp %in% ang_dbh_species), stan_data_lr_ang_dbh, "Angiosperm")

dbh_gym <- process_dataset(tallo_reduced_lr_df_ang_dbh |> filter(sp %in% gym_dbh_species), stan_data_lr_gym_dbh, "Gymnosperm")

```

```{r}
h_df_ang <- create_species_model_df(fit_nlr_nou_summary_weibull_ang_h, "Tree Height", "Weibull", h_ang$sp)
head(h_df_ang)

h_df_gym <- create_species_model_df(fit_nlr_nou_summary_weibull_gym_h, "Tree Height", "Weibull", h_gym$sp)
head(h_df_gym)

cr_df_ang <- create_species_model_df(fit_lr_nou_summary_pl_ang_cr, "Crown Radius", "Power-Law", cr_ang$sp)
head(cr_df_ang)

cr_df_gym <- create_species_model_df(fit_nlr_nou_summary_gmm_gym_cr, "Crown Radius", "gMM", cr_gym$sp)
head(cr_df_gym)

dbh_df_ang <- create_species_model_df(fit_lr_nou_summary_pl_ang_dbh, "DBH", "Power-Law", dbh_ang$sp)
head(dbh_df_ang)

dbh_df_gym <- create_species_model_df(fit_lr_nou_summary_pl_gym_dbh, "DBH", "Power-Law", dbh_gym$sp)
head(dbh_df_gym)

```

```{r}
sp_posterior_df <- bind_rows(
  h_df_ang |> mutate(Division = "Angiosperm"),
  h_df_gym |> mutate(Division = "Gymnosperm"),
  cr_df_ang |> mutate(Division = "Angiosperm"),
  cr_df_gym |> mutate(Division = "Gymnosperm"),
  dbh_df_ang |> mutate(Division = "Angiosperm"),
  dbh_df_gym |> mutate(Division = "Gymnosperm")
)

head(sp_posterior_df)
```

```{r}
  filter_species <- function(dataset, division) {
    dataset |> 
      filter(division == division, !is.na(dbh), !is.na(cr), !is.na(h)) |> 
      group_by(sp) |> 
      summarise(count = n()) |> 
      filter(count >= 20) |> 
      pull(sp)
  }

  # Helper function to extract posterior summaries and prepare species-specific output
  create_species_model_df <- function(beta_data, dependent_variable, functional_form, sp_list) {
    if (functional_form == "Weibull" || functional_form == "gMM") {
      a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
      a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q97.5))
      a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q2.5))

      b_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
      b_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q97.5)
      b_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q2.5)

      k_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
      k_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q97.5)
      k_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q2.5)

      # Ensure species list and parameters match
      if (length(sp_list) != length(a_median)) {
        stop("Mismatch in species list and parameter lengths.")
      }

      ci_95 <- data.frame(
        sp = sp_list,
        a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
        b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
        k = paste0(round(k_median, 2), " (", round(k_lower, 2), ", ", round(k_upper, 2), ")"),
        Dependent_variable = dependent_variable,
        stringsAsFactors = FALSE
      )
    } else if (functional_form == "Power-Law") {
      a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50))
      a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q97.5))
      a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q2.5))

      b_median <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
      b_upper <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q97.5)
      b_lower <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q2.5)

      c_median <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)
      c_upper <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q97.5)
      c_lower <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q2.5)

      # Ensure species list and parameters match
      if (length(sp_list) != length(a_median)) {
        stop("Mismatch in species list and parameter lengths.")
      }

      ci_95 <- data.frame(
        sp = sp_list,
        a = paste0(round(a_median, 2), " (", round(a_lower, 2), ", ", round(a_upper, 2), ")"),
        b = paste0(round(b_median, 2), " (", round(b_lower, 2), ", ", round(b_upper, 2), ")"),
        c = paste0(round(c_median, 2), " (", round(c_lower, 2), ", ", round(c_upper, 2), ")"),
        Dependent_variable = dependent_variable,
        stringsAsFactors = FALSE
      )
    }

    return(ci_95)
  }

  # Function to process datasets and align species IDs with Stan data
  process_dataset <- function(dataset, stan_data, division_filter) {
    # Filter the dataset by division
    dataset <- dataset |> filter(division == division_filter)

    # Prepare species data
    sp_data <- dataset |>
      group_by(sp) |>
      summarise(wd = mean(wd, na.rm = TRUE)) |>
      arrange(sp) |>
      mutate(sp_id = row_number()) |>
      mutate(wd_s = scale(wd) |> as.numeric()) |>
      select(sp_id, sp, wd, wd_s)

    # Verify Stan data structure
    if (!all(c("jj", "u") %in% names(stan_data))) {
      stop("Stan data must contain 'jj' and 'u'.")
    }

    # Handle `nlr` structure: `u` has rows indexed by `jj`
    if (nrow(stan_data$u) >= max(stan_data$jj, na.rm = TRUE)) {
      stan_processed <- data.frame(
        sp_id = as.factor(stan_data$jj),
        wd_s = stan_data$u[stan_data$jj, 2]  # Extract `wd_s` for species indices
      )
    } else if (nrow(stan_data$u) == 2) {
      stan_processed <- data.frame(
        sp_id = 1:ncol(stan_data$u),
        wd_s = stan_data$u[2, ]  # Use the second row for `wd_s`
      )
    } else {
      stop("Unexpected structure for 'u' in Stan data.")
    }

    # Add species names and calculate tree counts
    stan_processed <- stan_processed |>
      group_by(sp_id, wd_s) |>
      summarise(tree_count = n(), .groups = "drop") |>
      mutate(sp_id = as.integer(as.character(sp_id))) |>
      left_join(sp_data, by = "sp_id")

    return(list(sp_data = sp_data, stan_processed = stan_processed))
  }

  # Filter species
  ang_dbh_species <- filter_species(tallo_reduced_lr_df_ang_dbh, "Angiosperm")
  gym_dbh_species <- filter_species(tallo_reduced_lr_df_ang_dbh, "Gymnosperm")

  # Process datasets for Angiosperms and Gymnosperms
  h_ang <- process_dataset(tallo_reduced_lr_df_ang_h, stan_data_nlr_ang_h, "Angiosperm")
  h_gym <- process_dataset(tallo_reduced_lr_df_ang_h, stan_data_nlr_gym_h, "Gymnosperm")
  cr_ang <- process_dataset(tallo_reduced_lr_df_ang_cr, stan_data_lr_ang_cr, "Angiosperm")
  cr_gym <- process_dataset(tallo_reduced_lr_df_ang_cr, stan_data_nlr_gym_cr, "Gymnosperm")
  dbh_ang <- process_dataset(tallo_reduced_lr_df_ang_dbh |> filter(sp %in% ang_dbh_species), stan_data_lr_ang_dbh, "Angiosperm")
  dbh_gym <- process_dataset(tallo_reduced_lr_df_ang_dbh |> filter(sp %in% gym_dbh_species), stan_data_lr_gym_dbh, "Gymnosperm")

  # Generate posterior dataframes
  h_df_ang <- create_species_model_df(fit_nlr_nou_summary_weibull_ang_h, "Tree Height", "Weibull", h_ang$sp)
  h_df_gym <- create_species_model_df(fit_nlr_nou_summary_weibull_gym_h, "Tree Height", "Weibull", h_gym$sp)
  cr_df_ang <- create_species_model_df(fit_lr_nou_summary_pl_ang_cr, "Crown Radius", "Power-Law", cr_ang$sp)
  cr_df_gym <- create_species_model_df(fit_nlr_nou_summary_gmm_gym_cr, "Crown Radius", "gMM", cr_gym$sp)
  dbh_df_ang <- create_species_model_df(fit_lr_nou_summary_pl_ang_dbh, "DBH", "Power-Law", dbh_ang$sp)
  dbh_df_gym <- create_species_model_df(fit_lr_nou_summary_pl_gym_dbh, "DBH", "Power-Law", dbh_gym$sp)

  # # Combine results
  # sp_posterior_df <- bind_rows(
  #   h_df_ang |> mutate(Division = "Angiosperm"),
  #   h_df_gym |> mutate(Division = "Gymnosperm"),
  #   cr_df_ang |> mutate(Division = "Angiosperm"),
  #   cr_df_gym |> mutate(Division = "Gymnosperm"),
  #   dbh_df_ang |> mutate(Division = "Angiosperm"),
  #   dbh_df_gym |> mutate(Division = "Gymnosperm")
  # )

```

```{r}

```