### 2025-03-26 docs/check_H_gym.qmd

---
title: "Plotting H-DBH allometry"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r, include=FALSE}
library(targets)
library(here)
library(tidyverse)
library(scales)
options(width=150)
source("R/figs.R")
```

# Extract parameters of gymnosperms H-DBH from 3 functional forms

```{r}
create_species_model_df <- function(beta_data, dependent_variable, functional_form, sp_list) {
  if (functional_form == "Weibull" || functional_form == "gMM") {
    a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q50))
    a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q97.5))
    a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[\\d+,1\\]")) |> pull(q2.5))

    b_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q50)
    b_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q97.5)
    b_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,2\\]")) |> pull(q2.5)

    k_median <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q50)
    k_upper <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q97.5)
    k_lower <- beta_data |> filter(str_detect(variable, "beta\\[\\d+,3\\]")) |> pull(q2.5)

    ci_95 <- data.frame(
      sp = sp_list,
      a = paste0(format(a_median, scientific = FALSE), " (", format(a_lower, scientific = FALSE), ", ", format(a_upper, scientific = FALSE), ")"),
      b = paste0(format(b_median, scientific = FALSE), " (", format(b_lower, scientific = FALSE), ", ", format(b_upper, scientific = FALSE), ")"),
      k = paste0(format(k_median, scientific = FALSE), " (", format(k_lower, scientific = FALSE), ", ", format(k_upper, scientific = FALSE), ")"),
      Dependent_variable = dependent_variable,
      stringsAsFactors = FALSE
    )
  } else if (functional_form == "Power-Law") {
    a_median <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50))
    a_upper <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q97.5))
    a_lower <- exp(beta_data |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q2.5))

    b_median <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
    b_upper <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q97.5)
    b_lower <- beta_data |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q2.5)

    c_median <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)
    c_upper <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q97.5)
    c_lower <- beta_data |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q2.5)

  ci_95 <- data.frame(
        sp = sp_list,
        a = paste0(format(a_median, scientific = FALSE), " (", format(a_lower, scientific = FALSE), ", ", format(a_upper, scientific = FALSE), ")"),
        b = paste0(format(b_median, scientific = FALSE), " (", format(b_lower, scientific = FALSE), ", ", format(b_upper, scientific = FALSE), ")"),
        c = paste0(format(c_median, scientific = FALSE), " (", format(c_lower, scientific = FALSE), ", ", format(c_upper, scientific = FALSE), ")"),
        Dependent_variable = dependent_variable,
        stringsAsFactors = FALSE
    )
  }

  return(ci_95)
}
# Function to process datasets and align species IDs with Stan data
process_dataset <- function(dataset, stan_data, division_filter) {
# Filter the dataset by division
dataset <- dataset |> filter(division == division_filter)

# Prepare species data
sp_data <- dataset |>
  group_by(sp) |>
  summarise(wd = mean(wd, na.rm = TRUE)) |>
  arrange(sp) |>
  mutate(sp_id = row_number()) |>
  mutate(wd_s = scale(wd) |> as.numeric()) |>
  dplyr::select(sp_id, sp, wd, wd_s)

if (!all(c("jj", "u") %in% names(stan_data))) {
  stop("Stan data must contain 'jj' and 'u'.")
}

# Handle `nlr` structure: `u` has rows indexed by `jj`
if (nrow(stan_data$u) >= max(stan_data$jj, na.rm = TRUE)) {
  stan_processed <- data.frame(
    sp_id = as.factor(stan_data$jj),
    wd_s = stan_data$u[stan_data$jj, 2]  # Extract `wd_s` for species indices
  )
} 
# Handle `lr` structure: `u` has two rows (variables) and columns for species
else if (nrow(stan_data$u) == 2) {
  stan_processed <- data.frame(
    sp_id = 1:ncol(stan_data$u),
    wd_s = stan_data$u[2, ]  # Use the second row for `wd_s`
  )
} 
# Error handling for unexpected structures
else {
  stop("Unexpected structure for 'u' in Stan data.")
}

# Add species names and calculate tree counts
stan_processed <- stan_processed |>
  group_by(sp_id, wd_s) |>
  summarise(tree_count = n(), .groups = "drop") |>
  mutate(sp_id = as.integer(as.character(sp_id))) |>
  left_join(sp_data, by = "sp_id")

return(list(sp_data = sp_data, stan_processed = stan_processed))
}
```

```{r}
stan_data_nlr_gym_h <- tar_read(stan_data_nlr_gym_h)
stan_data_lr_gym_h <- tar_read(stan_data_lr_gym_h)

tallo_reduced_nlr_df_gym_h <- tar_read(tallo_reduced_nlr_df_gym_h)
tallo_reduced_nlr_df_gym_h

tallo_reduced_lr_df_gym_h <- tar_read(tallo_reduced_lr_df_gym_h)
tallo_reduced_lr_df_gym_h

fit_nlr_nou_summary_weibull_gym_h <- tar_read(fit_nlr_nou_summary_weibull_gym_h)
fit_nlr_nou_summary_weibull_gym_h

fit_lr_nou_summary_pl_gym_h <- tar_read(fit_lr_nou_summary_pl_gym_h)
fit_lr_nou_summary_pl_gym_h

fit_nlr_nou_summary_gmm_gym_h <- tar_read(fit_nlr_nou_summary_gmm_gym_h)
fit_nlr_nou_summary_gmm_gym_h

h_wb_gym <- process_dataset(tallo_reduced_nlr_df_gym_h, stan_data_nlr_gym_h, "Gymnosperm")
head(h_wb_gym)

h_pl_gym <- process_dataset(tallo_reduced_lr_df_gym_h, stan_data_lr_gym_h, "Gymnosperm")
head(h_pl_gym)

h_gmm_gym <- process_dataset(tallo_reduced_nlr_df_gym_h, stan_data_nlr_gym_h, "Gymnosperm")
head(h_gmm_gym)

h_wb_df_gym <- create_species_model_df(fit_nlr_nou_summary_weibull_gym_h, "Tree Height", "Weibull", h_wb_gym$sp_data$sp)
head(h_wb_df_gym)

h_pl_df_gym <- create_species_model_df(fit_lr_nou_summary_pl_gym_h, "Tree Height", "Power-Law", h_pl_gym$sp_data$sp)
head(h_pl_df_gym)

h_gmm_df_gym <- create_species_model_df(fit_nlr_nou_summary_gmm_gym_h, "Tree Height", "gMM", h_gmm_gym$sp_data$sp)
head(h_gmm_df_gym)

```

```{r}
tallo_reduced_lr_df_gym_h <- tar_read(tallo_reduced_lr_df_gym_h)
tallo_reduced_lr_df_gym_h

h_data <- tallo_reduced_lr_df_gym_h |> 
    filter(division == "Gymnosperm") |>
    filter(!is.na(h)) |>
    group_by(sp) |>
    ungroup()

h_data
num_species = 10

species_names <- sample(unique(h_data$sp), min(num_species, n_distinct(h_data$sp)))

# Split the dataset by species
sub_datasets <- h_data |>
  filter(sp %in% species_names) |>
  group_split(sp)

# Assign species names as dataset names
names(sub_datasets) <- species_names

```

```{r}
generate_plot <- function(sub_datasets, gmm_df, weibull_df, pl_df) {
  combined_data <- data.frame()

  for (i in seq_along(sub_datasets)) {
    species_data <- sub_datasets[[i]]
    species_name <- names(sub_datasets)[i]

    # Filter data for each model
    gmm_post <- gmm_df |> filter(sp == species_name)
    weibull_post <- weibull_df |> filter(sp == species_name)
    pl_post <- pl_df |> filter(sp == species_name)

    if (nrow(gmm_post) == 0 || nrow(weibull_post) == 0 || nrow(pl_post) == 0) {
      warning(paste("No posterior data for species:", species_name))
      next
    }

    # Use only observed DBH range for predictions
    dbh_values <- species_data$dbh

    # Power-Law model
    a_pl <- as.numeric(sub(" \\(.*\\)", "", pl_post$a))
    b_pl <- as.numeric(sub(" \\(.*\\)", "", pl_post$b))
    pl_pred <- a_pl * (dbh_values^b_pl)

    # gMM model
    a_gmm <- as.numeric(sub(" \\(.*\\)", "", gmm_post$a))
    b_gmm <- as.numeric(sub(" \\(.*\\)", "", gmm_post$b))
    k_gmm <- as.numeric(sub(" \\(.*\\)", "", gmm_post$k))
    gmm_pred <- a_gmm * (dbh_values^b_gmm) / (k_gmm + dbh_values^b_gmm)

    # Weibull model
    a_wb <- as.numeric(sub(" \\(.*\\)", "", weibull_post$a))
    b_wb <- as.numeric(sub(" \\(.*\\)", "", weibull_post$b))
    k_wb <- as.numeric(sub(" \\(.*\\)", "", weibull_post$k))
    wb_pred <- a_wb * (1 - exp(-b_wb * dbh_values^k_wb))

    # Create fitted curves for each model
    model_predictions <- list()
    model_predictions[["Power-Law"]] <- data.frame(DBH = dbh_values, H = pl_pred, sp = species_name, Model = "Power-Law")
    model_predictions[["gMM"]] <- data.frame(DBH = dbh_values, H = gmm_pred, sp = species_name, Model = "gMM")
    model_predictions[["Weibull"]] <- data.frame(DBH = dbh_values, H = wb_pred, sp = species_name, Model = "Weibull")

    # Combine observed and fitted data
    fitted_data <- bind_rows(model_predictions)
    observed_data <- species_data |>
      mutate(DBH = dbh, H = h, Model = "Observed", sp = species_name) |>
      dplyr::select(DBH, H, sp, Model)

    combined_species_data <- bind_rows(fitted_data, observed_data)
    combined_data <- bind_rows(combined_data, combined_species_data)
  }

  # Create the plot
  p <- ggplot(combined_data, aes(x = DBH, y = H, color = Model)) +
    geom_point(data = combined_data |> filter(Model == "Observed"), alpha = 0.5, color = "gray") +
    geom_line(data = combined_data |> filter(Model != "Observed"), aes(linetype = Model), linewidth = 0.6, linetype = "solid") +
    facet_wrap(~ sp, scales = "free", ncol = 3) +
    labs(
      x = "DBH (cm)",
      y = "Tree Height (m)"
    ) +
    scale_color_manual(values = c("gMM" = "#1b9e77", "Weibull" = "#0f92e9", "Power-Law" = "#FF5733", "Observed" = "gray")) +
    my_theme()

  return(p)
}
```

```{r, fig.width=30, fig.height=60, out.width='100%', out.height='100%'}
p <- generate_plot(sub_datasets, h_gmm_df_gym, h_wb_df_gym, h_pl_df_gym)
p
```