### 2025-03-26 docs/com.qmd

---
title: "Model comparison figure"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
# Required Libraries
library(targets)
library(tidyverse)
library(scales)
library(MASS)
library(viridis)
library(patchwork)
library(ggplot2)
source("R/figs.R")
```

# Read in data

```{r}
# Load Data for H-DBH
stan_data_lr_h <- tar_read(stan_data_lr_h)
stan_data_nlr_h <- tar_read(stan_data_nlr_h)
gmm_summary_h <- tar_read(fit_nlr_summary_gmm1_h)
wb_summary_h <- tar_read(fit_nlr_summary_weibull1_h)
pl_summary_h <- tar_read(fit_lr2_summary_power_law1_h)

# Load Data for CR-DBH
stan_data_lr_cr <- tar_read(stan_data_lr_cr)
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
gmm_summary_cr <- tar_read(fit_nlr_summary_gmm1_cr)
wb_summary_cr <- tar_read(fit_nlr_summary_weibull1_cr)
pl_summary_cr <- tar_read(fit_lr2_summary_power_law1_cr)
```

# Extract scaling parameters
```{r}
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)

data2 <- tallo_wd_df_200 |>
  filter(!is.na(dbh), !is.na(h), !is.na(cr), !is.na(wd_s)) |>
  mutate(log_dbh = log(dbh),
         log_h = log(h),
         log_cr = log(cr)) |>
  mutate(log_dbh_s = scale(log_dbh) |> as.numeric(),
         log_h_s = scale(log_h) |> as.numeric(),
         log_cr_s = scale(log_cr) |> as.numeric()) |>
  mutate(dbh_s = scale(dbh) |> as.numeric())

# Extract scaling parameters for unscaling
log_dbh_sd <- sd(data2$log_dbh)
log_dbh_mean <- mean(data2$log_dbh)
log_h_sd <- sd(data2$log_h)
log_h_mean <- mean(data2$log_h)
log_cr_sd <- sd(data2$log_cr)
log_cr_mean <- mean(data2$log_cr)
```

# Function to Compute Fitted Values

```{r}
compute_fitted_values <- function(model_summary, x_seq, log_y_mean, log_y_sd, log_x_sd, log_x_mean, model_type, log_x_seq = NULL) {
  gamma <- model_summary |> filter(str_detect(variable, "gamma"))

  if (model_type == "pl") {
    # Power-Law model: use log_x_seq for predictions
    log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
    b <- gamma |> filter(variable == "gamma[2,1]") |> pull(q50)
    log_y_pred_scaled <- log_a + b * log_x_seq

    # Unscale the predicted log-transformed y
    log_y_pred <- log_y_pred_scaled * log_y_sd + log_y_mean

    # Unscale the x variable and back-transform to the original scale
    log_x_unscaled <- (log_x_seq * log_x_sd) + log_x_mean
    fitted_values <- tibble(y = exp(log_y_pred), x = exp(log_x_unscaled))

  } else if (model_type == "gmm") {
    # Generalized Michaelis-Menten model
    log_log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
    b <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)
    k <- gamma |> filter(variable == "gamma[1,3]") |> pull(q50)

    log_a <- exp(log_log_a)
    b <- exp(b)
    k <- exp(k)

    log_y_pred <- log_a + b * log(x_seq) - log(k + x_seq^b)
    fitted_values <- tibble(y = exp(log_y_pred), x = x_seq)

  } else if (model_type == "weibull") {
    # Weibull model
    log_log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(q50)
    b <- gamma |> filter(variable == "gamma[1,2]") |> pull(q50)
    k <- gamma |> filter(variable == "gamma[1,3]") |> pull(q50)

    log_a <- exp(log_log_a)
    b <- exp(b)
    k <- exp(k)

    log_y_pred <- log_a + log(1 - exp(-b * (x_seq ^ k)))
    fitted_values <- tibble(y = exp(log_y_pred), x = x_seq)

  } else {
    stop("Invalid model type specified.")
  }

  return(fitted_values)
}
```

# Extract fitted values
```{r}
# Define sequences for Power-Law x variables (scaled log x)
scaled_x_range_h <- stan_data_lr_h$log_x[, 2] |> range()
log_x_seq_h <- seq(scaled_x_range_h[1], scaled_x_range_h[2], length.out = 100)

scaled_x_range_cr <- stan_data_lr_cr$log_x[, 2] |> range()
log_x_seq_cr <- seq(scaled_x_range_cr[1], scaled_x_range_cr[2], length.out = 100)

# Define sequences for gMM and Weibull x variables (original x)
x_nlr_h <- stan_data_nlr_h$x
x_nlr_range_h <- range(x_nlr_h)
x_nlr_seq_h <- seq(x_nlr_range_h[1], x_nlr_range_h[2], length.out = 100)

x_nlr_cr <- stan_data_nlr_cr$x
x_nlr_range_cr <- range(x_nlr_cr)
x_nlr_seq_cr <- seq(x_nlr_range_cr[1], x_nlr_range_cr[2], length.out = 100)

# Compute fitted values for Power-Law model (unscaled)
df_pl_h <- compute_fitted_values(pl_summary_h, NULL, log_h_mean, log_h_sd, log_dbh_sd, log_dbh_mean, model_type = "pl", log_x_seq = log_x_seq_h)
df_pl_cr <- compute_fitted_values(pl_summary_cr, NULL, log_cr_mean, log_cr_sd, log_dbh_sd, log_dbh_mean, model_type = "pl", log_x_seq = log_x_seq_cr)

# Compute fitted values for gMM model (original x)
df_gmm_h <- compute_fitted_values(gmm_summary_h, x_nlr_seq_h, log_h_mean, log_h_sd, log_dbh_sd, log_dbh_mean, model_type = "gmm")
df_gmm_cr <- compute_fitted_values(gmm_summary_cr, x_nlr_seq_cr, log_cr_mean, log_cr_sd, log_dbh_sd, log_dbh_mean, model_type = "gmm")

# Compute fitted values for Weibull model (original x)
df_wb_h <- compute_fitted_values(wb_summary_h, x_nlr_seq_h, log_h_mean, log_h_sd, log_dbh_sd, log_dbh_mean, model_type = "weibull")
df_wb_cr <- compute_fitted_values(wb_summary_cr, x_nlr_seq_cr, log_cr_mean, log_cr_sd, log_dbh_sd, log_dbh_mean, model_type = "weibull")
```

# Plotting

## Plotting function
```{r}
# Plotting function
plot_model_comparison <- function(data, df_pl, df_gmm, df_wb, x_var, y_var, x_label, y_label, log_scale = TRUE) {
  # Compute density for the scatter plot
  if (log_scale) {
    # Log-scale density calculation
    dens <- with(data, kde2d(log10(data[[x_var]]), log10(data[[y_var]]), n = 300))
    ix <- findInterval(log10(data[[x_var]]), dens$x)
    iy <- findInterval(log10(data[[y_var]]), dens$y)
  } else {
    # Linear-scale density calculation
    dens <- with(data, kde2d(data[[x_var]], data[[y_var]], n = 300))
    ix <- findInterval(data[[x_var]], dens$x)
    iy <- findInterval(data[[y_var]], dens$y)
  }

  data$density <- dens$z[cbind(ix, iy)]

  # Base plot
  p <- ggplot(data, aes_string(x = x_var, y = y_var, color = "density")) +
    geom_point(size = 0.5, alpha = 0.8) +
    scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
    labs(x = x_label, y = y_label) +
    theme_minimal()

  # Add model lines
  p <- p +
    geom_line(data = df_pl, aes(x = x, y = y), color = "green", size = 1) +
    geom_line(data = df_gmm, aes(x = x, y = y), color = "blue", size = 1) +
    geom_line(data = df_wb, aes(x = x, y = y), color = "red", size = 1)

  # Handle log scale if requested
  if (log_scale) {
    p <- p + scale_x_log10() + scale_y_log10()
  }

  return(p)
}
```

## Plot
```{r}
# Plot for H-DBH using Power-Law model type
combined_plot_h <- plot_model_comparison(
  data = data2,
  df_pl = df_pl_h,
  df_gmm = df_gmm_h,
  df_wb = df_wb_h,
  x_var = "dbh",
  y_var = "h",
  x_label = "DBH (cm)",
  y_label = "Height (m)",
  log_scale = TRUE    # or FALSE, depending on your preference
)

# Plot for CR-DBH using Power-Law model type
combined_plot_cr <- plot_model_comparison(
  data = data2,
  df_pl = df_pl_cr,
  df_gmm = df_gmm_cr,
  df_wb = df_wb_cr,
  x_var = "dbh",
  y_var = "cr",
  x_label = "DBH (cm)",
  y_label = "Crown Radius (m)",
  log_scale = TRUE    # or FALSE
)

```

# Combining plots

```{r}
# Increase the size of axis titles
combined_plot_h <- combined_plot_h + 
  theme(
    plot.title = element_text(hjust = 0, face = "bold"),  
    axis.title.x = element_blank(),                       # Remove x-axis title from the top plot
    axis.text.x = element_blank(),                        # Remove x-axis ticks from the top plot
    axis.ticks.x = element_blank(),                       # Remove x-axis ticks from the top plot
    axis.title.y = element_text(size = 16),               # Increase y-axis title size
    plot.margin = margin(5, 5, 5, 5),                     # Adjust margins if needed
    legend.position = "none"                              # Remove the legend from the top plot
  ) + ggtitle("(a)")

combined_plot_cr <- combined_plot_cr + 
  theme(
    plot.title = element_text(hjust = 0, face = "bold"),  # Left-align the title and make it bold
    axis.title.x = element_text(size = 16),               # Increase x-axis title size
    axis.title.y = element_text(size = 16),               # Increase y-axis title size
    plot.margin = margin(5, 5, 5, 5)                      # Adjust margins if needed
  ) + ggtitle("(b)")

combined_plots <- combined_plot_h / combined_plot_cr + 
  plot_layout(ncol = 1, heights = c(1, 1))

ggsave("combined_plots_with_larger_labels.png", plot = combined_plots, width = 10, height = 12, dpi = 300)
```

```{r}
# Plot for H-DBH using Power-Law model type
combined_plot_h_non <- plot_model_comparison(
  data = data2,
  df_pl = df_pl_h,
  df_gmm = df_gmm_h,
  df_wb = df_wb_h,
  x_var = "dbh",
  y_var = "h",
  x_label = "DBH (cm)",
  y_label = "H (m)",
  log_scale = FALSE    
)

# Plot for CR-DBH using Power-Law model type
combined_plot_cr_non <- plot_model_comparison(
  data = data2,
  df_pl = df_pl_cr,
  df_gmm = df_gmm_cr,
  df_wb = df_wb_cr,
  x_var = "dbh",
  y_var = "cr",
  x_label = "DBH (cm)",
  y_label = "CR (m)",
  log_scale = FALSE
)

print(combined_plot_h_non)
print(combined_plot_cr_non)
```