### 2025-03-26 docs/cov_beta.qmd

---
title: "Checking the covariance between beta parameter of gmm_nou_h"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load required packages 
```{r}
library(cmdstanr)
library(posterior)
library(targets)
```

# Data preparation
```{r}
gmm_nou_h <- tar_read(fit_nlr7_mcmc_gmm_nou_h)
```

# Extract correlations among each parameter 
```{r}
# Convert posterior samples to an array
posterior_samples <- as_draws_array(gmm_nou_h$draws())

# Number of posterior samples
num_samples <- dim(posterior_samples)[1]
print(num_samples)

# Number of predictors (K)
K <- 3

# Extract L_Sigma_beta samples
# Adjust the regex pattern if the variable names differ
L_Sigma_beta_samples <- posterior_samples[, , grep("^L_Sigma_beta", dimnames(posterior_samples)$variable)]

# Initialize array to store covariance matrices
cov_matrices <- array(NA, dim = c(K, K, num_samples))

# Calculate covariance matrix for each sample
for (i in 1:num_samples) {
  L_Sigma_beta <- matrix(L_Sigma_beta_samples[i, , ], nrow = K, byrow = TRUE)
  cov_matrices[, , i] <- L_Sigma_beta %*% t(L_Sigma_beta)
  }

```

```{r}
# Calculate the mean covariance matrix
mean_cov_matrix <- apply(cov_matrices, c(1, 2), mean)

# Print the mean covariance matrix
print(mean_cov_matrix)

# Calculate the mean correlation matrix
mean_cor_matrix <- cov2cor(mean_cov_matrix)

# Print the mean correlation matrix
print(mean_cor_matrix)
```

# Interpretation:

- Diagonal Element (Variance):
Variance of beta1: 853.0958
Variance of beta 2: 112.9586
Variance of beta 3: 478.6792
- Off-Diagonal Elements (Covariance):
Covariance between beta1 and beta2: 306.1926
Covariance between beta1 and beta3: 636.5709
Covariance between beta2 and beta3: 230.3332
- The Mean Correlation Matrix
Correlation Coefficients:
Correlation between beta1 and beta2: 0.9863616
Correlation between beta1 and beta3: 0.9961519
Correlation between beta2 and beta3: 0.9905461

# Summary
- All correlations are very high (close to 1), indicating strong positive relationship between beta parameters.
- Betas move together strongly in the same direction (in gmm_nou_h model)

# Checking 95% CI of the correlations coefficients
```{r}
# Calculate correlation matrices for each sample
cor_matrices <- array(NA, dim = c(K, K, num_samples))
for (i in 1:num_samples) {
  L_Sigma_beta <- matrix(L_Sigma_beta_samples[i, , ], nrow = K, byrow = TRUE)
  cov_matrix <- L_Sigma_beta %*% t(L_Sigma_beta)
  cor_matrices[, , i] <- cov2cor(cov_matrix)
}

# Initialize matrices to store the lower and upper bounds of the 95% CI
lower_ci <- matrix(NA, nrow = K, ncol = K)
upper_ci <- matrix(NA, nrow = K, ncol = K)

# Calculate the 95% CI for each correlation coefficient
for (row in 1:K) {
  for (col in 1:K) {
    cor_values <- cor_matrices[row, col, ]
    lower_ci[row, col] <- quantile(cor_values, probs = 0.025)
    upper_ci[row, col] <- quantile(cor_values, probs = 0.975)
  }
}

# Print the 95% CI for each correlation coefficient
print(lower_ci)
print(upper_ci)

```

```{r}
# Assuming you have the posterior samples already extracted
posterior_samples <- as_draws_array(gmm_h$draws())

# Number of posterior samples
num_samples <- dim(posterior_samples)[1]

# Define the number of tree-level predictors (K)
K <- 2

# Extract L_Sigma_beta samples
L_Sigma_beta_samples <- posterior_samples[, , grep("^L_Sigma_beta", dimnames(posterior_samples)$variable)]

# Initialize array to store covariance matrices
cov_matrices <- array(NA, dim = c(K, K, num_samples))

# Calculate covariance matrix for each sample
for (i in 1:num_samples) {
  L_Sigma_beta <- matrix(L_Sigma_beta_samples[i, , ], nrow = K, byrow = TRUE)
  cov_matrices[,,i] <- L_Sigma_beta %*% t(L_Sigma_beta)
}

# Calculate the mean covariance matrix
mean_cov_matrix <- apply(cov_matrices, c(1, 2), mean)

# Print the mean covariance matrix
print(mean_cov_matrix)

# Calculate the correlations
mean_cor_matrix <- cov2cor(mean_cov_matrix)

# Print the mean correlation matrix
print(mean_cor_matrix)

# Calculate 95% credible intervals for the correlations
cor_samples <- array(NA, dim = c(K, K, num_samples))
for (i in 1:num_samples) {
  cor_samples[,,i] <- cov2cor(cov_matrices[,,i])
}

# Calculate the 95% credible intervals
lower_ci <- apply(cor_samples, c(1, 2), function(x) quantile(x, 0.025))
upper_ci <- apply(cor_samples, c(1, 2), function(x) quantile(x, 0.975))

# Print the 95% credible intervals
print(lower_ci)
print(upper_ci)
```