### 2025-03-26 docs/cr_3functions.qmd

---
title: "CR models - plotting 3 community-lines (pl, wd, gMM in a single panel)"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(cmdstanr)
library(posterior)
library(loo)
library(dplyr)
library(ggplot2)
library(stringr)
library(reshape2)
library(scales)

# Define and run the targets pipeline
# tar_make()
```

# Unscale power-law's parameters

```{r}
tallo_wd_df_200 <- tar_read(tallo_wd_df_200)
tallo_wd_df_200

data2 <- tallo_wd_df_200 |>
    filter(!is.na(dbh)) |>
    filter(!is.na(cr)) |>
    filter(!is.na(h)) |>
    filter(!is.na(wd_s))

data2 <- data2 |>
    mutate(log_dbh = log(dbh),
           log_h = log(h),
           log_cr = log(cr)) |>
    mutate(log_dbh_s = scale(log_dbh) |> as.numeric(),
           log_h_s = scale(log_h) |> as.numeric(),
           log_cr_s = scale(log_cr) |> as.numeric()) |>
    mutate(dbh_s = scale(dbh) |> as.numeric())

data2

# Extract scaling parameters for unscaling
scaling_params <- data2 |>
    summarise(mean_log_dbh = mean(log_dbh), sd_log_dbh = sd(log_dbh),
              mean_log_cr = mean(log_cr), sd_log_cr = sd(log_cr))

scaling_params
```

# Plotting 3 community-lines (pl, wd, gMM in a single panel)

```{r}
# Read in the data (pl)
stan_data_lr_cr <- tar_read(stan_data_lr_cr)
str(stan_data_lr_cr)

# Extract relevant data from stan_data_lr_cr (log_y, x, jj)
log_y_lr <- stan_data_lr_cr$log_y
log_x_lr <- stan_data_lr_cr$log_x[, 2]
sp_lr <- as.factor(stan_data_lr_cr$jj)

unscaled_dbh_pl <- exp((log_x_lr * scaling_params$sd_log_dbh) + scaling_params$mean_log_dbh)
unscaled_cr_pl <- exp((log_y_lr * scaling_params$sd_log_cr) + scaling_params$mean_log_cr)

# Read in the data (gMM, wb)
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

# Extract relevant data from stan_data_nlr_cr (log_y, x, jj)
log_y_nlr <- stan_data_nlr_cr$log_y
x_nlr <- stan_data_nlr_cr$x
sp_nlr <- as.factor(stan_data_nlr_cr$jj)

# Read in the model summaries for generalized Michaelis-Menten (gMM), Weibull (wb), and power-law (pl)
gmm_summary <- tar_read(fit_nlr_summary_gmm1_cr)
wb_summary <- tar_read(fit_nlr_summary_weibull1_cr)
pl_summary <- tar_read(fit_lr2_summary_power_law1_cr)

gmm_summary
wb_summary
pl_summary

# Extract beta parameters for power-law model
beta_pl <- pl_summary |>
  filter(grepl("beta\\[", variable))

beta_pl

# Extract log_a (beta[1, ]) and b (beta[2, ])
log_a_pl <- beta_pl |>
  filter(str_detect(variable, "beta\\[1,\\d+\\]")) |>
  pull(mean)

b_pl <- beta_pl |>
  filter(str_detect(variable, "beta\\[2,\\d+\\]")) |>
  pull(mean)

# Print extracted values to verify
head(log_a_pl)
head(b_pl)

# Extract beta parameters for gMM model
beta_gmm <- gmm_summary |>
  filter(grepl("beta\\[", variable))

beta_gmm

# Extract log_a, b, k from beta_gmm
log_a_gmm <- beta_gmm |>
  filter(str_detect(variable, "beta\\[\\d+,1\\]")) |>
  pull(mean)

b_gmm <- beta_gmm |>
  filter(str_detect(variable, "beta\\[\\d+,2\\]")) |>
  pull(mean)

k_gmm <- beta_gmm |>
  filter(str_detect(variable, "beta\\[\\d+,3\\]")) |>
  pull(mean)

head(log_a_gmm)
head(b_gmm)
head(k_gmm)

# Extract beta parameters for Weibull model
beta_wb <- wb_summary |>
  filter(grepl("beta\\[", variable))

beta_wb

# Extract log_a, b, k from beta_wb
log_a_wb <- beta_wb |>
  filter(str_detect(variable, "beta\\[\\d+,1\\]")) |>
  pull(mean)

b_wb <- beta_wb |>
  filter(str_detect(variable, "beta\\[\\d+,2\\]")) |>
  pull(mean)

k_wb <- beta_wb |>
  filter(str_detect(variable, "beta\\[\\d+,3\\]")) |>
  pull(mean)

head(log_a_wb)
head(b_wb)
head(k_wb)

```

# Extract community-level parameters of each function
(I will change by median later)
```{r}
##pl
pl_summary
gamma_pl <- pl_summary |> filter(grepl("gamma", variable))
gamma_pl

community_log_a_hat_pl <- gamma_pl |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat_pl <- gamma_pl |> filter(variable == "gamma[2,1]") |> pull(mean)

community_log_a_pl <- exp(community_log_a_hat_pl)
community_b_pl <- exp(community_b_hat_pl)

community_log_a_pl
community_b_pl

## gMM
gmm_summary
gamma_gmm <- gmm_summary |> filter(grepl("gamma", variable))
gamma_gmm
community_log_a_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,2]") |> pull(mean)
community_k_hat_gmm <- gamma_gmm |> filter(variable == "gamma[1,3]") |> pull(mean)

community_log_a_gmm <- exp(community_log_a_hat_gmm)
community_b_gmm <- exp(community_b_hat_gmm) 
community_k_gmm <- exp(community_k_hat_gmm)

community_log_a_gmm
community_b_gmm
community_k_gmm

# Weibull
wb_summary
gamma_wb <- wb_summary |> filter(grepl("gamma", variable))
gamma_wb
community_log_a_hat_wb <- gamma_wb |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_hat_wb <- gamma_wb |> filter(variable == "gamma[1,2]") |> pull(mean)
community_k_hat_wb <- gamma_wb |> filter(variable == "gamma[1,3]") |> pull(mean)

community_log_a_wb <- exp(community_log_a_hat_wb)
community_b_wb <- exp(community_b_hat_wb) 
community_k_wb <- exp(community_k_hat_wb)

community_log_a_wb
community_b_wb
community_b_wb

# Computing Fitted Values for Community-level Parameters
#pl
community_fitted_log_y_cr_pl <- community_log_a_pl + community_b_pl * log_x_lr
community_fitted_cr_pl <- exp(community_fitted_log_y_cr_pl)

head(community_fitted_cr_pl)

#gMM

community_fitted_log_y_cr_gmm <- community_log_a_gmm + community_b_gmm * log(x_nlr) - log(community_k_gmm + x_nlr^community_b_gmm)
community_fitted_cr_gmm <- exp(community_fitted_log_y_cr_gmm)

head(community_fitted_cr_gmm)

# wb

community_fitted_log_y_cr_wb <- community_log_a_wb + log(1 - exp(-community_b_wb * (x_nlr ^ community_k_wb)))
community_fitted_cr_wb <- exp(community_fitted_log_y_cr_wb)

head(community_fitted_cr_wb)

# Create a data frame for plotting
data_cr_pl_unscaled <- data.frame(CR = unscaled_cr_pl, DBH = unscaled_dbh_pl, sp = as.factor(stan_data_lr_cr$jj), community_fitted_CR_pl = community_fitted_cr_pl)

head(data_cr_pl_unscaled)

data_cr_gb <- data.frame(CR = exp(log_y_nlr), DBH = x_nlr, sp = sp_nlr)
head(data_cr_gb)

data_cr_gb$community_fitted_CR_gmm <- community_fitted_cr_gmm
data_cr_gb$community_fitted_CR_wb <- community_fitted_cr_wb

head(data_cr_pl_unscaled)
head(data_cr_gb)

# Combine the data frames
data_combined <- merge(data_cr_pl_unscaled, data_cr_gb, by = "sp", all = TRUE)
head(data_combined)

# Separate the community-fitted values for plotting
data_community_pl <- data_combined |> select(DBH.x, community_fitted_CR_pl) |> distinct()
data_community_gmm <- data_combined |> select(DBH.y, community_fitted_CR_gmm) |> distinct()
data_community_wb <- data_combined |> select(DBH.y, community_fitted_CR_wb) |> distinct()

head(data_community_pl)
head(data_community_gmm)
head(data_community_wb)
```

# Plot the results
```{r}
combined_plot <- ggplot(data_combined) +
  geom_point(aes(x = DBH.x, y = CR.x), alpha = 0.5, color = "grey") +
  geom_point(aes(x = DBH.y, y = CR.y), alpha = 0.5, color = "grey") +
  geom_line(data = data_community_pl, aes(x = DBH.x, y = community_fitted_CR_pl, color = "Power-law"), size = 1) +
  geom_line(data = data_community_gmm, aes(x = DBH.y, y = community_fitted_CR_gmm, color = "gMM"), size = 1) +
  geom_line(data = data_community_wb, aes(x = DBH.y, y = community_fitted_CR_wb, color = "Weibull"), size = 1) +
  labs(title = "Community-level Model Fits for Power-law, gMM, and Weibull Models",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal() +
  scale_color_manual(name = "Model", values = c("Power-law" = "blue", "gMM" = "red", "Weibull" = "green"))

# Save the plot
ggsave(filename = "figs/cr/DBH_CR_community_combined.png", plot = combined_plot, width = 12, height = 8, dpi = 300, bg = "white")

print(combined_plot)
```