### 2025-03-26 docs/cr_conv.qmd

---
title: "CR-DBH allometry convergence's checking"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(dplyr)
```

# Power-law

## Without wood density

```{r}

fit_lr2_summary_power_law1_nou_cr <- tar_read(fit_lr2_summary_power_law1_nou_cr)
fit_lr2_summary_power_law1_nou_cr
beta_pl_nou_cr <- fit_lr2_summary_power_law1_nou_cr |> filter(grepl("beta\\[", variable))
summary(beta_pl_nou_cr)
```

- No problem

# With wood density
```{r}
fit_lr2_summary_power_law1_cr <- tar_read(fit_lr2_summary_power_law1_cr)
fit_lr2_summary_power_law1_cr
beta_pl_cr <- fit_lr2_summary_power_law1_cr |> filter(grepl("beta\\[", variable))
beta_pl_cr
summary(beta_pl_cr)
```

- No problem

# gMM

## Without wood density

```{r}
fit_nlr_summary_gmm1_nou_cr <- tar_read(fit_nlr_summary_gmm1_nou_cr)
fit_nlr_summary_gmm1_nou_cr
beta_gmm_nou_cr <- fit_nlr_summary_gmm1_nou_cr |> filter(grepl("beta\\[", variable))
summary(beta_gmm_nou_cr)

low_ess_gmm_nou_cr_beta <- beta_gmm_nou_cr |>
  filter(ess_bulk < 300)

low_ess_gmm_nou_cr_beta
summary(low_ess_gmm_nou_cr_beta)
```

- 5 species has convergent issue but ESS >= 250 so I think it OK.

## With wood density
```{r}
fit_nlr_summary_gmm1_cr <- tar_read(fit_nlr_summary_gmm1_cr)
fit_nlr_summary_gmm1_cr
beta_gmm_cr <- fit_nlr_summary_gmm1_cr |> filter(grepl("beta\\[", variable))
summary(beta_gmm_cr)

low_ess_gmm_cr_beta <- beta_gmm_cr |>
  filter(ess_bulk < 300)

print(low_ess_gmm_cr_beta, n = Inf)

low_ess_gmm_cr_beta <- low_ess_gmm_cr_beta |>
  mutate(sp_id = sub("beta\\[([0-9]+),[0-9]+\\]", "\\1", variable),
         dimension = sub("beta\\[[0-9]+,([0-9]+)\\]", "\\1", variable))

species_counts <- low_ess_gmm_cr_beta |>
  group_by(sp_id) |>
  summarise(dimension_count = n_distinct(dimension))

species_counts

num_species <- low_ess_gmm_cr_beta |>
  distinct(sp_id) |>
  nrow()
```

- 22 species has convergent issue but ESS >= 250

# Weibull

## Without wood density

```{r}
fit_nlr_summary_weibull1_nou_cr <- tar_read(fit_nlr_summary_weibull1_nou_cr)
fit_nlr_summary_weibull1_nou_cr
beta_weibull_nou_cr <- fit_nlr_summary_weibull1_nou_cr |> filter(grepl("beta\\[", variable))
summary(beta_weibull_nou_cr)

low_ess_weibull_nou_cr_beta <- beta_weibull_nou_cr |>
  filter(ess_bulk < 300)

low_ess_weibull_nou_cr_beta
summary(low_ess_weibull_nou_cr_beta)

low_ess_weibull_nou_cr_beta <- low_ess_weibull_nou_cr_beta |>
  mutate(sp_id = sub("beta\\[([0-9]+),[0-9]+\\]", "\\1", variable))

num_species_low_ess <- low_ess_weibull_nou_cr_beta |>
  distinct(sp_id) |>
  nrow()

print(num_species_low_ess)
```

# With wood density

```{r}
fit_nlr_summary_weibull1_cr <- tar_read(fit_nlr_summary_weibull1_cr)
fit_nlr_summary_weibull1_cr
beta_weibull_cr <- fit_nlr_summary_weibull1_cr |> filter(grepl("beta\\[", variable))
summary(beta_weibull_cr)

low_ess_weibull_cr_beta <- beta_weibull_cr |>
  filter(ess_bulk < 300)

low_ess_weibull_cr_beta

summary(low_ess_weibull_cr_beta)
print(low_ess_weibull_cr_beta, n = Inf)

low_ess_weibull_cr_beta <- low_ess_weibull_cr_beta |>
  mutate(sp_id = sub("beta\\[([0-9]+),[0-9]+\\]", "\\1", variable))

num_species_low_ess <- low_ess_weibull_cr_beta |>
  distinct(sp_id) |>
  nrow()

print(num_species_low_ess)
```

- 400 sp have convergent issue.