### 2025-03-26 docs/cr_new.qmd

---
title: "Comparing CR models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r, include=FALSE}
library(targets)
library(tidyverse)
library(scales)
library(MASS)
```

# Read in data

```{r}
options(width = 200)
tallo_wd_df_cr <- tar_read(tallo_wd_df_cr)
tallo_wd_df_cr

ang_cr_df <- tallo_wd_df_cr |> filter(division == "Angiosperm")
ang_cr_df

gym_cr_df <- tallo_wd_df_cr |> filter(division == "Gymnosperm")
gym_cr_df

```

# Power-law

## Ang
```{r}
fit_lr_nou_data_ang_cr <- tar_read(fit_lr_nou_data_ang_cr)
str(fit_lr_nou_data_ang_cr)

pl_ang_cr <- tar_read(fit_lr_nou_summary_pl_ang_cr)
pl_ang_cr

gamma_ang_pl <- pl_ang_cr |> filter(grepl("gamma", variable))

community_log_a_ang_pl <- gamma_ang_pl |> filter(variable == "gamma[1]") |> pull(q50)
community_b_ang_pl <- gamma_ang_pl |> filter(variable == "gamma[2]") |> pull(q50)

log_x_range_ang <- fit_lr_nou_data_ang_cr$log_x[,2] |> range()
log_x_seq_ang <- seq(log_x_range_ang[1], log_x_range_ang[2], length.out = 100)
fitted_log_y_seq_ang <- community_log_a_ang_pl + community_b_ang_pl * log_x_seq_ang
y_pred_ang <- exp(fitted_log_y_seq_ang)
df_ang_pl_cr <- tibble(y = y_pred_ang, x = exp(log_x_seq_ang))

ggplot() +
  geom_bin2d(data = ang_cr_df, aes(x = dbh, y = cr), bins = 100)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_ang_pl_cr, aes(x = x, y = y)) +
  labs(x = "DBH (cm)", y = "CR (m)")

```

## Gym

```{r}
fit_lr_nou_data_gym_cr <- tar_read(fit_lr_nou_data_gym_cr)
str(fit_lr_nou_data_gym_cr)

# pl_gym_cr <- tar_read(fit_lr_nou_summary_pl_biome_gym_cr)
# pl_gym_cr

pl_gym_cr <- tar_read(fit_lr_nou_summary_pl_gym_cr)
pl_gym_cr

gamma_gym_pl <- pl_gym_cr |> filter(grepl("gamma", variable))

community_log_a_gym_pl <- gamma_gym_pl |> filter(variable == "gamma[1]") |> pull(q50)
community_b_gym_pl <- gamma_gym_pl |> filter(variable == "gamma[2]") |> pull(q50)

log_x_range_gym_cr <- fit_lr_nou_data_gym_cr$log_x[,2] |> range()
log_x_seq_gym_cr <- seq(log_x_range_gym_cr[1], log_x_range_gym_cr[2], length.out = 100)

fitted_log_y_seq_gym <- community_log_a_gym_pl + community_b_gym_pl * log_x_seq_gym_cr
y_pred_gym_cr <- exp(fitted_log_y_seq_gym)

df_gym_pl_cr <- tibble(y = y_pred_gym_cr, x = exp(log_x_seq_gym_cr))
df_gym_pl_cr

ggplot() +
  geom_bin2d(data = gym_cr_df, aes(x = dbh, y = cr), bins = 100) +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gym_pl_cr, aes(x = x, y = y))

```

# gMM

## Ang
```{r}
fit_nlr_nou_data_ang_cr <- tar_read(fit_nlr_nou_data_ang_cr)
str(fit_nlr_nou_data_ang_cr)

x_nlr_range_ang <- range(fit_nlr_nou_data_ang_cr$x)
x_nlr_seq_ang <- seq(x_nlr_range_ang[1], x_nlr_range_ang[2], length.out = 100)

gmm_ang_cr <- tar_read(fit_nlr_nou_summary_gmm_ang_cr)
gamma_ang_gmm <- gmm_ang_cr |> filter(grepl("gamma(?!_hat)", variable, perl = TRUE))
community_log_a_ang_gmm <- gamma_ang_gmm |> filter(variable == "gamma[1]") |> pull(q50)
community_b_ang_gmm <- gamma_ang_gmm |> filter(variable == "gamma[2]") |> pull(q50)
community_k_ang_gmm <- gamma_ang_gmm |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_cr_gmm_seq <- community_log_a_ang_gmm + community_b_ang_gmm * log(x_nlr_seq_ang) - log(community_k_ang_gmm + x_nlr_seq_ang^community_b_ang_gmm)
y_ang_gmm_pred <- exp(community_fitted_log_y_cr_gmm_seq)
df_ang_gmm_cr <- tibble(y = y_ang_gmm_pred, x = x_nlr_seq_ang)

ggplot() +
  geom_bin2d(data = ang_cr_df, aes(x = dbh, y = cr), bins = 100) +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_ang_gmm_cr, aes(x = x, y = y)) +
  labs(x = "DBH (cm)", y = "CR (m)")

```

## Gym

```{r}
fit_nlr_nou_data_gym_cr <- tar_read(fit_nlr_nou_data_gym_cr)
str(fit_nlr_nou_data_gym_cr)

x_nlr_range_gym_cr <- range(fit_nlr_nou_data_gym_cr$x)
x_nlr_seq_gym_cr  <- seq(x_nlr_range_gym_cr[1], x_nlr_range_gym_cr[2], length.out = 100)

gmm_gym_cr <- tar_read(fit_nlr_nou_summary_gmm_gym_cr)
gmm_gym_cr

gamma_gym_gmm <- gmm_gym_cr |> filter(grepl("gamma(?!_hat)", variable, perl = TRUE))
gamma_gym_gmm

community_log_a_gym_gmm <- gamma_gym_gmm |> filter(variable == "gamma[1]") |> pull(q50)
community_b_gym_gmm <- gamma_gym_gmm |> filter(variable == "gamma[2]") |> pull(q50)
community_k_gym_gmm <- gamma_gym_gmm |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_cr_gym_gmm_seq <- community_log_a_gym_gmm + community_b_gym_gmm * log(x_nlr_seq_gym_cr) - log(community_k_gym_gmm + x_nlr_seq_gym_cr^community_b_gym_gmm)

y_gym_gmm_pred_cr <- exp(community_fitted_log_y_cr_gym_gmm_seq)

df_gym_gmm_cr <- tibble(y = y_gym_gmm_pred_cr, x = x_nlr_seq_gym_cr)
df_gym_gmm_cr

ggplot() +
  geom_bin2d(data = gym_cr_df, aes(x = dbh, y = cr), bins = 100)  +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gym_gmm_cr, aes(x = x, y = y))
```

# Weibull

## Ang
```{r}
fit_nlr_nou_data_ang_cr <- tar_read(fit_nlr_nou_data_ang_cr)
x_nlr_range_ang_cr <- range(fit_nlr_nou_data_ang_cr$x)
x_nlr_seq_ang_cr <- seq(x_nlr_range_ang_cr[1], x_nlr_range_ang_cr[2], length.out = 100)

weibull_ang_cr <- tar_read(fit_nlr_nou_summary_weibull_ang_cr)
weibull_ang_cr

gamma_ang_wb <- weibull_ang_cr |> filter(grepl("gamma", variable))
community_log_a_ang_wb <- gamma_ang_wb |> filter(variable == "gamma[1]") |> pull(q50)
community_b_ang_wb <- gamma_ang_wb |> filter(variable == "gamma[2]") |> pull(q50)
community_k_ang_wb <- gamma_ang_wb |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_cr_wb_ang_seq <- community_log_a_ang_wb + log(1 - exp(-community_b_ang_wb * (x_nlr_seq_ang ^ community_k_ang_wb)))

y_ang_wb_pred_cr <- exp(community_fitted_log_y_cr_wb_ang_seq)

df_ang_wb_cr <- tibble(y = y_ang_wb_pred_cr, x = x_nlr_seq_ang_cr)

ggplot() +
  geom_bin2d(data = ang_cr_df, aes(x = dbh, y = cr), bins = 100) +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_ang_wb_cr, aes(x = x, y = y)) +
  labs(x = "DBH (cm)", y = "CR (m)")
```

## Gym

```{r}
fit_nlr_nou_data_gym_cr <- tar_read(fit_nlr_nou_data_gym_cr)
str(fit_nlr_nou_data_gym_cr)

x_nlr_range_gym_cr <- range(fit_nlr_nou_data_gym_cr$x)
x_nlr_seq_gym_cr <- seq(x_nlr_range_gym_cr[1], x_nlr_range_gym_cr[2], length.out = 100)

weibull_gym_cr <- tar_read(fit_nlr_nou_summary_weibull_gym_cr)
gamma_gym_wb <- weibull_gym_cr |> filter(grepl("gamma", variable))

community_log_a_gym_wb <- gamma_gym_wb |> filter(variable == "gamma[1]") |> pull(q50)
community_b_gym_wb <- gamma_gym_wb |> filter(variable == "gamma[2]") |> pull(q50)
community_k_gym_wb <- gamma_gym_wb |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_cr_wb_gym_seq <- community_log_a_gym_wb + log(1 - exp(-community_b_gym_wb * (x_nlr_seq_gym_cr ^ community_k_gym_wb)))

y_gym_wb_pred_cr <- exp(community_fitted_log_y_cr_wb_gym_seq)

df_gym_wb_cr <- tibble(y = y_gym_wb_pred_cr, x = x_nlr_seq_gym_cr)

ggplot() +
  geom_bin2d(data = gym_cr_df, aes(x = dbh, y = cr), bins = 100) +
  scale_fill_viridis_c(option = "D") +
  geom_line(data = df_gym_wb_cr, aes(x = x, y = y))
```

# Plotting with density-based coloring log-scale

## Ang

```{r}
dens <- with(ang_cr_df, kde2d(log10(dbh), log10(cr), n = 300))
ix <- findInterval(log10(ang_cr_df$dbh), dens$x)
iy <- findInterval(log10(ang_cr_df$cr), dens$y)
ang_cr_df$density <- dens$z[cbind(ix, iy)]

p_ang_cr_den <- ggplot(data = ang_cr_df, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_gym_pl_cr, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_ang_gmm_cr, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_ang_wb_cr, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(x = "DBH_ang (cm)", y = "CR_ang (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

p_ang_cr_den

ggsave("figs/p_cr_ang_den.png", plot = p_ang_cr_den, width = 8, height = 6, dpi = 300, bg = "white")
```

## Gym

```{r}
dens <- with(gym_cr_df, kde2d(log10(dbh), log10(cr), n = 300))
ix <- findInterval(log10(gym_cr_df$dbh), dens$x)
iy <- findInterval(log10(gym_cr_df$cr), dens$y)
gym_cr_df$density <- dens$z[cbind(ix, iy)]

# Create the plot with log scale
p_gym_cr_den_log <- ggplot(data = gym_cr_df, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_gym_pl_cr, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_gym_gmm_cr, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_gym_wb_cr, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(x = "DBH_gym (cm)", y = "CR_gym (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

p_gym_cr_den_log

ggsave("figs/p_gym_cr_log_den.png", plot = p_gym_cr_den_log, width = 8, height = 6, dpi = 300, bg = "white")
```

# Plotting with density-based coloring non-log scale

## Ang

```{r}
dens <- with(ang_cr_df, kde2d(dbh, cr, n = 300))
ix <- findInterval(ang_cr_df$dbh, dens$x)
iy <- findInterval(ang_cr_df$cr, dens$y)
ang_cr_df$density <- dens$z[cbind(ix, iy)]

p_ang_cr_den_non <- ggplot(data = ang_cr_df, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_ang_pl_cr, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_ang_gmm_cr, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_ang_wb_cr, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(x = "DBH_ang (cm)", y = "CR_ang (m)") +
  theme_minimal()

p_ang_cr_den_non

ggsave("figs/p_ang_cr_non_den_non.png", plot = p_ang_cr_den_non, width = 8, height = 6, dpi = 300, bg = "white")
```

## Gym

```{r}
dens <- with(gym_cr_df, kde2d(dbh, cr, n = 300))
ix <- findInterval(gym_cr_df$dbh, dens$x)
iy <- findInterval(gym_cr_df$cr, dens$y)
gym_cr_df$density <- dens$z[cbind(ix, iy)]

p_gym_cr_den_non <- ggplot(data = gym_cr_df, aes(x = dbh, y = cr, color = density)) +
  geom_point(size = 0.5, alpha = 0.8) +
  scale_color_gradientn(colors = c("black", "purple", "orange", "yellow"), name = "Density") +
  geom_line(data = df_gym_pl_cr, aes(x = x, y = y), color = "#FF5733", size = 0.8) +
  geom_line(data = df_gym_gmm_cr, aes(x = x, y = y), color = "#2ECC71", size = 0.8) +
  geom_line(data = df_gym_wb_cr, aes(x = x, y = y), color = "#0f92e9", size = 0.8) +
  labs(x = "DBH_gym (cm)", y = "CR_gym (m)") +
  theme_minimal()

p_gym_cr_den_non

ggsave("figs/p_gym_cr_non_den.png", plot = p_gym_cr_den_non, width = 8, height = 6, dpi = 300, bg = "white")
```