### 2025-03-26 docs/dbh.qmd

---
title: "Plotting DBH models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(cmdstanr)
library(posterior)
library(loo)
library(dplyr)
library(ggplot2)
library(stringr)
library(reshape2)
library(scales)
# Define and run the targets pipeline
# tar_make()
```

# Plot DBH = aCR^b * H^c model (1, 5) (no WD)

```{r}
# Extract relevant data for model 1 (loo_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)

log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]
log_x_h <- stan_data_lr_dbh$log_x[, 3]
sp <- as.factor(stan_data_lr_dbh$jj)

str(log_x_cr)
str(log_x_h)

# Read in the model summary for model 
dbh_nou_summary <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
dbh_nou_summary

# Filter rows where the variable is "beta"
beta <- dbh_nou_summary |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(mean)

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

fitted_dbh <- exp(fitted_log_y_dbh)

gamma <- dbh_nou_summary |>
  filter(grepl("gamma", variable))
gamma

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a <- gamma |> filter(variable == "gamma[1]") |> pull(mean)
community_b_cr <- gamma |> filter(variable == "gamma[2]") |> pull(mean)
community_b_h <- gamma |> filter(variable == "gamma[3]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CR = exp(log_x_cr), H = exp(log_x_h), sp = sp)
data_dbh$community_fitted_DBH <- community_fitted_DBH
head(data_dbh)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Compute CR^b and H^c for each observation
data_dbh$CR_b <- data_dbh$CR^community_b_cr
data_dbh$H_c <- data_dbh$H^community_b_h
data_dbh$CR_H <- data_dbh$CR_b*data_dbh$H_c

head(data_dbh)

data_long <- reshape2::melt(data_dbh, id.vars = c("DBH", "CR_H", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

dbh_plot <- ggplot(data_long, aes(x = CR_H, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CR_H, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*CR^b * H^c",
       x = "CR_H",
       y = "DBH") +
  scale_x_log10(labels = label_number()) +
  scale_y_log10(labels = label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/H_CR_DBH1.png", plot = dbh_plot, width = 12, height = 8, dpi = 300, bg = "white")

dbh_plot
```

- Note: Plot looks weird and I am struggling with this TT_TT

```{r}
# Extract relevant data for model 1 (loo_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)

log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]
log_x_h <- stan_data_lr_dbh$log_x[, 3]
sp <- as.factor(stan_data_lr_dbh$jj)

str(log_x_cr)
str(log_x_h)

# Read in the model summary for model 
dbh_nou_summary <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
dbh_nou_summary

# Filter rows where the variable is "beta"
beta <- dbh_nou_summary |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(mean)

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

gamma <- dbh_nou_summary |>
  filter(grepl("gamma", variable))
gamma

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a <- gamma |> filter(variable == "gamma[1]") |> pull(mean)
community_b_cr <- gamma |> filter(variable == "gamma[2]") |> pull(mean)
community_b_h <- gamma |> filter(variable == "gamma[3]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h

# Create a data frame for plotting with log-transformed values
data_dbh <- data.frame(log_DBH = log_y, log_CR = log_x_cr, log_H = log_x_h, sp = sp)
data_dbh$community_fitted_log_DBH <- community_fitted_log_y_dbh

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_log_DBH_sp_", i)]] <- fitted_log_y_dbh[, i]
}

# Compute CR^b and H^c for each observation in the log scale
data_dbh$log_CR_b <- community_b_cr * data_dbh$log_CR
data_dbh$log_H_c <- community_b_h * data_dbh$log_H
data_dbh$log_CR_H <- data_dbh$log_CR_b + data_dbh$log_H_c

# Reshape the data for plotting
data_long <- reshape2::melt(data_dbh, id.vars = c("log_DBH", "log_CR_H", "sp", "community_fitted_log_DBH"), measure.vars = grep("fitted_log_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

# Plot with log-transformed values
dbh_plot <- ggplot(data_long, aes(x = log_CR_H, y = log_DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = log_CR_H, y = value, group = variable), color = "#4682B4", alpha = 0.3) +  # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_log_DBH), color = "white", linewidth = 1.2) +
  labs(title = "log(DBH) = log(a) + b*log(CR) + c*log(H)",
       x = "log(CR^b * H^c)",
       y = "log(DBH)") +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/H_CR_DBH1_log.png", plot = dbh_plot, width = 12, height = 8, dpi = 300, bg = "white")

dbh_plot

```

# Plot DBH = a(CR*H)^b model (6, 2) (with WD)

```{r}
# Read the data
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_data_lr_dbh1)

# Extract relevant data
log_y <- stan_data_lr_dbh1$log_y
log_x <- stan_data_lr_dbh1$log_x[, 2]
sp <- as.factor(stan_data_lr_dbh1$jj)

str(log_x)
head(log_x)

# Read in the model summary
dbh1 <- tar_read(fit_lr2_summary_power_law1_dbh1)
dbh1

# Filter rows where the variable is "beta"
beta <- dbh1 |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)

print(length(log_a))
print(length(b))
print(length(sp))

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b[i] * log_x
})

fitted_dbh <- exp(fitted_log_y_dbh)

# Extract community-level parameters directly
community_log_a <- dbh1 |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b <- dbh1 |> filter(variable == "gamma[2,1]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b * log_x

community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh1 <- data.frame(DBH = exp(log_y), CRxH = exp(log_x), sp = sp)
data_dbh1$community_fitted_DBH <- community_fitted_DBH

head(data_dbh1)
summary(data_dbh1)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh1[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

data_long <- melt(data_dbh1, id.vars = c("DBH", "CRxH", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh1), value = TRUE))

head(data_long)
summary(data_long)
str(data_long)

# Plotting
dbh_plot1 <- ggplot(data_dbh1, aes(x = CRxH, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CRxH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CR*H",
       y = "DBH") +
  scale_x_log10(labels = label_number()) +
  scale_y_log10(labels = label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/CR*H_DBH1.png", plot = dbh_plot1, width = 12, height = 8, dpi = 300, bg = "white")

dbh_plot1
```

# Fix the range of the CR*H axis

```{r}
# Function to create CR*H sequence for each species based on actual range
create_crxh_sequence <- function(species, log_x, length_out = 100) {
  crxh_values <- exp(log_x[stan_data_lr_dbh1$jj == species])
  min_crxh <- min(crxh_values)
  max_crxh <- max(crxh_values)
  seq(min_crxh, max_crxh, length.out = length_out)
}

# Calculate fitted DBH values for each species using gamma parameters
fitted_dbh_list <- lapply(1:length(log_a), function(i) {
  crxh_seq <- create_crxh_sequence(i, log_x)
  fitted_log_dbh <- log_a[i] + b[i] * log(crxh_seq)
  fitted_dbh <- exp(fitted_log_dbh)

  data.frame(
    CRxH = crxh_seq,
    DBH = fitted_dbh,
    variable = paste0("fitted_DBH_sp_", i)
  )
})

# Combine all species data frames into one
data_long <- do.call(rbind, fitted_dbh_list)
head(data_long)

# Original data for points
data_dbh1 <- data.frame(DBH = exp(log_y), CRxH = exp(log_x), sp = sp)
head(data_dbh1)

# Plotting the DBH against CR*H
dbh_plot1 <- ggplot(data_dbh1, aes(x = CRxH, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CRxH, y = DBH, group = variable), color = "#4682B4", alpha = 0.3) +
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CR*H",
       y = "DBH") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/CRxH_DBH1.png", plot = dbh_plot1, width = 12, height = 8, dpi = 300, bg = "white")

dbh_plot1
```