### 2025-03-26 docs/dbh_up.qmd

---
title: "Comparing DBH models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(tidyverse)
library(ggplot2)
options(width = 120)
```

# Read in data

```{r}
loo_tbl <- tar_read(loo_tbl)
print(loo_tbl, n = Inf)

dbh_models <- loo_tbl |> 
  filter(grepl("dbh($|1|2|3)", model)) |> 
  arrange(looic)
dbh_models

```

```{r}
# Extract relevant data for model 1 (loo_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)

log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]
log_x_h <- stan_data_lr_dbh$log_x[, 3]
sp <- as.factor(stan_data_lr_dbh$jj)

# Read in the model summary for model 
dbh_nou_summary <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
dbh_nou_summary

# Filter rows where the variable is "beta"
beta <- dbh_nou_summary |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

fitted_dbh <- exp(fitted_log_y_dbh)

gamma <- dbh_nou_summary |>
  filter(grepl("gamma", variable))
gamma

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a <- gamma |> filter(variable == "gamma[1]") |> pull(q50)
community_b_cr <- gamma |> filter(variable == "gamma[2]") |> pull(q50)
community_b_h <- gamma |> filter(variable == "gamma[3]") |> pull(q50)

community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CR = exp(log_x_cr), H = exp(log_x_h), sp = sp)
data_dbh$community_fitted_DBH <- community_fitted_DBH
head(data_dbh)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Compute CR^b and H^c for each observation
data_dbh$CR_b <- data_dbh$CR^community_b_cr
data_dbh$H_c <- data_dbh$H^community_b_h
data_dbh$CR_H <- data_dbh$CR_b*data_dbh$H_c

head(data_dbh)

data_long <- reshape2::melt(data_dbh, id.vars = c("DBH", "CR_H", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

data_long <- data_long |> 
  mutate(partial_residuals = DBH - value)

data_summary <- data_long |> 
  group_by(sp, CR_H) |> 
  summarize(mean_residuals = mean(partial_residuals),
            sd_residuals = sd(partial_residuals),
            .groups = 'drop')me(legend.position = "none")

head(data_summary)

ggplot(data_summary, aes(x = CR_H, y = mean_residuals)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_residuals - sd_residuals, ymax = mean_residuals + sd_residuals), width = 0.01) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "CR^b * H^c", y = "Partial Residuals") +
  theme_minimal()

```

```{r}
# Extract relevant data for model (stan_data_lr_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)
# Ensure the correct indexing for variables
log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]  
log_x_h <- stan_data_lr_dbh$log_x[, 3]   
sp <- as.factor(stan_data_lr_dbh$jj)

# Read in the model summary
dbh_nou_summary <- tar_read(fit_lr2_summary_power_law1_nou_dbh)

# Filter rows where the variable is "beta" and extract parameters
beta <- dbh_nou_summary |> filter(grepl("beta", variable))

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(q50)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(q50)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(q50)

# Compute fitted values for DBH in log scale using the multiple regression formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

# Extract gamma values for community-level effects
gamma <- dbh_nou_summary |> filter(grepl("gamma", variable))
community_log_a <- gamma |> filter(variable == "gamma[1]") |> pull(q50)
community_b_cr <- gamma |> filter(variable == "gamma[2]") |> pull(q50)
community_b_h <- gamma |> filter(variable == "gamma[3]") |> pull(q50)

# Calculate community-level fitted line in log scale
community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h

# Create a data frame for plotting (log scale)
data_dbh <- data.frame(log_DBH = log_y, log_CR = log_x_cr, log_H = log_x_h, sp = sp)
data_dbh$community_fitted_log_DBH <- community_fitted_log_y_dbh

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_log_DBH_sp_", i)]] <- fitted_log_y_dbh[, i]
}

# Reshape data for plotting
data_long <- reshape2::melt(data_dbh, id.vars = c("log_DBH", "log_CR", "log_H", "sp", "community_fitted_log_DBH"),
                            measure.vars = grep("fitted_log_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

dim(data_long)
```

# Visualization

## Scatter Plot of Observed vs. Fitted Values
```{r}
> ggplot(data_long, aes(x = value, y = log_DBH, color = sp)) +
+   geom_point() +
+   geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
+   labs(x = "Fitted Log(DBH)", y = "Observed Log(DBH)") +
+   theme_minimal() +
+   theme(legend.position = "none")
```

## Partial Residuals Plot
```{r}
data_long <- data_long |> 
  mutate(partial_residuals = log_DBH - value)

head(data_long)

ggplot(data_long, aes(x = sp, y = partial_residuals, color = sp)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Species", y = "Partial Residuals (Log Scale)") +
  theme_minimal() +
  theme(legend.position = "none")

```

## Residuals vs. Fitted Plot

```{r}
ggplot(data_long, aes(x = value, y = partial_residuals, color = sp)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Fitted Log(DBH)", y = "Partial Residuals") +
  theme_minimal() +
  theme(legend.position = "none")

```