### 2025-03-26 docs/diag.qmd

---
title: "Diagnostic checking"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(cmdstanr)
# tar_make()
```

# Diagnostic-checking

- Note: List of all potential diagnostic file names (lr2, nlr)
  "fit_lr2_diagnostics_power_law1_cr", "fit_lr2_diagnostics_power_law1_dbh",
  "fit_lr2_diagnostics_power_law1_dbh1", "fit_lr2_diagnostics_power_law1_dbh2",
  "fit_lr2_diagnostics_power_law1_dbh3", "fit_lr2_diagnostics_power_law1_h",
  "fit_lr2_diagnostics_power_law1_nou_cr", "fit_lr2_diagnostics_power_law1_nou_dbh",
  "fit_lr2_diagnostics_power_law1_nou_dbh1", "fit_lr2_diagnostics_power_law1_nou_dbh2",
  "fit_lr2_diagnostics_power_law1_nou_dbh3", "fit_lr2_diagnostics_power_law1_nou_h",
  "fit_nlr_diagnostics_gmm1_cr", "fit_nlr_diagnostics_gmm1_h",
  "fit_nlr_diagnostics_gmm1_nou_cr", "fit_nlr_diagnostics_gmm1_nou_h",
  "fit_nlr_diagnostics_weibull1_cr", "fit_nlr_diagnostics_weibull1_h",
  "fit_nlr_diagnostics_weibull1_nou_cr", "fit_nlr_diagnostics_weibull1_nou_h"

```{r}
options(width = 200)
# Load diagnostics data for specific fit_lr2 and fit_nlr models
pl1_cr <- tar_read(fit_lr2_diagnostics_power_law1_cr)
pl1_dbh <- tar_read(fit_lr2_diagnostics_power_law1_dbh)
pl1_dbh1 <- tar_read(fit_lr2_diagnostics_power_law1_dbh1)
pl1_dbh2 <- tar_read(fit_lr2_diagnostics_power_law1_dbh2)
pl1_dbh3 <- tar_read(fit_lr2_diagnostics_power_law1_dbh3)
pl1_h <- tar_read(fit_lr2_diagnostics_power_law1_h)
pl1_nou_cr <- tar_read(fit_lr2_diagnostics_power_law1_nou_cr)
pl1_nou_dbh <- tar_read(fit_lr2_diagnostics_power_law1_nou_dbh)
pl1_nou_dbh1 <- tar_read(fit_lr2_diagnostics_power_law1_nou_dbh1)
pl1_nou_dbh2 <- tar_read(fit_lr2_diagnostics_power_law1_nou_dbh2)
pl1_nou_dbh3 <- tar_read(fit_lr2_diagnostics_power_law1_nou_dbh3)
pl1_nou_h <- tar_read(fit_lr2_diagnostics_power_law1_nou_h)
gmm1_cr <- tar_read(fit_nlr_diagnostics_gmm1_cr)
gmm1_h <- tar_read(fit_nlr_diagnostics_gmm1_h)
gmm1_nou_cr <- tar_read(fit_nlr_diagnostics_gmm1_nou_cr)
gmm1_nou_h <- tar_read(fit_nlr_diagnostics_gmm1_nou_h)
weibull1_cr <- tar_read(fit_nlr_diagnostics_weibull1_cr)
weibull1_h <- tar_read(fit_nlr_diagnostics_weibull1_h)
weibull1_nou_cr <- tar_read(fit_nlr_diagnostics_weibull1_nou_cr)
weibull1_nou_h <- tar_read(fit_nlr_diagnostics_weibull1_nou_h)
```

```{r}
summary(pl1_cr)
summary(pl1_dbh)
summary(pl1_dbh1)
summary(pl1_dbh2)
summary(pl1_dbh3)
summary(pl1_h)
summary(pl1_nou_cr)
summary(pl1_nou_dbh)
summary(pl1_nou_dbh1)
summary(pl1_nou_dbh2)
summary(pl1_nou_dbh3)
summary(pl1_nou_h)

# fit_nlr models summaries
summary(gmm1_cr)
summary(gmm1_h)
summary(gmm1_nou_cr)
summary(gmm1_nou_h)
summary(weibull1_cr)
summary(weibull1_h)
summary(weibull1_nou_cr)
summary(weibull1_nou_h)

```

# Checking ESS_bulk and R_hat
```{r}
pl1_cr <- tar_read(fit_lr2_summary_power_law1_cr)
pl1_dbh <- tar_read(fit_lr2_summary_power_law1_dbh)
pl1_dbh1 <- tar_read(fit_lr2_summary_power_law1_dbh1)
pl1_dbh2 <- tar_read(fit_lr2_summary_power_law1_dbh2)
pl1_dbh3 <- tar_read(fit_lr2_summary_power_law1_dbh3)
pl1_h <- tar_read(fit_lr2_summary_power_law1_h)
pl1_nou_cr <- tar_read(fit_lr2_summary_power_law1_nou_cr)
pl1_nou_dbh <- tar_read(fit_lr2_summary_power_law1_nou_dbh)
pl1_nou_dbh1 <- tar_read(fit_lr2_summary_power_law1_nou_dbh1)
pl1_nou_dbh2 <- tar_read(fit_lr2_summary_power_law1_nou_dbh2)
pl1_nou_dbh3 <- tar_read(fit_lr2_summary_power_law1_nou_dbh3)
pl1_nou_h <- tar_read(fit_lr2_summary_power_law1_nou_h)

gmm1_cr <- tar_read(fit_nlr_summary_gmm1_cr)
gmm1_h <- tar_read(fit_nlr_summary_gmm1_h)
gmm1_nou_cr <- tar_read(fit_nlr_summary_gmm1_nou_cr)
gmm1_nou_h <- tar_read(fit_nlr_summary_gmm1_nou_h)
weibull1_cr <- tar_read(fit_nlr_summary_weibull1_cr)
weibull1_h <- tar_read(fit_nlr_summary_weibull1_h)
weibull1_nou_cr <- tar_read(fit_nlr_summary_weibull1_nou_cr)
weibull1_nou_h <- tar_read(fit_nlr_summary_weibull1_nou_h)
```

```{r}
pl1_cr
pl1_dbh
pl1_dbh1
pl1_dbh2
pl1_dbh3
pl1_h
pl1_nou_cr
pl1_nou_dbh
pl1_nou_dbh1
pl1_nou_dbh2
pl1_nou_dbh3
pl1_nou_h
gmm1_cr
gmm1_h
gmm1_nou_cr
gmm1_nou_h
weibull1_cr
weibull1_h
weibull1_nou_cr
weibull1_nou_h
```

```{r}
summary(pl1_cr)
summary(pl1_dbh)
summary(pl1_dbh1)
summary(pl1_dbh2)
summary(pl1_dbh3)
summary(pl1_h)
summary(pl1_nou_cr)
summary(pl1_nou_dbh)
summary(pl1_nou_dbh1)
summary(pl1_nou_dbh2)
summary(pl1_nou_dbh3)
summary(pl1_nou_h)

# fit_nlr models summaries
summary(gmm1_cr)
summary(gmm1_h)
summary(gmm1_nou_cr)
summary(gmm1_nou_h)
summary(weibull1_cr)
summary(weibull1_h)
summary(weibull1_nou_cr)
summary(weibull1_nou_h)
```

# Divergent transition checking
```{r}
library(targets)
library(posterior)
library(bayesplot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(cmdstanr)
library(stringr)

```

```{r}
divergent_tbl <- tar_read(divergent_tbl)
print(divergent_tbl, n = Inf)
```

# Parallel coordinates plot
```{r}
pl1_h_sam <- tar_read(fit_lr2_draws_power_law1_h)
pl1_h_sam_df <- as_draws_df(pl1_h_sam)
pl1_h_sam_plot <- pl1_h_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]

pl1_h_pairs <- mcmc_pairs(pl1_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_h_pairs
```

```{r}
pl1_cr_sam <- tar_read(fit_lr2_draws_power_law1_cr)
pl1_cr_sam_df <- as_draws_df(pl1_cr_sam)
pl1_cr_sam_df
pl1_cr_sam_plot <- pl1_cr_sam_df[,c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]
pl1_cr_pairs <- mcmc_pairs(pl1_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_cr_pairs

```

```{r}
pl1_dbh_sam <- tar_read(fit_lr2_draws_power_law1_dbh)
pl1_dbh_sam_df <- as_draws_df(pl1_dbh_sam)
pl1_dbh_sam_df
pl1_dbh_sam_plot <- pl1_dbh_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[3,1]",  "gamma[1,2]", "gamma[2,2]", "gamma[3,2]")]
pl1_dbh_pairs <- mcmc_pairs(pl1_dbh_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_dbh_pairs
```

```{r}
pl1_dbh1_sam <- tar_read(fit_lr2_draws_power_law1_dbh1)
pl1_dbh1_sam_df <- as_draws_df(pl1_dbh1_sam)
pl1_dbh1_sam_df
pl1_dbh1_sam_plot <- pl1_dbh1_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]
pl1_dbh1_pairs <- mcmc_pairs(pl1_dbh1_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_dbh1_pairs

```

```{r}
pl1_dbh2_sam <- tar_read(fit_lr2_draws_power_law1_dbh2)
pl1_dbh2_sam_df <- as_draws_df(pl1_dbh2_sam)
pl1_dbh2_sam_df
pl1_dbh2_sam_plot <- pl1_dbh2_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]
pl1_dbh2_pairs <- mcmc_pairs(pl1_dbh2_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_dbh2_pairs
```

```{r}
pl1_dbh3_sam <- tar_read(fit_lr2_draws_power_law1_dbh3)
pl1_dbh3_sam_df <- as_draws_df(pl1_dbh3_sam)
pl1_dbh3_sam_df
pl1_dbh3_sam_plot <- pl1_dbh3_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]")]
pl1_dbh3_pairs <- mcmc_pairs(pl1_dbh3_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_dbh3_pairs
```

```{r}
pl1_nou_h_sam <- tar_read(fit_lr2_draws_power_law1_nou_h)
pl1_nou_h_sam_df <- as_draws_df(pl1_nou_h_sam)
pl1_nou_h_sam_df
pl1_nou_h_sam_plot <- pl1_nou_h_sam_df[, c("gamma[1]", "gamma[2]")]
pl1_nou_h_pairs <- mcmc_pairs(pl1_nou_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_h_pairs
```

```{r}
pl1_nou_cr_sam <- tar_read(fit_lr2_draws_power_law1_nou_cr)
pl1_nou_cr_sam_df <- as_draws_df(pl1_nou_cr_sam)
pl1_nou_cr_sam_df
pl1_nou_cr_sam_plot <- pl1_nou_cr_sam_df[, c("gamma[1]", "gamma[2]")]
pl1_nou_cr_pairs <- mcmc_pairs(pl1_nou_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_cr_pairs
```

```{r}
pl1_nou_dbh_sam <- tar_read(fit_lr2_draws_power_law1_nou_dbh)
pl1_nou_dbh_sam_df <- as_draws_df(pl1_nou_dbh_sam)
pl1_nou_dbh_sam_df
pl1_nou_dbh_sam_plot <- pl1_nou_dbh_sam_df[, c("gamma[1]", "gamma[2]", "gamma[3]")]
pl1_nou_dbh_pairs <- mcmc_pairs(pl1_nou_dbh_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_dbh_pairs
```

```{r}
pl1_nou_dbh1_sam <- tar_read(fit_lr2_draws_power_law1_nou_dbh1)
pl1_nou_dbh1_sam_df <- as_draws_df(pl1_nou_dbh1_sam)
pl1_nou_dbh1_sam_df 
pl1_nou_dbh1_sam_plot <- pl1_nou_dbh1_sam_df[, c("gamma[1]", "gamma[2]")]
pl1_nou_dbh1_pairs <- mcmc_pairs(pl1_nou_dbh1_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_dbh1_pairs
```

```{r}
pl1_nou_dbh2_sam <- tar_read(fit_lr2_draws_power_law1_nou_dbh2)
pl1_nou_dbh2_sam_df <- as_draws_df(pl1_nou_dbh2_sam)
pl1_nou_dbh2_sam_df
pl1_nou_dbh2_sam_plot <- pl1_nou_dbh2_sam_df[, c("gamma[1]", "gamma[2]")]
pl1_nou_dbh2_pairs <- mcmc_pairs(pl1_nou_dbh2_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_dbh2_pairs
```

```{r}
pl1_nou_dbh3_sam <- tar_read(fit_lr2_draws_power_law1_nou_dbh3)
pl1_nou_dbh3_sam_df <- as_draws_df(pl1_nou_dbh3_sam)
pl1_nou_dbh3_sam_df
pl1_nou_dbh3_sam_plot <- pl1_nou_dbh3_sam_df[, c("gamma[1]", "gamma[2]")]
pl1_nou_dbh3_pairs <- mcmc_pairs(pl1_nou_dbh3_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
pl1_nou_dbh3_pairs
```

```{r}
gmm1_h_sam <- tar_read(fit_nlr_draws_gmm1_h)
gmm1_h_sam_df <- as_draws_df(gmm1_h_sam)
gmm1_h_sam_df
gmm1_h_sam_plot <- gmm1_h_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]", "gamma[1,3]",  "gamma[2,3]")]
gmm1_h_pairs <- mcmc_pairs(gmm1_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
gmm1_h_pairs
```

```{r}
gmm1_cr_sam <- tar_read(fit_nlr_draws_gmm1_cr)
gmm1_cr_sam_df <- as_draws_df(gmm1_cr_sam)
gmm1_cr_sam_df
gmm1_cr_sam_plot <- gmm1_cr_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]", "gamma[1,3]",  "gamma[2,3]")]
gmm1_cr_pairs <- mcmc_pairs(gmm1_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
gmm1_cr_pairs
```

```{r}
gmm1_nou_h_sam <- tar_read(fit_nlr_draws_gmm1_nou_h)
gmm1_nou_h_sam_df <- as_draws_df(gmm1_nou_h_sam)
gmm1_nou_h_sam_df
gmm1_nou_h_sam_plot <- gmm1_nou_h_sam_df[, c("gamma[1]", "gamma[2]", "gamma[3]")]
gmm1_nou_h_pairs <- mcmc_pairs(gmm1_nou_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
gmm1_nou_h_pairs
```

```{r}
gmm1_nou_cr_sam <- tar_read(fit_nlr_draws_gmm1_nou_cr)
gmm1_nou_cr_sam_df <- as_draws_df(gmm1_nou_cr_sam)
gmm1_nou_cr_sam_df
gmm1_nou_cr_sam_plot <- gmm1_nou_cr_sam_df[, c("gamma[1]", "gamma[2]", "gamma[3]")]
gmm1_nou_cr_pairs <- mcmc_pairs(gmm1_nou_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
gmm1_nou_cr_pairs
```

```{r}
weibull1_h_sam <- tar_read(fit_nlr_draws_weibull1_h)
weibull1_h_sam_df <- as_draws_df(weibull1_h_sam)
weibull1_h_sam_df
weibull1_h_sam_plot <- weibull1_h_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]", "gamma[1,3]",  "gamma[2,3]")]
weibull1_h_pairs <- mcmc_pairs(weibull1_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
weibull1_h_pairs
```

```{r}
weibull1_cr_sam <- tar_read(fit_nlr_draws_weibull1_cr)
weibull1_cr_sam_df <- as_draws_df(weibull1_cr_sam)
weibull1_cr_sam_df
weibull1_cr_sam_plot <- weibull1_cr_sam_df[, c("gamma[1,1]", "gamma[2,1]", "gamma[1,2]", "gamma[2,2]", "gamma[1,3]",  "gamma[2,3]")]
weibull1_cr_pairs <- mcmc_pairs(weibull1_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
weibull1_cr_pairs
```

```{r}
weibull1_nou_h_sam <- tar_read(fit_nlr_draws_weibull1_nou_h)
weibull1_nou_h_sam_df <- as_draws_df(weibull1_nou_h_sam)
weibull1_nou_h_sam
weibull1_nou_h_sam_plot <- weibull1_nou_h_sam_df[, c("gamma[1]", "gamma[2]", "gamma[3]")]
weibull1_nou_h_pairs <- mcmc_pairs(weibull1_nou_h_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
weibull1_nou_h_pairs
```

```{r}
weibull1_nou_cr_sam <- tar_read(fit_nlr_draws_weibull1_nou_cr)
weibull1_nou_cr_sam_df <- as_draws_df(weibull1_nou_cr_sam)
weibull1_nou_cr_sam_df
weibull1_nou_cr_sam_plot <- weibull1_nou_cr_sam_df[, c("gamma[1]", "gamma[2]", "gamma[3]")]
weibull1_nou_cr_pairs <- mcmc_pairs(weibull1_nou_cr_sam_plot, diag_fun = "hist", off_diag_fun = "hex")
weibull1_nou_cr_pairs
```