### 2025-03-26 docs/gmm_cr.qmd

---
title: "Checking gmm model"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load the required packages

```{r}
library(dplyr)
library(readr)
library(targets)
library(stringr)
library(cmdstanr)
library(tidyverse)
library(reshape2)
library(bayesplot)
library(cmdstanr) 
library(posterior) 
```

# Data preparation
```{r}
tallo_wd_200_df <- read_csv("data/tallo_wd_200.csv")
tallo_wd_200_df
stan_data_nlr_cr <- tar_read(stan_data_nlr_cr)
str(stan_data_nlr_cr)

options(width = 100)
gmm_cr <- tar_read(fit_nlr_mcmc_gmm_cr)
gmm_cr

gmm2_cr <- tar_read(fit_nlr_mcmc_gmm2_cr)
gmm2_cr

gmm_beta <- gmm_cr$summary(variable = "beta")
gmm2_beta <- gmm2_cr$summary(variable = "beta")

gmm_beta
gmm2_beta

summary(gmm_beta)
summary(gmm2_beta)

# Filter beta parameters with small ESS
small_ess_gmm_beta <- gmm_beta |>
  filter(ess_bulk < 300) |>
  arrange(ess_bulk) |>
  mutate(species_id = as.integer(str_extract(variable, "\\d+")))

small_ess_gmm1_beta <- gmm1_beta |>
  filter(ess_bulk < 300) |>
  arrange(ess_bulk)
small_ess_gmm1_beta

high_rhat_gmm1_beta <- gmm1_beta |>
   filter(rhat > 1.1) |> 
   arrange(rhat)

high_rhat_gmm1_beta
```

# Species with low ESS of gmm_cr and gmm2_cr model

```{r}
beta1 <- gmm_beta |> filter(str_detect(variable, "beta\\[\\d+,1\\]"))
beta1
summary(beta1)

small_ess_beta1 <- beta1 |>
  filter(ess_bulk < 300) |>
  arrange(ess_bulk) |>
  mutate(species_id = as.integer(str_extract(variable, "\\d+")))

small_ess_beta1

beta2 <- gmm_beta |> filter(str_detect(variable, "beta\\[\\d+,2\\]"))

beta2
summary(beta2)

beta3 <- gmm_beta |> filter(str_detect(variable, "beta\\[\\d+,3\\]"))

beta3
summary(beta3)
```

```{r}
gamma_trace <- gmm_cr$draws("gamma")
mcmc_trace(gamma_trace)

mcmc_trace(gamma_trace) + ggtitle("Trace Plot for gmm_cr Gamma")

gamma2_trace <- gmm2_cr$draws("gamma")
mcmc_trace(gamma2_trace) + ggtitle("Trace Plot for gmm2_cr Gamma")

mcmc_trace(gamma_trace) + ggtitle("Trace plot of gmm1_cr")
```