### 2025-03-26 docs/h_cr_bb.qmd

---
title: "Overlaps species"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r}
library(readr)
library(ggplot2)
library(tidyverse)
```

# Read the CSV file
```{r}
hcr <- read_csv("data/bb_allometry_re.csv")
summary(hcr)
head(hcr, n=100)
summary(hcr)
```

# Convert sp to a factor
```{r}
hcr$sp <- factor(hcr$sp)
hcr$life <- factor(hcr$life)
#Construct the histogram 
par(mfrow=c(2,1))
hist(hcr$tree_height)
hist(hcr$crown_width_we) # positive skewed
```

#log transform
```{r}
hcr$crown_width_we.ln <- log(hcr$crown_width_we)
hcr$tree_height.ln <- log(hcr$tree_height)
summary(hcr)
```

#check
```{r}
par(mfrow=c(2,1))
hist(hcr$tree_height.ln) #look better
hist(hcr$crown_width_we.ln) 
```

#plot the relationship of interest
```{r}
plot(crown_width_we.ln ~ tree_height.ln, data = hcr)
plot(hcr$tree_height.ln ,hcr$crown_width_we.ln ,col= c("red", "blue", "green")[as.numeric(hcr$sp)])
```

# Remove rows with -Inf and NA in crown_width_we.ln, tree_height.ln
```{r}
hcr <- hcr[is.finite(hcr$crown_width_we.ln) & !is.na(hcr$crown_width_we.ln), ]
hcr <- hcr[is.finite(hcr$tree_height.ln) & !is.na(hcr$tree_height.ln), ]

# Check the summary again
summary(hcr$crown_width_we.ln)
summary(hcr$tree_height.ln)
```

# Fit a linear model
```{r}
hcr.rg <- lm(crown_width_we.ln ~ tree_height.ln, data = hcr)
summary(hcr.rg)
plot(hcr.rg)
# Create a line graph using ggplot2 with regression lines for each species
ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = sp, colour = sp, shape = sp)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1, aes(group = sp)) +
  theme(legend.position = "none")  # Optional: To remove legend for shapes due to the large number of levels
```

# Relationship between Tree Height and Crown Width by Life
```{r}
ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = life, colour = life, shape = life)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1, aes(group = life)) +
  theme_minimal() +
  labs(title = "Relationship between Tree Height and Crown Width by Life",
       subtitle = "Regression lines for each life, colored and shaped by life",
       color = "Life", shape = "Life")
```

#separate plots for each level of the life variable
```{r}
ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = life, colour = life)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1, aes(group = life)) +
  facet_wrap(~life, scales = "free") +
  theme_minimal() +
  labs(title = "Relationship between Tree Height and Crown Width by Life",
       subtitle = "Regression lines for each life, colored by life",
       color = "Life")
```

# 
```{r}
ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = life, colour = life)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1, aes(group = sp)) +
  facet_wrap(~life, scales = "free") +
  theme_minimal() +
  labs(title = "Relationship between Tree Height and Crown Width by Life",
       subtitle = "Regression lines for each life, colored by life",
       color = "Life")
```

# facetting the plot by the life variable and showing regression lines for each life category, colored by the life variable.
```{r}
bb_plot <- ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = interaction(sp, life), colour = life)) +
  geom_point(size = 5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 1, aes(group = interaction(sp, life))) +
  facet_wrap(~life, scales = "free") +
  theme_minimal() +
  labs(title = "Relationship between Tree Height and Crown Width by Life",
       subtitle = "Regression lines for each species within different life categories",
       color = "Life")

# Save the plot
ggsave("figs/bb.png", plot = bb_plot, width = 10, height = 6)

```

```{r}

bb <- ggplot(data = hcr, aes(y = crown_width_we.ln, x = tree_height.ln, group = interaction(sp, life), colour = life)) +
  geom_point(size = 7, alpha = 0.7) +  # Increase point size and add transparency
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, size = 0.7, aes(group = interaction(sp, life))) +  # Reduce line size
  facet_wrap(~life, scales = "free") +
  theme_minimal() +
  labs(title = "Relationship between Tree Height and Crown Width by Life",
       subtitle = "Regression lines for each species within different life categories",
       color = "Life")

# Save the plot
ggsave("figs/bb_2.png", plot = bb, width = 10, height = 6)
```