### 2025-03-26 docs/pl_cr.qmd

---
title: "Example for Power-law DBH-CR"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries
```{r}
library(targets)
library(ggplot2)
library(dplyr)
library(stringr)
library(reshape2)
library(gridExtra)
library(readr)
```

# Data preparation
```{r}
# Load MCMC results and Stan data
pl_cr <- tar_read(fit_lr_mcmc_power_law_cr)
stan_lr_data_cr <- tar_read(stan_data_lr_cr)
str(stan_lr_data_cr)

```

# Extract relevant data
```{r}
log_y <- stan_lr_data_cr$log_y
log_x <- stan_lr_data_cr$log_x
sp <- as.factor(stan_lr_data_cr$jj)

str(log_x)
head(log_x) # Use log_x[, 2] for the fitted values
summary(exp(log_y))
summary(exp(log_x))

# Extract beta parameters
beta <- pl_cr$summary(variable = "beta")
beta

# Extract the mean values of beta parameters for each species
log_a1 <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b1 <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)

# Verify the lengths of the extracted parameters
print(length(log_a1))
print(length(b1))
print(length(sp))

# Compute fitted values for CR using the power law formula
fitted_log_y_cr <- sapply(1:length(log_a1), function(i) {
  log_a1[i] + b1[i] * log_x[, 2]  
})

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a1 <- mean(log_a1)
community_b1 <- mean(b1)
community_fitted_log_y_cr <- community_log_a1 + community_b1 * log_x[, 2]
community_fitted_CR <- exp(community_fitted_log_y_cr)

# Create a data frame for plotting
data_cr <- data.frame(CR = exp(log_y), DBH = exp(log_x[, 2]), sp = sp)
head(data_cr)
summary(data_cr)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a1)) {
  data_cr[[paste0("fitted_CR_sp_", i)]] <- exp(fitted_log_y_cr[, i])
}

data_long <- melt(data_cr, id.vars = c("CR", "DBH", "sp"), measure.vars = grep("fitted_CR_sp_", names(data_cr), value = TRUE))

head(data_long)
summary(data_long)
str(data_long)
str(data_cr)

```

# Plot observed CR vs DBH with fitted lines for each species and community-level line
```{r}
pl_cr_plot <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10(labels = scales::comma) +
  scale_y_log10(labels = scales::comma) +
  theme_minimal()

# Save the plot
ggsave("figs/plot_CR_vs_DBH.png", plot = pl_cr_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

# Display the plot
print(pl_cr_plot)

```

```{r}
pl_cr_plot1 <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_continuous(trans = "log10", breaks = scales::trans_breaks("log10", function(x) 10^x),
                     labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_y_continuous(trans = "log10", breaks = scales::trans_breaks("log10", function(x) 10^x),
                     labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  theme_minimal()

# Save the plot
ggsave("figs/plot1_CR_vs_DBH.png", plot = pl_cr_plot1, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

# Display the plot
print(pl_cr_plot1)
```

```{r}
library(scales)

# Plot observed CR vs DBH with fitted lines for each species and community-level line
pl_cr_plot5 <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10(labels = comma) +
  scale_y_log10(labels = comma) +
  theme_minimal()

# Save the plot
ggsave("figs/plot5_CR_vs_DBH.png", plot = pl_cr_plot5, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

```

# Plot for some sp

```{r}
# Identify the top 5 species with the most observations
top_5_species <- data_cr |>
  count(sp, sort = TRUE) |>
  top_n(5, n) |>
  pull(sp)
top_5_species
# Create and save plots for the top 5 species
output_dir <- "figs/top_5_species_plots"

if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

for (species in top_5_species) {
  species_data_cr <- data_cr %>% filter(sp == species)
  species_data_long <- data_long %>% filter(sp == species)

  species_plot <- ggplot(species_data_cr, aes(x = DBH, y = CR)) +
    geom_point(alpha = 0.5, color = "grey") +
    geom_line(data = species_data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) +
    geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
    labs(title = paste("Scatter Plot of CR vs DBH for Species", species),
         x = "DBH (cm)",
         y = "CR (m)") +
    scale_x_log10(labels = scales::comma, limits = c(1, 150)) +
    scale_y_log10(labels = scales::comma) +
    theme_minimal()

  plot_filename <- file.path(output_dir, paste0("plot_CR_vs_DBH_species_", species, ".png"))
  ggsave(plot_filename, plot = species_plot, width = 8, height = 6, units = "in", dpi = 300, bg = "white")
    print(species_plot)
}
```

# How wood density effect tree architecture differentiation

```{r}
library(dplyr)
str(stan_lr_data_cr)
u <- stan_lr_data_cr$u[2, ]

# # Load the tallo_wf_df_200 dataset
# tallo_wf_df_200 <- read_csv("data/tallo_wd_200.csv")

# # Calculate mean wd for each species and filter species with at least 5 observations
# wd_df <- tallo_wf_df_200 |>
#   group_by(sp) |>
#   filter(n() >= 5) |>
#   select(sp, mean_wd)

# Create a data frame for plotting
data_wd_plot <- data.frame(
  a1 = exp(log_a1),
  b1 = b1,
  wd = u
)

head(data_wd_plot)
summary(data_wd_plot)

# Plot for intercepts (a)
plot_a <- ggplot(data_wd_plot, aes(x = wd, y = a1)) +
  geom_point(color = "grey", alpha = 0.5) +
  geom_smooth(method = "lm", color = "black", fill = "grey") +
  labs(title = "(a)", x = "WD_s", y = "Intercept a") +
  theme_minimal()

plot_a

# Plot for exponents (b)
plot_b <- ggplot(data_wd_plot, aes(x = wd, y = b1)) +
  geom_point(color = "orange", alpha = 0.5) +
  geom_smooth(method = "lm", color = "black", fill = "orange") +
  labs(title = "(b)", x = "WD_s", y = "Exponent b") +
  theme_minimal()

plot_b

# Arrange the plots vertically
combined_plot <- grid.arrange(plot_a, plot_b, ncol = 1)

# Save the plot
ggsave("figs/intercept_exponent_plot.png", plot = combined_plot, width = 8, height = 10, units = "in", dpi = 300, bg = "white")

# Display the plot
print(combined_plot)
```

```{r}
pl_cr_plot2 <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of CR vs DBH with Fitted Lines for Each Species and Community-Level Line",
       x = "DBH (cm)",
       y = "CR (m)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

# Save the plot
ggsave("figs/plot2_CR_vs_DBH.png", plot = pl_cr_plot2, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

# Display the plot
print(pl_cr_plot2)
```

```{r}
# Plot observed CR vs DBH with fitted lines for each species and community-level line
pl_cr_plot3 <- ggplot(data_cr, aes(x = DBH, y = CR)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = DBH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_CR), color = "white", linewidth = 1.2) +
  labs(title = "Scatter Plot of Crown Radius vs Diameter with Fitted Lines for Each Species and Community-Level Line",
       x = "Diameter at Breast Height (cm)",
       y = "Crown Radius (m)") +
  scale_x_log10(labels = scales::comma) +
  scale_y_log10(labels = scales::comma) +
  annotation_logticks(sides = "bl") +  # Add logarithmic ticks on both axes
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

# Save the plot
ggsave("figs/plot3_CR_vs_DBH.png", plot = pl_cr_plot3, width = 8, height = 6, units = "in", dpi = 300, bg = "white")

pl_cr_plot3
```

# Comparison of pl_cr and pl_nou_cr

```{r}
# Load necessary libraries
library(targets)
library(cmdstanr)
library(posterior)
library(loo)
```
```{r}
pl_cr <- tar_read(fit_lr2_mcmc_power_law_cr)
pl_nou_cr <- tar_read(fit_lr2_mcmc_power_law_nou_cr)
```

```{r}
# Extract log-likelihood for both models
log_lik_cr <- pl_cr$draws(variables = "log_lik", format = "draws_matrix")
log_lik_cr_matrix <- as_draws_matrix(log_lik_cr)

log_lik_nou_cr <- pl_nou_cr$draws(variables = "log_lik", format = "draws_matrix")
log_lik_nou_cr_matrix <- as_draws_matrix(log_lik_nou_cr)

# Calculate LOO for both models
loo_cr <- loo(log_lik_cr_matrix, reloo = TRUE, save_psis = TRUE)
loo_nou_cr <- loo(log_lik_nou_cr_matrix, reloo = TRUE, save_psis = TRUE)

# Print LOO results
print(loo_cr)
print(loo_nou_cr)

# Compare models
model_comparison <- loo_compare(loo_cr, loo_nou_cr)
print(model_comparison)
```