### 2025-03-26 docs/pl_dbh.qmd

---
title: "Comparing DBH models"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/workspaces/tree-allometry")
```

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

# Load necessary libraries

```{r}
library(targets)
library(cmdstanr)
library(posterior)
library(loo)
library(dplyr)
library(ggplot2)
library(stringr)
library(reshape2)
library(scales)
# Define and run the targets pipeline
# tar_make()
```

- Note:

MCMC setting:
  chains = 3,
  parallel_chains = 3,
  iter_warmup = 1000,
  iter_sampling = 1000,
  adapt_delta = 0.9,
  max_treedepth = 15,

# Diagnostics checking

```{r}
options(width = 100)
# Assuming you have CmdStanMCMC or CmdStanFit objects for each model
# Load diagnostics for each model
diagnostics_dbh <- tar_read(fit_lr1_diagnostics_power_law_dbh)
diagnostics_dbh1 <- tar_read(fit_lr1_diagnostics_power_law_dbh1)
diagnostics_dbh2 <- tar_read(fit_lr1_diagnostics_power_law_dbh2)
diagnostics_dbh3 <- tar_read(fit_lr1_diagnostics_power_law_dbh3)
diagnostics_nou_dbh <- tar_read(fit_lr1_diagnostics_power_law_nou_dbh)
diagnostics_nou_dbh1 <- tar_read(fit_lr1_diagnostics_power_law_nou_dbh1)
diagnostics_nou_dbh2 <- tar_read(fit_lr1_diagnostics_power_law_nou_dbh2)
diagnostics_nou_dbh3 <- tar_read(fit_lr1_diagnostics_power_law_nou_dbh3)

summary(diagnostics_dbh)
summary(diagnostics_dbh1)
summary(diagnostics_dbh2)
summary(diagnostics_dbh3)
summary(diagnostics_nou_dbh)
summary(diagnostics_nou_dbh1)
summary(diagnostics_nou_dbh2)
summary(diagnostics_nou_dbh3)
```

- All models exhibit stable treedepth, no divergences, and consistent behavior across chains and iterations. The mean energy levels, acceptance statistics, and step sizes vary slightly among the models but remain within acceptable ranges, indicating well-mixed chains and reliable diagnostics.

# Extract log-likelihood values

```{r}
draws_dbh <- tar_read(fit_lr1_draws_power_law_dbh)
draws_dbh1 <- tar_read(fit_lr1_draws_power_law_dbh1)
draws_dbh2 <- tar_read(fit_lr1_draws_power_law_dbh2)
draws_dbh3 <- tar_read(fit_lr1_draws_power_law_dbh3)
draws_nou_dbh <- tar_read(fit_lr1_draws_power_law_nou_dbh)
draws_nou_dbh1 <- tar_read(fit_lr1_draws_power_law_nou_dbh1)
draws_nou_dbh2 <- tar_read(fit_lr1_draws_power_law_nou_dbh2)
draws_nou_dbh3 <- tar_read(fit_lr1_draws_power_law_nou_dbh3)
```

```{r}
log_lik_dbh <- as.matrix(draws_dbh |> select(starts_with("log_lik")))
log_lik_dbh1 <- as.matrix(draws_dbh1 |> select(starts_with("log_lik")))
log_lik_dbh2 <- as.matrix(draws_dbh2 |> select(starts_with("log_lik")))
log_lik_dbh3 <- as.matrix(draws_dbh3 |> select(starts_with("log_lik")))
log_lik_nou_dbh <- as.matrix(draws_nou_dbh |> select(starts_with("log_lik")))
log_lik_nou_dbh1 <- as.matrix(draws_nou_dbh1 |> select(starts_with("log_lik")))
log_lik_nou_dbh2 <- as.matrix(draws_nou_dbh2 |> select(starts_with("log_lik")))
log_lik_nou_dbh3 <- as.matrix(draws_nou_dbh3 |> select(starts_with("log_lik")))
```

# Compute LOOIC for each model

```{r}
loo_dbh <- loo(log_lik_dbh)
loo_dbh1 <- loo(log_lik_dbh1)
loo_dbh2 <- loo(log_lik_dbh2)
loo_dbh3 <- loo(log_lik_dbh3)
loo_nou_dbh <- loo(log_lik_nou_dbh)
loo_nou_dbh1 <- loo(log_lik_nou_dbh1)
loo_nou_dbh2 <- loo(log_lik_nou_dbh2)
loo_nou_dbh3 <- loo(log_lik_nou_dbh3)

# loo_dbh
# loo_dbh1
# loo_dbh2
# loo_dbh3
# loo_nou_dbh
# loo_nou_dbh1
# loo_nou_dbh2
# loo_nou_dbh3
```

```{r}
# Function to save LOO plots
save_loo_plot <- function(loo_object, filename) {
  png(filename, width = 8, height = 6, units = "in", res = 300)
  plot(loo_object)
  dev.off()
}

# Create and save the plots with titles
save_loo_plot(loo_dbh, "figs/dbh/dbh_loo_plot.png")
save_loo_plot(loo_dbh1, "figs/dbh/dbh1_loo_plot.png")
save_loo_plot(loo_dbh2, "figs/dbh/dbh2_loo_plot.png")
save_loo_plot(loo_dbh3, "figs/dbh/dbh3_loo_plot.png")
save_loo_plot(loo_nou_dbh, "figs/dbh/nou_dbh_loo_plot.png")
save_loo_plot(loo_nou_dbh1, "figs/dbh/nou_dbh1_loo_plot.png")
save_loo_plot(loo_nou_dbh2, "figs/dbh/nou_dbh2_loo_plot.png")
save_loo_plot(loo_nou_dbh3, "figs/dbh/nou_dbh3_loo_plot.png")
```

# Compare models using LOOIC
```{r}
comparison <- loo_compare(loo_dbh, loo_dbh1, loo_dbh2, loo_dbh3, loo_nou_dbh, loo_nou_dbh1, loo_nou_dbh2, loo_nou_dbh3)

# Print comparison results
print(comparison)
```

- Note:

print(comparison)
       elpd_diff se_diff 
model1      0.0       0.0
model5     -0.9       2.1
model6  -2705.2      93.2
model2  -2707.0      93.2
model8 -10973.9     169.4
model4 -10975.6     169.4
model7 -17706.9     213.3
model3 -17707.4     213.3

```{r}
# Create a data frame for comparison
model_names <- c("loo_dbh", "loo_dbh1", "loo_dbh2", "loo_dbh3", "loo_nou_dbh", "loo_nou_dbh1", "loo_nou_dbh2", "loo_nou_dbh3")
elpd_diff <- comparison[, "elpd_diff"]
se_diff <- comparison[, "se_diff"]

# Map the results back to the model names based on loo_compare output
comparison_df <- data.frame(
  Model = model_names[c(1, 5, 6, 2, 8, 4, 7, 3)],  # Reorder based on loo_compare output order
  ELPD_Difference = round(elpd_diff, 1),
  SE_Difference = round(se_diff, 1),
  Formula = c(
    "DBH = a * (CR^b * H^c)",
    "DBH = a * (CR^b * H^c)",
    "DBH = a * (CR * H)^b",
    "DBH = a * (CR * H)^b",
    "DBH = a * H^b",
    "DBH = a * H^b",
    "DBH = a * CR^b",
    "DBH = a * CR^b"
  )
)
comparison_df
# Print the comparison table
library(knitr)
kable(comparison_df, caption = "LOOIC Comparison of DBH Models")
```

- Best Model:
Model 1 (loo_dbh) with the formula DBH = a * (CR^b * H^c) is the best-performing model.

The best-performing models (1, 5, 6, 2) use the formula DBH = a * (CR^b * H^c), DBH = a * (CR * H)^b, indicating that including both CR and H is crucial for accurate DBH estimation.

Models without wood density (indicated by the prefix “nou” in  table) do not show significantly worse performance compared to their counterparts with wood density. For example, model 1 (loo_dbh) and model 5 (loo_nou_dbh) have very similar ELPD differences, suggesting that simpler models without wood density are nearly as effective.

While simpler models without wood density perform well, the model incorporating wood density as a species-level predictor (model 1: DBH = a * (CR^b * H^c)) shows the best overall performance. This indicates that, despite the minimal difference in performance, including wood density may still provide some benefit.

- Key findings:

Accurate DBH Estimation: Incorporating both tree height (H) and crown radius (CR) is essential for accurate DBH estimation.

Wood Density: While the inclusion of wood density does not significantly improve the model’s performance, it provides a slight benefit, making the model with wood density the best performer overall.

# Plot DBH = aCR^b * H^c model (1, 5)

```{r}
# Extract relevant data for model 1 (loo_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)

log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]
log_x_h <- stan_data_lr_dbh$log_x[, 3]
sp <- as.factor(stan_data_lr_dbh$jj)

str(log_x_cr)
str(log_x_h)

# Read in the model summary for model 
dbh_summary <- tar_read(fit_lr1_summary_power_law_dbh)
colnames(dbh_summary)

# Filter rows where the variable is "beta"
beta <- dbh_summary |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(mean)

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

fitted_dbh <- exp(fitted_log_y_dbh)

gamma <- dbh_summary |>
  filter(grepl("gamma", variable))
gamma

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_cr <- gamma |> filter(variable == "gamma[2,1]") |> pull(mean)
community_b_h <- gamma |> filter(variable == "gamma[3,1]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CR = exp(log_x_cr), H = exp(log_x_h), sp = sp)
data_dbh$community_fitted_DBH <- community_fitted_DBH
head(data_dbh)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Combine CR and H for plotting
data_dbh$CR_H <- data_dbh$CR^community_b_cr * data_dbh$H^community_b_h

data_long <- reshape2::melt(data_dbh, id.vars = c("DBH", "CR_H", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

dbh_plot <- ggplot(data_dbh, aes(x = CR_H, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CR_H, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*CR^b * H^c",
       x = "CR_H",
       y = "DBH") +
  scale_x_log10(labels = label_number()) +
  scale_y_log10(labels = label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/H_CR_DBH1.png", plot = dbh_plot, width = 12, height = 8, dpi = 300, bg = "white")
```

- Note: Plot looks weird

# Fixxing code
```{r}
# Extract relevant data for model 1 (loo_dbh)
stan_data_lr_dbh <- tar_read(stan_data_lr_dbh)
str(stan_data_lr_dbh)
head(stan_data_lr_dbh$log_x)

log_y <- stan_data_lr_dbh$log_y
log_x_cr <- stan_data_lr_dbh$log_x[, 2]
log_x_h <- stan_data_lr_dbh$log_x[, 3]
sp <- as.factor(stan_data_lr_dbh$jj)

head(log_x_cr)
head(log_x_h)

# Read in the model summary for model 
dbh_summary <- tar_read(fit_lr1_summary_power_law_dbh)

# Filter rows where the variable is "beta"
beta <- dbh_summary |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b_cr <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)
b_h <- beta |> filter(str_detect(variable, "beta\\[3,\\d+\\]")) |> pull(mean)

# Debug: Check the lengths of log_a, b_cr, and b_h
cat("Lengths of log_a, b_cr, b_h:\n", length(log_a), length(b_cr), length(b_h), "\n")

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b_cr[i] * log_x_cr + b_h[i] * log_x_h
})

fitted_dbh <- exp(fitted_log_y_dbh)
gamma <- dbh_summary |>
  filter(grepl("gamma", variable))
gamma

# Calculate community-level fitted line by averaging the species-level coefficients
community_log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b_cr <- gamma |> filter(variable == "gamma[2,1]") |> pull(mean)
community_b_h <- gamma |> filter(variable == "gamma[3,1]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b_cr * log_x_cr + community_b_h * log_x_h
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CR = exp(log_x_cr), H = exp(log_x_h), sp = sp)
data_dbh$community_fitted_DBH <- community_fitted_DBH
head(data_dbh)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Combine CR and H for plotting
data_dbh$CR_H <- data_dbh$CR^community_b_cr * data_dbh$H^community_b_h

data_long <- reshape2::melt(data_dbh, id.vars = c("DBH", "CR_H", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)

# Plotting with no background
dbh_plot <- ggplot(data_dbh, aes(x = CR_H, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CR_H, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(x = CR_H, y = community_fitted_DBH), color = "red", linewidth = 1.2) + # Community fit line in red
  labs(title = "DBH = a * CR^b * H^c",
       x = "CR^b * H^c",
       y = "DBH") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_number()) +
  theme_minimal() +
  theme(panel.background = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"),
        plot.background = element_blank(),
        legend.background = element_blank(),
        legend.box.background = element_blank())

# Save the plot
ggsave(filename = "figs/dbh/H_CR_DBH1_2.png", plot = dbh_plot, width = 12, height = 8, dpi = 300, bg = "transparent")

# Print the plot to visualize in R
print(dbh_plot)
```

- Note: still weird

```{r}
# Create a data frame for plotting
data_dbh <- data.frame(log_DBH = log_y, log_CR = log_x_cr, log_H = log_x_h, sp = sp)
head(data_dbh)

# Fit the overall model
fit <- lm(log_DBH ~ log_CR + log_H, data = data_dbh)
summary(fit)

# Generate overall fitted values
data_dbh$fitted_log_DBH <- predict(fit, newdata = data_dbh)
head(data_dbh)

species_fitted_values <- data_dbh |>
  group_by(sp) |>
  do({
    fit_sp <- lm(log_DBH ~ log_CR + log_H, data = .)
    data.frame(
      log_DBH = .$log_DBH,
      fitted_log_DBH_sp = predict(fit_sp, newdata = .),
      sp = .$sp
    )
  })

# Merge the species fitted values back to the main data frame
data_dbh <- merge(data_dbh, species_fitted_values, by = c("log_DBH", "sp"))
head(data_dbh)

plot_observed_vs_predicted <- ggplot(data_dbh, aes(x = fitted_log_DBH, y = log_DBH)) +
  geom_point(alpha = 0.3, color = "grey") +
  geom_abline(intercept = 0, slope = 1, color = "#4682B4", linetype = "dashed") + # Perfect fit line
  geom_smooth(aes(group = sp, color = ), method = "lm", se = FALSE, alpha = 0.5) + # Species-specific fitted lines
  labs(title = "DBH = aCR^b*H^c",
       x = "log(CR) + log(H)",
       y = "log(DBH)") +
  theme_minimal() +
  guides(color = "none") # Remove the legend
plot_observed_vs_predicted
```

# Plot DBH = a(CR*H)^b model (6, 2)

```{r}
comparison1 <- loo_compare(loo_dbh1, loo_dbh2, loo_dbh3, loo_nou_dbh1, loo_nou_dbh2, loo_nou_dbh3)

# Print comparison results
print(comparison1)
```

# Read the data

```{r}
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_data_lr_dbh1)

# Extract relevant data
log_y <- stan_data_lr_dbh1$log_y
log_x <- stan_data_lr_dbh1$log_x
sp <- as.factor(stan_data_lr_dbh1$jj)
# u <- stan_data_lr_dbh1$u

str(log_x)
head(log_x)

# Read in the model summary
dbh1 <- tar_read(fit_lr1_summary_power_law_dbh1)
dbh1

# Filter rows where the variable is "beta"
beta <- dbh1 |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)

print(length(log_a))
print(length(b))
print(length(sp))

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b[i] * log_x[, 2]
})

fitted_dbh <- exp(fitted_log_y_dbh)

# Extract the gamma values
gamma <- dbh1 |> filter(grepl("gamma", variable))

# Compute the community-level intercept and slope using gamma
community_log_a <- gamma |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b <- gamma |> filter(variable == "gamma[2,1]") |> pull(mean)

# Compute fitted values for the community-level line
community_fitted_log_y_dbh <- community_log_a + community_b * log_x[, 2]
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CRxH = exp(log_x[, 2]), sp = sp)

data_dbh$community_fitted_DBH <- community_fitted_DBH

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

# Transform data for plotting
data_long <- data_dbh |> melt(id.vars = c("DBH", "CRxH", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)
summary(data_long)
str(data_long)
```

```{r}
# Plotting
dbh_plot1 <- ggplot(data_dbh, aes(x = CRxH, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CRxH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CR*H",
       y = "DBH") +
  scale_x_log10(labels = label_number()) +
  scale_y_log10(labels = label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/CR*H_DBH.png", plot = dbh_plot1, width = 12, height = 8, dpi = 300, bg = "white")

```

```{r}
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_data_lr_dbh1)

# Extract relevant data
log_y <- stan_data_lr_dbh1$log_y
log_x <- stan_data_lr_dbh1$log_x
sp <- as.factor(stan_data_lr_dbh1$jj)
# u <- stan_data_lr_dbh1$u

str(log_x)
head(log_x)

# Read in the model summary
dbh1 <- tar_read(fit_lr1_summary_power_law_dbh1)
dbh1

# Community parameters
community_log_a <- dbh1 |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b <- dbh1 |> filter(variable == "gamma[1,2]") |> pull(mean)
community_log_a
community_b

community_fitted_log_y_dbh <- community_log_a + community_b * log_x[, 2]
community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CRxH = exp(log_x[, 2]), community_fitted_DBH, sp = sp)

data_dbh$community_fitted_DBH <- community_fitted_DBH
head(data_dbh)
str(data_dbh)

dbh <- ggplot(data_dbh, aes(x = CRxH, y = DBH)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(x = CRxH, y = community_fitted_DBH, color = sp, group = sp), show.legend = FALSE) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CRxH ",
       y = "DBH (cm)") +
  scale_x_log10() +
  scale_y_log10() +
  theme_minimal()

dbh
```

# Plot DBH = a(CR*H)^b model (6, 2) (Using gamma as community-level parameters)

```{r}
# Read the data
stan_data_lr_dbh1 <- tar_read(stan_data_lr_dbh1)
str(stan_data_lr_dbh1)

# Extract relevant data
log_y <- stan_data_lr_dbh1$log_y
log_x <- stan_data_lr_dbh1$log_x
sp <- as.factor(stan_data_lr_dbh1$jj)

str(log_x)
head(log_x)

# Read in the model summary
dbh1 <- tar_read(fit_lr1_summary_power_law_dbh1)
dbh1

# Filter rows where the variable is "beta"
beta <- dbh1 |>
  filter(grepl("beta", variable))

beta

log_a <- beta |> filter(str_detect(variable, "beta\\[1,\\d+\\]")) |> pull(mean)
b <- beta |> filter(str_detect(variable, "beta\\[2,\\d+\\]")) |> pull(mean)

print(length(log_a))
print(length(b))
print(length(sp))

# Compute fitted values for DBH using the power law formula
fitted_log_y_dbh <- sapply(1:length(log_a), function(i) {
  log_a[i] + b[i] * log_x[, 2]
})

fitted_dbh <- exp(fitted_log_y_dbh)

# Extract community-level parameters directly
community_log_a <- dbh1 |> filter(variable == "gamma[1,1]") |> pull(mean)
community_b <- dbh1 |> filter(variable == "gamma[2,1]") |> pull(mean)

community_fitted_log_y_dbh <- community_log_a + community_b * log_x[, 2]

community_fitted_DBH <- exp(community_fitted_log_y_dbh)

# Create a data frame for plotting
data_dbh <- data.frame(DBH = exp(log_y), CRxH = exp(log_x[, 2]), sp = sp)
data_dbh$community_fitted_DBH <- community_fitted_DBH

head(data_dbh)
summary(data_dbh)

# Add species-specific fitted values to the data frame
for (i in 1:length(log_a)) {
  data_dbh[[paste0("fitted_DBH_sp_", i)]] <- fitted_dbh[, i]
}

data_long <- melt(data_dbh, id.vars = c("DBH", "CRxH", "sp", "community_fitted_DBH"), measure.vars = grep("fitted_DBH_sp_", names(data_dbh), value = TRUE))

head(data_long)
summary(data_long)
str(data_long)

# Plotting
dbh_plot1 <- ggplot(data_dbh, aes(x = CRxH, y = DBH)) +
  geom_point(alpha = 0.5, color = "grey") +
  geom_line(data = data_long, aes(x = CRxH, y = value, group = variable), color = "#4682B4", alpha = 0.3) + # SteelBlue color for species fits
  geom_line(aes(y = community_fitted_DBH), color = "white", linewidth = 1.2) +
  labs(title = "DBH = a*(CR*H)^b",
       x = "CR*H",
       y = "DBH") +
  scale_x_log10(labels = label_number()) +
  scale_y_log10(labels = label_number()) +
  theme_minimal()

# Save the plot
ggsave(filename = "figs/dbh/CR*H_DBH1.png", plot = dbh_plot1, width = 12, height = 8, dpi = 300, bg = "white")

```