### 2025-03-26 docs/wd_ef_CIs.qmd

---
title: "Wood density's effect"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r, include=FALSE}
library(ggplot2)
library(targets)
library(tidyverse)
library(patchwork)
source("R/figs.R")
options(width = 150)
library(cowplot)

```

# H-DBH angiosperms

```{r}
tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)
tallo_reduced_nlr_df_ang_h

wd_h_ang_dat <- tallo_reduced_nlr_df_ang_h |>
  filter(division == "Angiosperm") |>
  dplyr::select(sp, dbh, h, cr, wd) |>
  group_by(sp) |>
  summarize(wd = mean(wd)) |>
  mutate(wd_s = scale(wd) |> as.numeric())
wd_h_ang_dat

light_h_ang_wd_s <- quantile(wd_h_ang_dat$wd_s, 0.1, na.rm = TRUE)
dense_h_ang_wd_s <- quantile(wd_h_ang_dat$wd_s, 0.9, na.rm = TRUE)
light_h_ang_wd_s
dense_h_ang_wd_s

fit_wd_ang_h_draws_weibull_wd <- tar_read(fit_wd_ang_h_draws_weibull_wd)
fit_wd_ang_h_draws_weibull_wd

gamma_params_h <- fit_wd_ang_h_draws_weibull_wd |>
  dplyr::select(starts_with("gamma_int"), starts_with("gamma_slope"))

gamma_params_h

median_a <- median(exp(gamma_params_h$`gamma_int[1]`), na.rm = TRUE)
median_b <- median(gamma_params_h$`gamma_int[2]`, na.rm = TRUE)

gamma_params_h <- gamma_params_h |>
  mutate(
    k_light = `gamma_int[3]` + `gamma_slope[3]` * light_h_ang_wd_s,
    k_dense = `gamma_int[3]` + `gamma_slope[3]` * dense_h_ang_wd_s
  )
gamma_params_h

dbh_seq <- seq(1, 150, length.out = 100)

pred_h <- gamma_params_h |>
  rowwise() |>
  mutate(
    h_light = list(tibble(
      dbh = dbh_seq,
      h_light = median_a * (1 - exp(-median_b * dbh^k_light))
    )),
    h_dense = list(tibble(
      dbh = dbh_seq,
      h_dense = median_a * (1 - exp(-median_b * dbh^k_dense))
    ))
  ) |>
  select(h_light, h_dense) |>
  unnest(cols = c(h_light, h_dense), names_sep = "_") |>
  rename(
    dbh = h_light_dbh,
    h_light = h_light_h_light,
    h_dense_dbh = h_dense_dbh,
    h_dense = h_dense_h_dense
  )

pred_h
summary(pred_h)
str(pred_h)

h_ang_light <- pred_h |>
  group_by(dbh) |>
  summarize(
    h_median_light = median(h_light),
    h_lower_light = quantile(h_light, 0.025),
    h_upper_light = quantile(h_light, 0.975),
    .groups = "drop"
  )

h_ang_light

h_ang_dense <- pred_h |>
  group_by(dbh) |>
  summarize(
    h_median_dense = median(h_dense),
    h_lower_dense = quantile(h_dense, 0.025),
    h_upper_dense = quantile(h_dense, 0.975),
    .groups = "drop"
  )

h_ang_dense

h_ang_dat <- left_join(
  h_ang_light,
  h_ang_dense,
  by = "dbh"
)
h_ang_dat

p1 <- ggplot(h_ang_dat, aes(x = dbh)) +
  geom_ribbon(aes(ymin = h_lower_light, ymax = h_upper_light, fill = "Light WD species"), alpha = 0.2) +
  geom_line(aes(y = h_median_light, color = "Light WD species")) +
  geom_ribbon(aes(ymin = h_lower_dense, ymax = h_upper_dense, fill = "Dense WD species"), alpha = 0.2) +
  geom_line(aes(y = h_median_dense, color = "Dense WD species")) +
  scale_color_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_fill_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(x = "DBH (cm)", y = "Height (m)") +
  my_theme()

p1

```

# CR-DBH angiosperms

```{r}
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr

wd_cr_ang_dat <- tallo_reduced_lr_df_ang_cr |>
  filter(division == "Angiosperm") |>
  dplyr::select(sp, wd) |>
  group_by(sp) |>
  summarize(wd = mean(wd)) |>
  mutate(wd_s = scale(wd) |> as.numeric())
wd_cr_ang_dat

light_cr_ang_wd_s <- quantile(wd_cr_ang_dat$wd_s, 0.1, na.rm = TRUE)
dense_cr_ang_wd_s <- quantile(wd_cr_ang_dat$wd_s, 0.9, na.rm = TRUE)
light_cr_ang_wd_s
dense_cr_ang_wd_s

fit_wd_ang_cr_draws_pl_wd <- tar_read(fit_wd_ang_cr_draws_pl_wd)
fit_wd_ang_cr_draws_pl_wd

gamma_params_cr_ang <- fit_wd_ang_cr_draws_pl_wd |>
  dplyr::select(starts_with("gamma"))

gamma_params_cr_ang

gamma_params_cr_ang <- gamma_params_cr_ang |>
  mutate(
    a_light = exp(`gamma[1,1]` + `gamma[1,2]` * light_cr_ang_wd_s),
    a_dense = exp(`gamma[1,1]` + `gamma[1,2]` * dense_cr_ang_wd_s),
    b_light = `gamma[2,1]` + `gamma[2,2]` * light_cr_ang_wd_s,
    b_dense = `gamma[2,1]` + `gamma[2,2]` * dense_cr_ang_wd_s
  )

gamma_params_cr_ang

dbh_cr_ang_seq <- seq(1, 150, length.out = 100)

pred_ang_cr <- gamma_params_cr_ang |>
  rowwise() |>
  mutate(
    cr_light = list(tibble(
      dbh = dbh_cr_ang_seq,
      cr_light = a_light * dbh^b_light
    )),
    cr_dense = list(tibble(
      dbh = dbh_cr_ang_seq,
      cr_dense = a_dense * dbh^b_dense
    ))
  ) |>
  select(cr_light, cr_dense) |>
  unnest(cols = c(cr_light, cr_dense), names_sep = "_") |>
  rename(
    dbh = cr_light_dbh,
    cr_light = cr_light_cr_light,
    cr_dense_dbh = cr_dense_dbh,
    cr_dense = cr_dense_cr_dense
  )

pred_ang_cr
summary(pred_ang_cr)
str(pred_ang_cr)

cr_ang_light <- pred_ang_cr |>
  group_by(dbh) |>
  summarize(
    cr_median_light = median(cr_light),
    cr_lower_light = quantile(cr_light, 0.025),
    cr_upper_light = quantile(cr_light, 0.975),
    .groups = "drop"
  )

cr_ang_light

cr_ang_dense <- pred_ang_cr |>
  group_by(dbh) |>
  summarize(
    cr_median_dense = median(cr_dense),
    cr_lower_dense = quantile(cr_dense, 0.025),
    cr_upper_dense = quantile(cr_dense, 0.975),
    .groups = "drop"
  )

cr_ang_dense

cr_ang_dat <- left_join(
  cr_ang_light,
  cr_ang_dense,
  by = "dbh"
)

cr_ang_dat

p2 <- ggplot(cr_ang_dat, aes(x = dbh)) +
  geom_ribbon(aes(ymin = cr_lower_light, ymax = cr_upper_light, fill = "Light WD species"), alpha = 0.2) +
  geom_line(aes(y =  cr_median_light, color = "Light WD species")) +
  geom_ribbon(aes(ymin = cr_lower_dense, ymax = cr_upper_dense, fill = "Dense WD species"), alpha = 0.2) +
  geom_line(aes(y = cr_median_dense, color = "Dense WD species")) +
  scale_color_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_fill_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(x = "DBH (cm)", y = "Crown Radius (m)") +
  my_theme()

p2
```

# CR-DBH gymnosperms

```{r}
tallo_reduced_nlr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)
tallo_reduced_nlr_df_gym_cr

wd_cr_gym_dat <- tallo_reduced_nlr_df_gym_cr |>
  filter(division == "Gymnosperm") |>
  dplyr::select(sp, dbh, h, cr, wd) |>
  group_by(sp) |>
  summarize(wd = mean(wd)) |>
  mutate(wd_s = scale(wd) |> as.numeric())
wd_cr_gym_dat

light_cr_gym_wd_s <- quantile(wd_cr_gym_dat$wd_s, 0.1, na.rm = TRUE)
dense_cr_gym_wd_s <- quantile(wd_cr_gym_dat$wd_s, 0.9, na.rm = TRUE)
light_cr_gym_wd_s
dense_cr_gym_wd_s

fit_wd_gym_cr_draws_gmm_wd <- tar_read(fit_wd_gym_cr_draws_gmm_wd)
fit_wd_gym_cr_draws_gmm_wd

gamma_params_cr_gym <- fit_wd_gym_cr_draws_gmm_wd |>
  dplyr::select(starts_with("gamma_int"), starts_with("gamma_slope"))

gamma_params_cr_gym

median_cr_gym_a <- median(exp(gamma_params_cr_gym$`gamma_int[1]`), na.rm = TRUE)
median_cr_gym_k <- median(gamma_params_cr_gym$`gamma_int[3]`, na.rm = TRUE)

gamma_params_cr_gym <- gamma_params_cr_gym |>
  mutate(
    b_light = `gamma_int[2]` + `gamma_slope[2]` * light_cr_gym_wd_s,
    b_dense = `gamma_int[2]` + `gamma_slope[2]` * dense_cr_gym_wd_s
  )
gamma_params_cr_gym

dbh_cr_gym_seq <- seq(1, 150, length.out = 100)

pred_gym_cr <- gamma_params_cr_gym |>
  rowwise() |>
  mutate(
    cr_light = list(tibble(
      dbh = dbh_cr_gym_seq,
      cr_light = median_cr_gym_a * dbh^b_light / (median_cr_gym_k + dbh^b_light)
    )),
    cr_dense = list(tibble(
      dbh = dbh_cr_gym_seq,
      cr_dense = median_cr_gym_a * dbh^b_dense / (median_cr_gym_k + dbh^b_dense)
    ))
  ) |>
  select(cr_light, cr_dense) |>
  unnest(cols = c(cr_light, cr_dense), names_sep = "_") |>
  rename(
    dbh = cr_light_dbh,
    cr_light = cr_light_cr_light,
    cr_dense_dbh = cr_dense_dbh,
    cr_dense = cr_dense_cr_dense
  )

pred_gym_cr
summary(pred_gym_cr)
str(pred_gym_cr)

cr_gym_light <- pred_gym_cr |>
  group_by(dbh) |>
  summarize(
    cr_median_light = median(cr_light),
    cr_lower_light = quantile(cr_light, 0.025),
    cr_upper_light = quantile(cr_light, 0.975),
    .groups = "drop"
  )

cr_gym_light

cr_gym_dense <- pred_gym_cr |>
  group_by(dbh) |>
  summarize(
    cr_median_dense = median(cr_dense),
    cr_lower_dense = quantile(cr_dense, 0.025),
    cr_upper_dense = quantile(cr_dense, 0.975),
    .groups = "drop"
  )

pred_gym_cr

cr_gym_dat <- left_join(
  cr_gym_light,
  cr_gym_dense,
  by = "dbh"
)

cr_gym_dat

p3 <- ggplot(cr_gym_dat, aes(x = dbh)) +
  geom_ribbon(aes(ymin = cr_lower_light, ymax = cr_upper_light, fill = "Light WD species"), alpha = 0.2) +
  geom_line(aes(y =  cr_median_light, color = "Light WD species")) +
  geom_ribbon(aes(ymin = cr_lower_dense, ymax = cr_upper_dense, fill = "Dense WD species"), alpha = 0.2) +
  geom_line(aes(y = cr_median_dense, color = "Dense WD species")) +
  scale_color_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_fill_manual(name = "Wood Type", values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(x = "DBH (cm)", y = "Crown Radius (m)") +
  my_theme()

p3
```

# Combine in 1 plot

```{r}
p <- p1 + p2 + p3 + plot_layout(nrow = 1, guides = "collect") +
      plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)"))) &
      theme(
      plot.tag = element_text(size = 8),
      axis.title = element_text(size = 9),
      text = element_text(size = 8),
      legend.title = element_blank(),
      plot.margin = margin(t = 2, r = 2, b = 2, l = 2, unit = "pt")
    )

p <- ggdraw() +
draw_plot(p, x = 0, y = 0, width = 1, height = 0.95) +  
draw_label("Angiosperm", x = 0.172, y = 0.94, fontface = 'bold', size = 10) +  
draw_label("Angiosperm", x = 0.429, y = 0.94, fontface = 'bold', size = 10) +  
draw_label("Gymnosperm", x = 0.689, y = 0.94, fontface = 'bold', size = 10)

```

# Writing function

```{r}
process_wd_data <- function(df, division, light_quantile = 0.1, dense_quantile = 0.9) {
  df_filtered <- df |>
    filter(division == division) |>
    dplyr::select(sp, dbh, h, cr, wd) |>
    group_by(sp) |>
    summarize(wd = mean(wd)) |>
    mutate(wd_s = scale(wd) |> as.numeric())

  light_wd_s <- quantile(df_filtered$wd_s, light_quantile, na.rm = TRUE)
  dense_wd_s <- quantile(df_filtered$wd_s, dense_quantile, na.rm = TRUE)

  return(list(df = df_filtered, light_wd_s = light_wd_s, dense_wd_s = dense_wd_s))
}

```

```{r}

wd_h_ang_dat <- process_wd_data(tallo_reduced_nlr_df_ang_h)
wd_h_ang_dat

wd_cr_ang_dat <- process_wd_data(tallo_reduced_lr_df_ang_cr)
wd_cr_ang_dat

wd_cr_gym_dat <- process_wd_data(tallo_reduced_nlr_df_gym_cr)
wd_cr_gym_dat

```