### 2025-03-26 docs/wd_ef_check.qmd

---
title: "Checking Wood density's effect plot"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r, include=FALSE}
library(ggplot2)
library(targets)
library(tidyverse)
library(patchwork)
source("R/figs.R")
options(width = 150)

```

# WD vs. CR-DBH

## Angiosperm

### No CIs
```{r}
sp_posterior_cr_ang_df_wd <- tar_read(sp_posterior_cr_ang_df_wd)
head(sp_posterior_cr_ang_df_wd)

tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)

ang_cr_post <- tallo_reduced_lr_df_ang_cr |>
  filter(division == "Angiosperm") |>
  left_join(sp_posterior_cr_ang_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, cr, wd, a, b) |>
  mutate(
    a = as.numeric(sub(" .*", "", a)),
    b = as.numeric(sub(" .*", "", b))
  )

ang_cr_post
summary(ang_cr_post)

wd_quantiles <- quantile(ang_cr_post$wd, probs = c(0.1, 0.9), na.rm = TRUE)

light_wood <- ang_cr_post |>
  filter(wd < wd_quantiles[1])

dense_wood <- ang_cr_post |>
  filter(wd > wd_quantiles[2])

dbh_seq <- seq(1, 150, length.out = 100)

mean_light_wood <- light_wood |>
  summarise(across(c(a, b), mean, na.rm = TRUE))

mean_dense_wood <- dense_wood |>
  summarise(across(c(a, b), mean, na.rm = TRUE))

predicted_cr_light <- mean_light_wood$a * dbh_seq^mean_light_wood$b
predicted_cr_dense <- mean_dense_wood$a * dbh_seq^mean_dense_wood$b

prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_cr = c(predicted_cr_light, predicted_cr_dense),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

ggplot(prediction_df_mean, aes(x = dbh, y = predicted_cr, color = wood_type)) +
  geom_line(alpha = 0.9) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Crown Radius (m)",
    color = "Wood Density"
  ) +
  my_theme()

```

### Adding CIs

```{r}
sp_posterior_cr_ang_df_wd <- tar_read(sp_posterior_cr_ang_df_wd)
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)

ang_cr_post <- tallo_reduced_lr_df_ang_cr |>
  filter(division == "Angiosperm") |>
  left_join(sp_posterior_cr_ang_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, cr, wd, a, b) |>
  mutate(
    a_median = as.numeric(str_extract(a, "^[^ ]+")),
    b_median = as.numeric(str_extract(b, "^[^ ]+")),

    a_lower = as.numeric(str_extract(a, "(?<=\\().*?(?=,)")),
    b_lower = as.numeric(str_extract(b, "(?<=\\().*?(?=,)")),

    a_upper = as.numeric(str_extract(a, "(?<=,).*?(?=\\))")),
    b_upper = as.numeric(str_extract(b, "(?<=,).*?(?=\\))"))
  ) |>
  dplyr::select(-a, - b)

ang_cr_post
summary(ang_cr_post)

light_wood <- ang_cr_post |>
  filter(wd < quantile(ang_cr_post$wd, probs = 0.1, na.rm = TRUE))

dense_wood <- ang_cr_post |>
  filter(wd > quantile(ang_cr_post$wd, probs = 0.9, na.rm = TRUE))

mean_light_wood <- light_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE)
  )

mean_dense_wood <- dense_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE)
  )

dbh_seq <- seq(1, 100, length.out = 100)

predicted_cr_light <- mean_light_wood$a_median * dbh_seq^mean_light_wood$b_median
predicted_cr_dense <- mean_dense_wood$a_median * dbh_seq^mean_dense_wood$b_median

# Calculate CI bounds for both wood types using lower and upper values of a and b
predicted_cr_light_lower <- mean_light_wood$a_lower * dbh_seq^mean_light_wood$b_lower
predicted_cr_light_upper <- mean_light_wood$a_upper * dbh_seq^mean_light_wood$b_upper

predicted_cr_dense_lower <- mean_dense_wood$a_lower * dbh_seq^mean_dense_wood$b_lower
predicted_cr_dense_upper <- mean_dense_wood$a_upper * dbh_seq^mean_dense_wood$b_upper

prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_cr = c(predicted_cr_light, predicted_cr_dense),
  predicted_cr_lower = c(predicted_cr_light_lower, predicted_cr_dense_lower),
  predicted_cr_upper = c(predicted_cr_light_upper, predicted_cr_dense_upper),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

head(prediction_df_mean)

ggplot(prediction_df_mean, aes(x = dbh, fill = wood_type)) +
  geom_ribbon(aes(ymin = predicted_cr_lower, ymax = predicted_cr_upper), alpha = 0.2) +
  geom_line(aes(y = predicted_cr, color = wood_type), alpha = 0.9) +
  scale_fill_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Crown Radius (m)",
    fill = "Wood Density",
    color = "Wood Density"
  ) +
  my_theme()

```

## Gymnosperms

### No CIs

```{r}
sp_posterior_cr_gym_df_wd <- tar_read(sp_posterior_cr_gym_df_wd)
head(sp_posterior_cr_gym_df_wd)

tallo_reduced_lr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

gym_cr_df <- tallo_reduced_lr_df_gym_cr |> filter(division == "Gymnosperm")
gym_cr_df

gym_cr_post <- gym_cr_df |>
  left_join(sp_posterior_cr_gym_df_wd, by = c("sp" = "sp")) |>
  dplyr::select(tree_id, sp, dbh, h, cr, wd, a, b, k) |>
  mutate(
    a = as.numeric(sub(" .*", "", a)),
    b = as.numeric(sub(" .*", "", b)),
    k = as.numeric(sub(" .*", "", k))
  )

gym_cr_post
summary(gym_cr_post)

sp_posterior_cr_gym_df_wd <- tar_read(sp_posterior_cr_gym_df_wd)
tallo_reduced_lr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

gym_cr_post <- tallo_reduced_lr_df_gym_cr |>
  filter(division == "Gymnosperm") |>
  left_join(sp_posterior_cr_gym_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, cr, wd, a, b, k) |>
  mutate(across(c(a, b, k), ~ as.numeric(sub(" .*", "", .))))

wd_quantiles <- quantile(gym_cr_post$wd, probs = c(0.1, 0.9), na.rm = TRUE)
light_wood <- gym_cr_post |> filter(wd < wd_quantiles[1])
dense_wood <- gym_cr_post |> filter(wd > wd_quantiles[2])

dbh_seq <- seq(1, 150, length.out = 100)

mean_light_wood <- light_wood |> summarise(across(c(a, b, k), mean, na.rm = TRUE))
mean_dense_wood <- dense_wood |> summarise(across(c(a, b, k), mean, na.rm = TRUE))

predicted_cr_light <- (mean_light_wood$a * dbh_seq^mean_light_wood$b) / (mean_light_wood$k + dbh_seq^mean_light_wood$b)
predicted_cr_dense <- (mean_dense_wood$a * dbh_seq^mean_dense_wood$b) / (mean_dense_wood$k + dbh_seq^mean_dense_wood$b)

prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_cr = c(predicted_cr_light, predicted_cr_dense),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

ggplot(prediction_df_mean, aes(x = dbh, y = predicted_cr, color = wood_type)) +
  geom_line(alpha = 0.9) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Crown Radius (m)",
    color = "Wood Density"
  ) +
  my_theme()

```

### Adding CIs

```{r}
sp_posterior_cr_gym_df_wd <- tar_read(sp_posterior_cr_gym_df_wd)
tallo_reduced_lr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

# tallo_reduced_nlr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

gym_cr_post <- tallo_reduced_lr_df_gym_cr |>
  filter(division == "Gymnosperm") |>
  left_join(sp_posterior_cr_gym_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, cr, wd, a, b, k) |>
  mutate(
    a_median = as.numeric(str_extract(a, "^[^ ]+")),
    b_median = as.numeric(str_extract(b, "^[^ ]+")),
    k_median = as.numeric(str_extract(k, "^[^ ]+")),

    a_lower = as.numeric(str_extract(a, "(?<=\\().*?(?=,)")),
    b_lower = as.numeric(str_extract(b, "(?<=\\().*?(?=,)")),
    k_lower = as.numeric(str_extract(k, "(?<=\\().*?(?=,)")),

    a_upper = as.numeric(str_extract(a, "(?<=,).*?(?=\\))")),
    b_upper = as.numeric(str_extract(b, "(?<=,).*?(?=\\))")),
    k_upper = as.numeric(str_extract(k, "(?<=,).*?(?=\\))"))
  ) |>
  select(-a, -b, -k)

gym_cr_post
wd_quantiles <- quantile(gym_cr_post$wd, probs = c(0.1, 0.9), na.rm = TRUE)
light_wood <- gym_cr_post |> filter(wd < wd_quantiles[1])
dense_wood <- gym_cr_post |> filter(wd > wd_quantiles[2])

mean_light_wood <- light_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    k_median = mean(k_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    k_lower = mean(k_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE),
    k_upper = mean(k_upper, na.rm = TRUE)
  )

mean_dense_wood <- dense_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    k_median = mean(k_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    k_lower = mean(k_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE),
    k_upper = mean(k_upper, na.rm = TRUE)
  )

dbh_seq <- seq(1, 150, length.out = 100)

predicted_cr_light <- (mean_light_wood$a_median * dbh_seq^mean_light_wood$b_median) / 
  (mean_light_wood$k_median + dbh_seq^mean_light_wood$b_median)

predicted_cr_dense <- (mean_dense_wood$a_median * dbh_seq^mean_dense_wood$b_median) / 
  (mean_dense_wood$k_median + dbh_seq^mean_dense_wood$b_median)

predicted_cr_light_lower <- (mean_light_wood$a_lower * dbh_seq^mean_light_wood$b_lower) / 
  (mean_light_wood$k_lower + dbh_seq^mean_light_wood$b_lower)

predicted_cr_light_upper <- (mean_light_wood$a_upper * dbh_seq^mean_light_wood$b_upper) / 
  (mean_light_wood$k_upper + dbh_seq^mean_light_wood$b_upper)

predicted_cr_dense_lower <- (mean_dense_wood$a_lower * dbh_seq^mean_dense_wood$b_lower) / 
  (mean_dense_wood$k_lower + dbh_seq^mean_dense_wood$b_lower)

predicted_cr_dense_upper <- (mean_dense_wood$a_upper * dbh_seq^mean_dense_wood$b_upper) / 
  (mean_dense_wood$k_upper + dbh_seq^mean_dense_wood$b_upper)

# Combine into a single data frame
prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_cr = c(predicted_cr_light, predicted_cr_dense),
  predicted_cr_lower = c(predicted_cr_light_lower, predicted_cr_dense_lower),
  predicted_cr_upper = c(predicted_cr_light_upper, predicted_cr_dense_upper),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

ggplot(prediction_df_mean, aes(x = dbh, fill = wood_type)) +
  geom_ribbon(aes(ymin = predicted_cr_lower, ymax = predicted_cr_upper), alpha = 0.2) +
  geom_line(aes(y = predicted_cr, color = wood_type), alpha = 0.9) +
  scale_fill_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Crown Radius (m)",
    fill = "Wood Density",
    color = "Wood Density"
  ) +
  my_theme()
```

# WD vs. H-DBH

## Angiosperm

### No CIs
```{r}
sp_posterior_h_ang_df_wd <- tar_read(sp_posterior_h_ang_df_wd)
# tallo_reduced_lr_df_h_ang <- tar_read(tallo_reduced_nlr_df_ang_h)

tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)

ang_h_post <- tallo_reduced_lr_df_h_ang |>
  filter(division == "Angiosperm") |>
  left_join(sp_posterior_h_ang_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, wd, a, b, k) |>
  mutate(across(c(a, b, k), ~ as.numeric(sub(" .*", "", .))))

wd_quantiles <- quantile(ang_h_post$wd, probs = c(0.25, 0.75), na.rm = TRUE)
light_wood <- filter(ang_h_post, wd < wd_quantiles[1])
dense_wood <- filter(ang_h_post, wd > wd_quantiles[2])

dbh_seq <- seq(1, 150, length.out = 100)
mean_light_wood <- summarise(light_wood, across(c(a, b, k), mean))
mean_dense_wood <- summarise(dense_wood, across(c(a, b, k), mean))

predicted_h_light <- mean_light_wood$a * (1 - exp(-mean_light_wood$b * dbh_seq^mean_light_wood$k))
predicted_h_dense <- mean_dense_wood$a * (1 - exp(-mean_dense_wood$b * dbh_seq^mean_dense_wood$k))

prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_h = c(predicted_h_light, predicted_h_dense),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

ggplot(prediction_df_mean, aes(x = dbh, y = predicted_h, color = wood_type)) +
  geom_line(alpha = 0.9) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Tree Height (m)",
    color = "Wood Density"
  ) +
  my_theme()

```

### Adding CIs

```{r}
# Load data
sp_posterior_h_ang_df_wd <- tar_read(sp_posterior_h_ang_df_wd)
tallo_reduced_lr_df_h_ang <- tar_read(tallo_reduced_nlr_df_ang_h)

# Process data and parse CIs
ang_h_post <- tallo_reduced_lr_df_h_ang |>
  filter(division == "Angiosperm") |>
  left_join(sp_posterior_h_ang_df_wd, by = "sp") |>
  select(tree_id, sp, dbh, h, wd, a, b, k) |>
  mutate(
    a_median = as.numeric(str_extract(a, "^[^ ]+")),
    b_median = as.numeric(str_extract(b, "^[^ ]+")),
    k_median = as.numeric(str_extract(k, "^[^ ]+")),

    a_lower = as.numeric(str_extract(a, "(?<=\\().*?(?=,)")),
    b_lower = as.numeric(str_extract(b, "(?<=\\().*?(?=,)")),
    k_lower = as.numeric(str_extract(k, "(?<=\\().*?(?=,)")),

    a_upper = as.numeric(str_extract(a, "(?<=,).*?(?=\\))")),
    b_upper = as.numeric(str_extract(b, "(?<=,).*?(?=\\))")),
    k_upper = as.numeric(str_extract(k, "(?<=,).*?(?=\\))"))
  ) |>
  select(-a, -b, -k)

# Quantiles for Wood Density
wd_quantiles <- quantile(ang_h_post$wd, probs = c(0.25, 0.75), na.rm = TRUE)
light_wood <- filter(ang_h_post, wd < wd_quantiles[1])
dense_wood <- filter(ang_h_post, wd > wd_quantiles[2])

# Summary statistics
mean_light_wood <- light_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    k_median = mean(k_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    k_lower = mean(k_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE),
    k_upper = mean(k_upper, na.rm = TRUE)
  )

mean_dense_wood <- dense_wood |>
  summarise(
    a_median = mean(a_median, na.rm = TRUE),
    b_median = mean(b_median, na.rm = TRUE),
    k_median = mean(k_median, na.rm = TRUE),
    a_lower = mean(a_lower, na.rm = TRUE),
    b_lower = mean(b_lower, na.rm = TRUE),
    k_lower = mean(k_lower, na.rm = TRUE),
    a_upper = mean(a_upper, na.rm = TRUE),
    b_upper = mean(b_upper, na.rm = TRUE),
    k_upper = mean(k_upper, na.rm = TRUE)
  )

# Generate predictions
dbh_seq <- seq(1, 150, length.out = 100)

# Median predictions
predicted_h_light <- mean_light_wood$a_median * (1 - exp(-mean_light_wood$b_median * dbh_seq^mean_light_wood$k_median))
predicted_h_dense <- mean_dense_wood$a_median * (1 - exp(-mean_dense_wood$b_median * dbh_seq^mean_dense_wood$k_median))

# Lower and upper bounds
predicted_h_light_lower <- mean_light_wood$a_lower * (1 - exp(-mean_light_wood$b_lower * dbh_seq^mean_light_wood$k_lower))
predicted_h_light_upper <- mean_light_wood$a_upper * (1 - exp(-mean_light_wood$b_upper * dbh_seq^mean_light_wood$k_upper))

predicted_h_dense_lower <- mean_dense_wood$a_lower * (1 - exp(-mean_dense_wood$b_lower * dbh_seq^mean_dense_wood$k_lower))
predicted_h_dense_upper <- mean_dense_wood$a_upper * (1 - exp(-mean_dense_wood$b_upper * dbh_seq^mean_dense_wood$k_upper))

# Combine into a single data frame
prediction_df_mean <- data.frame(
  dbh = rep(dbh_seq, 2),
  predicted_h = c(predicted_h_light, predicted_h_dense),
  predicted_h_lower = c(predicted_h_light_lower, predicted_h_dense_lower),
  predicted_h_upper = c(predicted_h_light_upper, predicted_h_dense_upper),
  wood_type = rep(c("Light WD species", "Dense WD species"), each = length(dbh_seq))
)

# Plot with CIs
ggplot(prediction_df_mean, aes(x = dbh, fill = wood_type)) +
  geom_ribbon(aes(ymin = predicted_h_lower, ymax = predicted_h_upper), alpha = 0.2) +
  geom_line(aes(y = predicted_h, color = wood_type), alpha = 0.9) +
  scale_fill_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  scale_color_manual(values = c("Light WD species" = "#f48d10", "Dense WD species" = "#7f590e")) +
  labs(
    x = "DBH (cm)",
    y = "Tree Height (m)",
    fill = "Wood Density",
    color = "Wood Density"
  ) +
  my_theme()
```