### 2025-03-26 docs/wd_para_cr_new.qmd

---
title: "Wood density vs. CR-DBH allometry parameters (Weibull)"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r, include=FALSE}
library(targets)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(patchwork)
library(posterior)
options(width = 150)
```

# Summary result

## Ang

```{r}

pl_cr_wd_ang <- tar_read(fit_wd_ang_cr_summary_pl_wd)
pl_cr_wd_ang

fit_wd_ang_cr_draws_pl_wd <- tar_read(fit_wd_ang_cr_draws_pl_wd)
fit_wd_ang_cr_draws_pl_wd

```

- Key finding: Wood density is signifcant effect the horizontal expansion of *Angiosperm* tree.

## Gym

```{r}
gmm_cr_wd_gym <- tar_read(fit_wd_gym_cr_summary_gmm_wd)
gmm_cr_wd_gym

gmm_cr_wd_gym <- gmm_cr_wd_gym |>
  filter(str_detect(variable, "gamma"))

gmm_cr_wd_gym

gmm_cr_wd_gym <- gmm_cr_wd_gym |>
  filter(str_detect(variable, "gamma_int") | str_detect(variable, "gamma_slope"))
gmm_cr_wd_gym

```

- Key finding: Wood density significantly affects parameter b, but not a and k of CR-DBH allometry of *Gymnosperm*.

# Plotting

## Data preparation

### Ang

```{r}
tallo_wd_df_cr <- tar_read(tallo_wd_df_cr)
tallo_wd_df_cr

ang_cr_df <- tallo_wd_df_cr |> filter(division == "Angiosperm")
ang_cr_df

wd_df_cr_ang <- ang_cr_df |>
  group_by(sp) |>
  summarize(wd = mean(wd), .groups = "drop") |>
  mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

wd_df_cr_ang

```

- 762 species

```{r}
pl_cr_wd_ang_draws <- tar_read(fit_wd_ang_cr_draws_pl_wd)
pl_cr_wd_ang_draws

gamma_ang_cr <- pl_cr_wd_ang_draws |>
  select(starts_with("gamma"))

gamma_ang_cr

beta_ang_cr <- pl_cr_wd_ang_draws |>
  select(starts_with("beta["))

beta_ang_cr

beta_ang_cr_long <- beta_ang_cr |>
  pivot_longer(
    cols = everything(),
    names_to = c("parameter", "species"),
    names_pattern = "beta\\[(\\d+),(\\d+)\\]",
    names_transform = list(parameter = as.integer, species = as.integer),
    values_to = "value"
  ) |>
  mutate(parameter = case_when(
    parameter == 1 ~ "log_a",  # Assign 'log_a' for [1,]
    parameter == 2 ~ "b"       # Assign 'b' for [2,]
  ),
  value = log(value))

beta_ang_cr_long

plot_ang_cr_data <- beta_ang_cr_long |>
  left_join(wd_df_cr_ang, by = c("species" = "sp_id"))

plot_ang_cr_data

```

### Gym

#### Species ID

```{r}

gym_cr_df <- tallo_wd_df_cr |> filter(division == "Gymnosperm")
gym_cr_df

wd_df_cr_gym <- gym_cr_df |>
  group_by(sp) |>
  summarize(wd = mean(wd), .groups = "drop") |>
  mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

wd_df_cr_gym

wd_seq_cr_gym <- seq(min(wd_df_cr_gym$wd_s), max(wd_df_cr_gym$wd_s), length.out = 100)
```

#### Extracting parameters

```{r}
gmm_cr_wd_gym_draws <- tar_read(fit_wd_gym_cr_draws_gmm_wd)
gmm_cr_wd_gym_draws

fit_wd_gym_cr_data <- tar_read(fit_wd_gym_cr_data)

gamma_gym_cr <- gmm_cr_wd_gym_draws |>
  select(starts_with("gamma"))

gamma_gym_cr

beta_gym_cr <- gmm_cr_wd_gym_draws |>
  select(starts_with("beta["))

beta_gym_cr

beta_gym_cr_long <- beta_gym_cr |>
  pivot_longer(
    cols = everything(),
    names_to = c("species", "parameter"),
    names_pattern = "beta\\[(\\d+),(\\d+)\\]",
    names_transform = list(species = as.integer, parameter = as.integer),
    values_to = "value"
  ) |>
  mutate(parameter = case_when(
    parameter == 1 ~ "log_a",
    parameter == 2 ~ "b",
    parameter == 3 ~ "k"
  ),
  value = log(value))

beta_gym_cr_long

plot_gym_cr_data <- beta_gym_cr_long |>
  left_join(wd_df_cr_gym, by = c("species" = "sp_id"))

plot_gym_cr_data
```

# Plotting function

```{r}
tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr

fit_wd_ang_cr_draws_pl_wd <- tar_read(fit_wd_ang_cr_draws_pl_wd)
fit_wd_ang_cr_draws_pl_wd

fit_wd_ang_cr_data <- tar_read(fit_wd_ang_cr_data)
str(fit_wd_ang_cr_data)

stan_data_lr_ang_cr <- tar_read(stan_data_lr_ang_cr)
str(stan_data_lr_ang_cr)

# same as fit_wd_ang_cr_data

tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr
str(tallo_reduced_lr_df_ang_cr)

ang_cr_df <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")

ang_cr_df

wd_df_ang <- ang_cr_df |>
    group_by(sp) |>
    summarize(wd = mean(wd, na.rm = TRUE)) |>
    mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

wd_seq_ang <- seq(min(wd_df_ang$wd_s), max(wd_df_ang$wd_s), length.out = 100)

wd_seq_df_ang <- data.frame(
    wd_s = wd_seq_ang,
    wd = approx(wd_df_ang$wd_s, wd_df_ang$wd, xout = wd_seq_ang)$y
  )

```

```{r}
fit_wd_ang_cr_draws_pl_wd <- tar_read(fit_wd_ang_cr_draws_pl_wd)
fit_wd_ang_cr_draws_pl_wd

gamma_ang_cr <- fit_wd_ang_cr_draws_pl_wd |>
  select(starts_with("gamma"))

gamma_ang_cr

beta_ang_cr <- fit_wd_ang_cr_draws_pl_wd |>
  select(starts_with("beta["))

beta_ang_cr

beta_ang_cr_long <- beta_ang_cr |>
  pivot_longer(
    cols = everything(),  # Pivot all columns
    names_to = c("parameter", "species"),  # Split column names into 'parameter' and 'species'
    names_pattern = "beta\\[(\\d+),(\\d+)\\]",  # Extract the two parts from column names like beta[1,1]
    names_transform = list(parameter = as.integer, species = as.integer),  # Convert to integers
    values_to = "value"  # Store values in 'value'
  ) |>
  mutate(
    parameter = case_when(
      parameter == 1 ~ "log_a",  # Map '1' to 'log_a'
      parameter == 2 ~ "b"       # Map '2' to 'b'
    )
  )

beta_ang_cr_long

plot_ang_cr_data <- beta_ang_cr_long |>
  left_join(wd_df_ang, by = c("species" = "sp_id"))

plot_ang_cr_data

summary_stats <- plot_ang_cr_data |>
  group_by(species, parameter, wd, wd_s) |>
  summarize(
    median = median(value),
    lower = quantile(value, 0.025),
    upper = quantile(value, 0.975),
    .groups = "drop"
  )

# Join summary_stats with wd_df_ang to get the `wd` values
# summary_stats <- summary_stats |>
#   left_join(wd_df_ang, by = c("wd_s"))

summary_stats

gamma_log_a <- gamma_ang_cr |>
  select(`gamma[1,1]`, `gamma[1,2]`)
gamma_b <- gamma_ang_cr |>
  select(`gamma[1,2]`, `gamma[2,2]`)

lines_data <- do.call(rbind, lapply(1:nrow(gamma_log_a), function(i) {
  data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    log_a = gamma_log_a$`gamma[1,1]`[i] + gamma_log_a$`gamma[1,2]`[i] * wd_seq_df_ang$wd_s,
    draw = i
  )
}))

median_intercept <- median(gamma_log_a$`gamma[1,1]`)
median_slope <- median(gamma_log_a$`gamma[1,2]`)

# Update `lines_data` to compute `a` from `log_a`
lines_data <- do.call(rbind, lapply(1:nrow(gamma_log_a), function(i) {
  data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    a = exp(gamma_log_a$`gamma[1,1]`[i] + gamma_log_a$`gamma[1,2]`[i] * wd_seq_df_ang$wd_s),  # Now using `a` instead of `log_a`
    draw = i
  )
}))

fit_line_a <- data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    a = exp(median_intercept + median_slope * wd_seq_df_ang$wd_s)
)

# Updated plot code
ggplot() +
  geom_line(data = lines_data, aes(x = wd, y = a, group = draw), alpha = 0.1, color = "blue") +  # Pile of lines using `a` for y
  geom_line(data = fit_line_a, aes(x = wd, y = a), color = "red", size = 1) +  # Median fit line using `a`
  geom_point(data = summary_stats |> filter(parameter == "log_a"), aes(x = wd, y = exp(median)), color = "green") +  # Species-specific points using `a`
  geom_errorbar(data = summary_stats |> filter(parameter == "log_a"), aes(x = wd, ymin = exp(lower), ymax = exp(upper)), width = 0.02, color = "green") +  # 95% CIs using `a`
  theme_minimal() +
  labs(
    x = "WD",  # Label for x-axis
    y = "a"  # Label for y-axis (now using `a`)
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )

```

### for b

```{r}
# Select gamma terms for `b` from gamma_ang_cr
gamma_b <- gamma_ang_cr |>
  select(`gamma[2,1]`, `gamma[2,2]`)

# Prepare data for lines corresponding to `b`
lines_data_b <- do.call(rbind, lapply(1:nrow(gamma_b), function(i) {
  data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    b = gamma_b$`gamma[2,1]`[i] + gamma_b$`gamma[2,2]`[i] * wd_seq_df_ang$wd_s,
    draw = i
  )
}))

# Median fit line for `b`
median_intercept_b <- median(gamma_b$`gamma[2,1]`)
median_slope_b <- median(gamma_b$`gamma[2,2]`)

fit_line_b <- data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    b = median_intercept_b + median_slope_b * wd_seq_df_ang$wd_s
)

# Updated plot for `b`
ggplot() +
  geom_line(data = lines_data_b, aes(x = wd, y = b, group = draw), alpha = 0.1, color = "blue") +  # Pile of lines using `b` for y
  geom_line(data = fit_line_b, aes(x = wd, y = b), color = "red", size = 1) +  # Median fit line using `b`
  geom_point(data = summary_stats |> filter(parameter == "b"), aes(x = wd, y = median), color = "green") +  # Species-specific points using `b`
  geom_errorbar(data = summary_stats |> filter(parameter == "b"), aes(x = wd, ymin = lower, ymax = upper), width = 0.02, color = "green") +  # 95% CIs using `b`
  theme_minimal() +
  labs(
    x = "WD",  # Label for x-axis
    y = "b"  # Label for y-axis (now using `b`)
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  )
```

## The scale of points is weird? WHYYY?

```{r}
fit_wd_ang_cr_data <- tar_read(fit_wd_ang_cr_data)

fit_wd_ang_cr_draws_pl_wd <- tar_read(fit_wd_ang_cr_draws_pl_wd)

tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr
str(tallo_reduced_lr_df_ang_cr)

# Filter Angiosperms
ang_cr_df <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")

# Group by species and calculate mean wood density for Angiosperms
wd_df_ang <- ang_cr_df |>
  group_by(sp) |>
  summarize(wd = mean(wd, na.rm = TRUE)) |>
  mutate(wd_s = scale(wd) |> as.numeric(), sp_id = row_number())

# Create the sequence for wood density for Angiosperms
wd_seq_ang <- seq(min(wd_df_ang$wd_s), max(wd_df_ang$wd_s), length.out = 100)

wd_seq_ang

wd_seq_df_ang <- data.frame(
  wd_s = wd_seq_ang,
  wd = approx(wd_df_ang$wd_s, wd_df_ang$wd, xout = wd_seq_ang)$y
)

# Prepare function for plotting
prepare_and_plot_parameter <- function(fit_data, fit_draws, wd_df, param_name, color_point, color_ribbon, line_color, show_ribbon = TRUE) {
  # Extract gamma parameters for the selected parameter
  gamma_params <- fit_draws |>
    dplyr::select(starts_with("gamma"))

  # Extract beta (species-specific) parameters
  beta_params <- fit_draws |>
    dplyr::select(starts_with("beta[")) |>
    pivot_longer(
      cols = everything(),
      names_to = c("parameter", "species"),
      names_pattern = "beta\\[(\\d+),(\\d+)\\]",
      names_transform = list(parameter = as.integer, species = as.integer),
      values_to = "value"
    ) |>
    mutate(parameter = case_when(
      parameter == 1 ~ "log_a",
      parameter == 2 ~ "b"
    ))

  # Combine species-specific beta with wd_df
  plot_data <- beta_params |>
    left_join(wd_df, by = c("species" = "sp_id"))

  # Summarize the posterior distributions for plotting
  summary_stats <- plot_data |>
    group_by(species, parameter, wd, wd_s) |>
    summarize(
      median = median(value, na.rm = TRUE),
      lower = quantile(value, 0.025, na.rm = TRUE),
      upper = quantile(value, 0.975, na.rm = TRUE),
      .groups = "drop"
    )

  # Prepare data for lines using gamma parameters
  gamma_param <- gamma_params |>
    dplyr::select(paste0("gamma[1,", which(c("log_a", "b") == param_name), "]"), 
            paste0("gamma[2,", which(c("log_a", "b") == param_name), "]"))

  lines_data <- do.call(rbind, lapply(1:nrow(gamma_param), function(i) {
    data.frame(
      wd_s = wd_seq_df_ang$wd_s,
      wd = wd_seq_df_ang$wd,
      value = if (param_name == "log_a") {
        exp(gamma_param[[1]][i] + gamma_param[[2]][i] * wd_seq_df_ang$wd_s)
      } else {
        gamma_param[[1]][i] + gamma_param[[2]][i] * wd_seq_df_ang$wd_s
      },
      draw = i
    )
  }))

  # Calculate the fit line and 95% CIs
  median_intercept <- median(gamma_param[[1]], na.rm = TRUE)
  median_slope <- median(gamma_param[[2]], na.rm = TRUE)

  fit_line <- data.frame(
    wd_s = wd_seq_df_ang$wd_s,
    wd = wd_seq_df_ang$wd,
    value = if (param_name == "log_a") {
      exp(median_intercept + median_slope * wd_seq_df_ang$wd_s)
    } else {
      median_intercept + median_slope * wd_seq_df_ang$wd_s
    }
  )

  ci_ribbon <- lines_data |>
    group_by(wd) |>
    summarize(
      lower = quantile(value, 0.025, na.rm = TRUE),
      upper = quantile(value, 0.975, na.rm = TRUE),
      .groups = "drop"
    )

  fit_with_ci <- left_join(fit_line, ci_ribbon, by = "wd")

  if (param_name == "log_a") {
    summary_stats <- summary_stats |>
      mutate(median = exp(median), lower = exp(lower), upper = exp(upper))
  }

  # Create plot with the selected colors
  plot <- ggplot() +
    geom_errorbar(data = summary_stats |> filter(parameter == param_name),
                  aes(x = wd, ymin = lower, ymax = upper), 
                  width = 0, linewidth = 0.2, color = color_ribbon) +
    geom_point(data = summary_stats |> filter(parameter == param_name), 
                aes(x = wd, y = median), color = color_point, size = 0.5) +
    geom_line(data = fit_line, aes(x = wd, y = value), 
              color = line_color, linewidth = 0.6) +
    theme_minimal() +
    labs(x = "Wood Density (g cm⁻³)", y = ifelse(param_name == "log_a", "Amplitude (a)", "Scale (b)")) +
    theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size = 9), axis.title.y = element_text(size = 9))

  if (show_ribbon) {
    plot <- plot + geom_ribbon(data = fit_with_ci, aes(x = wd, ymin = lower, ymax = upper), alpha = 0.5, fill = color_ribbon)
  }

  return(plot)
}

```

```{r}
# Generate plots using the custom colors for log_a and b
p1 <- prepare_and_plot_parameter(fit_wd_ang_cr_data, fit_wd_ang_cr_draws_pl_wd, wd_df_ang, "log_a", "#9e9ac8", "#cbc9e2", "#6a51a3", show_ribbon = TRUE)
p2 <- prepare_and_plot_parameter(fit_wd_ang_cr_data, fit_wd_ang_cr_draws_pl_wd, wd_df_ang, "b", "#74c476", "#bae4b3", "#238b45", show_ribbon = TRUE)

# Combine the plots into one layout
p <- (p1 + labs(tag = "(a)", x = NULL) + theme(axis.text.x = element_blank()) |
      p2 + labs(tag = "(b)", x = expression("Wood Density (g cm"^"-3"~")"))) +
  plot_layout(ncol = 1, nrow = 2) +
  plot_annotation(tag_levels = list(c("(a)", "(b)"))) &
  theme(
    plot.tag = element_text(face = "bold", size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9),
    plot.margin = margin(5, 5, 5, 5)
  )

# Add a label for Angiosperm
p <- ggdraw() +
  draw_plot(p, x = 0, y = 0, width = 1, height = 0.9) +  
  draw_label("Angiosperm", x = 0.75, y = 0.91, fontface = 'bold', size = 10)

```