### 2025-03-26 docs/wd_para_new.qmd

---
title: "WD's effect"
author: "Nguyen Thi Duyen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)

```

```{r, include=FALSE}
library(ggplot2)
library(targets)
library(tidyverse)
library(patchwork)
options(width = 150)
```

# WD vs. CR-DBH

## Angiosperm

### All species

```{r}
fit_wd_ang_cr_summary_pl_wd <- tar_read(fit_wd_ang_cr_summary_pl_wd)
fit_wd_ang_cr_summary_pl_wd
beta_ang_cr_pl  <- fit_wd_ang_cr_summary_pl_wd |> filter(grepl("beta", variable))
beta_ang_cr_pl

tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr

ang_cr_df <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")
ang_cr_df
unique_species_cr_df <- unique(ang_cr_df$sp)
sp_id_mapping <- ang_cr_df |>
  select(sp) |>
  distinct() |>
  mutate(sp_id = row_number())

angiosperm_parameters <- beta_ang_cr_pl |>
  filter(grepl("beta", variable)) |>
  mutate(param_type = case_when(
    grepl("1,", variable) ~ "log_a",
    grepl("2,", variable) ~ "b"
  ))

angiosperm_parameters <- angiosperm_parameters |>
  mutate(sp_id = rep(1:(nrow(angiosperm_parameters) / 2), each = 2))

angiosperm_params_wide <- angiosperm_parameters |>
  pivot_wider(
    names_from = param_type,
    values_from = mean,
    id_cols = sp_id
  )

angiosperm_params_wide

angiosperm_params_with_sp <- angiosperm_params_wide |>
  left_join(sp_id_mapping, by = "sp_id")

angiosperm_params_with_sp

angiosperm_data_full <- ang_cr_df |>
  left_join(angiosperm_params_with_sp, by = "sp") |>
  select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b, )

angiosperm_data_full

species_summary <- angiosperm_data_full |>
  group_by(sp) |>
  summarize(
    avg_dbh = mean(dbh, na.rm = TRUE),
    avr_cr = mean(cr, na.rm = TRUE),
    avg_log_a = mean(log_a, na.rm = TRUE),
    avg_b = mean(b, na.rm = TRUE),
    avg_wd = mean(wd, na.rm = TRUE)
  )

species_summary
summary(species_summary)

dbh_seq <- seq(min(species_summary$avg_dbh), max(species_summary$avg_dbh), length.out = 100)

prediction_df <- data.frame()

for(i in 1:nrow(species_summary)) {
  sp <- species_summary$sp[i]
  avg_log_a <- species_summary$avg_log_a[i]
  avg_b <- species_summary$avg_b[i]
  avg_wd <- species_summary$avg_wd[i]

  predicted_cr <- exp(avg_log_a) * (dbh_seq ^ avg_b)

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
  prediction_df <- rbind(prediction_df, temp_df)
}

# ggplot(prediction_df, aes(x = dbh, y = predicted_cr, group = sp, color = avg_wd)) +
#   geom_line(alpha = 0.7) +
#   scale_color_viridis_c() +
#   labs(x = "DBH (cm)", y = "Crown Radius (m)") +
#   theme_minimal() +
#   theme(legend.title = element_text(size = 12),
#         plot.title = element_text(size = 14)) +
#   guides(color = guide_colorbar(title = "Wood Density"))
```

### Light and Dense

```{r}
# Calculate the 25% and 75% quantiles of WD
wd_quantiles <- quantile(species_summary$avg_wd, probs = c(0.25, 0.75))

# Filter data for low and high WD species
low_wd_species <- species_summary |> filter(abs(avg_wd - wd_quantiles[1]) == min(abs(avg_wd - wd_quantiles[1])))
high_wd_species <- species_summary |> filter(abs(avg_wd - wd_quantiles[2]) == min(abs(avg_wd - wd_quantiles[2])))

# Combine low and high WD data
wd_filtered <- bind_rows(low_wd_species, high_wd_species)

# Create DBH sequence
dbh_seq <- seq(min(species_summary$avg_dbh), max(species_summary$avg_dbh), length.out = 100)

# Generate predictions for low and high WD
prediction_df <- data.frame()

for (i in 1:nrow(wd_filtered)) {
  sp <- wd_filtered$sp[i]
  avg_log_a <- wd_filtered$avg_log_a[i]
  avg_b <- wd_filtered$avg_b[i]
  avg_wd <- wd_filtered$avg_wd[i]

  predicted_cr <- exp(avg_log_a) * (dbh_seq ^ avg_b)

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
  prediction_df <- rbind(prediction_df, temp_df)
}

ggplot(prediction_df, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
  geom_line(alpha = 0.7) +
  scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
  labs(x = "DBH (cm)", y = "Crown Radius (m)", color = "WD Quantiles") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 14)
  )

p1 <- ggplot(prediction_df, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
  geom_line(alpha = 0.7) +
  scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
  labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +  # Removed units from labels
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 14),
    axis.text = element_blank()  # Remove axis tick labels
  )

p1

```

## Gymnosperm

### All species

```{r}
fit_wd_gym_cr_summary_gmm_wd <- tar_read(fit_wd_gym_cr_summary_gmm_wd)
fit_wd_gym_cr_summary_gmm_wd
beta_gym_cr_gmm  <- fit_wd_gym_cr_summary_gmm_wd |> filter(grepl("beta", variable))
beta_gym_cr_gmm

tallo_reduced_nlr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

gym_cr_df <- tallo_reduced_nlr_df_gym_cr |> filter(division == "Gymnosperm")
gym_cr_df

unique_species_cr_df_gym <- unique(gym_cr_df$sp)
sp_id_mapping_gym <- gym_cr_df |>
  select(sp) |>
  distinct() |>
  mutate(sp_id = row_number())

gymnosperm_parameters <- beta_gym_cr_gmm |>
  filter(grepl("beta", variable)) |>
  mutate(param_type = case_when(
    grepl("\\[.*,1\\]", variable) ~ "log_a",
    grepl("\\[.*,2\\]", variable) ~ "b",
    grepl("\\[.*,3\\]", variable) ~ "k"
  ))

gymnosperm_parameters <- gymnosperm_parameters |>
  mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))

gymnosperm_params_wide <- gymnosperm_parameters |>
  select(sp_id, param_type, mean) |>
  pivot_wider(names_from = param_type, values_from = mean)

gymnosperm_params_with_sp <- gymnosperm_params_wide |>
  left_join(sp_id_mapping_gym, by = "sp_id")

gymnosperm_data_full <- gym_cr_df |>
  left_join(gymnosperm_params_with_sp, by = "sp") |>
  select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b, k)

species_summary_gym <- gymnosperm_data_full |>
  group_by(sp) |>
  summarize(
    avg_dbh = mean(dbh, na.rm = TRUE),
    avr_cr = mean(cr, na.rm = TRUE),
    avg_log_a = mean(log_a, na.rm = TRUE),
    avg_b = mean(b, na.rm = TRUE),
    avg_k = mean(k, na.rm = TRUE),
    avg_wd = mean(wd, na.rm = TRUE)
  )

species_summary_gym

summary(species_summary_gym)

dbh_seq <- seq(min(species_summary_gym$avg_dbh), max(species_summary_gym$avg_dbh), length.out = 100)

prediction_df_gym <- data.frame()

for (i in 1:nrow(species_summary_gym)) {
  sp <- species_summary_gym$sp[i]
  avg_log_a <- species_summary_gym$avg_log_a[i]
  avg_b <- species_summary_gym$avg_b[i]
  avg_k <- species_summary_gym$avg_k[i]
  avg_wd <- species_summary_gym$avg_wd[i]

  predicted_cr <- (exp(avg_log_a) * dbh_seq^avg_b) / (avg_k + dbh_seq^avg_b)

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
  prediction_df_gym <- rbind(prediction_df_gym, temp_df)
}

ggplot(prediction_df_gym, aes(x = dbh, y = predicted_cr, group = sp, color = avg_wd)) +
  geom_line(alpha = 0.7) +
  scale_color_viridis_c() +
  labs(x = "DBH (cm)", y = "Crown Radius (CR)") +
  theme_minimal() +
  theme(legend.title = element_text(size = 12),
        plot.title = element_text(size = 14)) +
  guides(color = guide_colorbar(title = "Wood Density"))

```

### Light and Dense

```{r}
# Calculate the 25% and 75% quantiles of WD for Gymnosperms
wd_quantiles_gym <- quantile(species_summary_gym$avg_wd, probs = c(0.25, 0.75))

# Filter data for low and high WD species
low_wd_species_gym <- species_summary_gym |> filter(abs(avg_wd - wd_quantiles_gym[1]) == min(abs(avg_wd - wd_quantiles_gym[1])))
high_wd_species_gym <- species_summary_gym |> filter(abs(avg_wd - wd_quantiles_gym[2]) == min(abs(avg_wd - wd_quantiles_gym[2])))

# Combine low and high WD data
wd_filtered_gym <- bind_rows(low_wd_species_gym, high_wd_species_gym)

# Create predictions for Gymnosperms based on selected species
prediction_df_gym <- data.frame()

for (i in 1:nrow(wd_filtered_gym)) {
  sp <- wd_filtered_gym$sp[i]
  avg_log_a <- wd_filtered_gym$avg_log_a[i]
  avg_b <- wd_filtered_gym$avg_b[i]
  avg_k <- wd_filtered_gym$avg_k[i]
  avg_wd <- wd_filtered_gym$avg_wd[i]

  predicted_cr <- (exp(avg_log_a) * dbh_seq^avg_b) / (avg_k + dbh_seq^avg_b)

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
  prediction_df_gym <- rbind(prediction_df_gym, temp_df)
}

# ggplot(prediction_df_gym, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
#   geom_line(alpha = 0.7) +
#   scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
#   labs(x = "DBH (cm)", y = "Crown Radius (CR)", color = "WD Quantiles") +
#   theme_minimal() +
#   theme(
#     legend.title = element_text(size = 12),
#     plot.title = element_text(size = 14)
#   )

p2 <- ggplot(prediction_df_gym, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
  geom_line(alpha = 0.7) +
  scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
  labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +  # Removed units from labels
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 14),
    axis.text = element_blank()  # Remove axis tick labels
  )

p2
```

# WD vs. H-DBH

## Angiosperm

### All species
```{r}
fit_wd_ang_h_summary_weibull_wd <- tar_read(fit_wd_ang_h_summary_weibull_wd)
fit_wd_ang_h_summary_weibull_wd

beta_ang_h_wb  <- fit_wd_ang_h_summary_weibull_wd |> filter(grepl("beta", variable))
beta_ang_h_wb

tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)
tallo_reduced_nlr_df_ang_h

ang_h_df <- tallo_reduced_nlr_df_ang_h |> filter(division == "Angiosperm")
summary(ang_h_df)

unique_species_h_df_ang <- unique(ang_h_df$sp)
sp_id_mapping_ang <- ang_h_df |>
  select(sp) |>
  distinct() |>
  mutate(sp_id = row_number())

beta_ang_h_weibull <- fit_wd_ang_h_summary_weibull_wd |>
  filter(grepl("beta", variable)) |>
  mutate(param_type = case_when(
    grepl("\\[.*,1\\]", variable) ~ "log_a",
    grepl("\\[.*,2\\]", variable) ~ "b",
    grepl("\\[.*,3\\]", variable) ~ "k"
  ))

beta_ang_h_weibull <- beta_ang_h_weibull |>
  mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))

angiosperm_params_wide <- beta_ang_h_weibull |>
  select(sp_id, param_type, mean) |>
  pivot_wider(names_from = param_type, values_from = mean)

angiosperm_params_with_sp <- angiosperm_params_wide |>
  left_join(sp_id_mapping_ang, by = "sp_id")

angiosperm_data_full <- ang_h_df |>
  left_join(angiosperm_params_with_sp, by = "sp") |>
  select(tree_id, sp, sp_id, dbh, h, wd, log_a, b, k)

species_summary_ang <- angiosperm_data_full |>
  group_by(sp) |>
  summarize(
    avg_dbh = mean(dbh, na.rm = TRUE),
    avg_h = mean(h, na.rm = TRUE),
    avg_log_a = mean(log_a, na.rm = TRUE),
    avg_b = mean(b, na.rm = TRUE),
    avg_k = mean(k, na.rm = TRUE),
    avg_wd = mean(wd, na.rm = TRUE)
  )
species_summary_ang

summary(species_summary_ang)

dbh_seq <- seq(min(species_summary_ang$avg_dbh), max(species_summary_ang$avg_dbh), length.out = 100)

prediction_df_ang <- data.frame()

for (i in 1:nrow(species_summary_ang)) {
  sp <- species_summary_ang$sp[i]
  avg_log_a <- species_summary_ang$avg_log_a[i]
  avg_b <- species_summary_ang$avg_b[i]
  avg_k <- species_summary_ang$avg_k[i]
  avg_wd <- species_summary_ang$avg_wd[i]

  predicted_h <- exp(avg_log_a) * (1 - exp(-avg_b * dbh_seq^avg_k))

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_h = predicted_h, avg_wd = avg_wd)
  prediction_df_ang <- rbind(prediction_df_ang, temp_df)
}

ggplot(prediction_df_ang, aes(x = dbh, y = predicted_h, group = sp, color = avg_wd)) +
  geom_line(alpha = 0.7) +
  scale_color_viridis_c() +
  labs(x = "DBH (cm)", y = "Height (H)") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 14)
  ) +
  guides(color = guide_colorbar(title = "Wood Density"))
```

### Light and Dense

```{r}
# Using Weibull model, calculate the 25% and 75% quantiles for WD in Angiosperms
wd_quantiles_ang <- quantile(species_summary_ang$avg_wd, probs = c(0.25, 0.75))

# Filter data for low and high WD species
low_wd_species_ang <- species_summary_ang |> filter(abs(avg_wd - wd_quantiles_ang[1]) == min(abs(avg_wd - wd_quantiles_ang[1])))
high_wd_species_ang <- species_summary_ang |> filter(abs(avg_wd - wd_quantiles_ang[2]) == min(abs(avg_wd - wd_quantiles_ang[2])))

# Combine low and high WD data
wd_filtered_ang <- bind_rows(low_wd_species_ang, high_wd_species_ang)

# Generate height predictions for low and high WD
prediction_df_ang <- data.frame()

for (i in 1:nrow(wd_filtered_ang)) {
  sp <- wd_filtered_ang$sp[i]
  avg_log_a <- wd_filtered_ang$avg_log_a[i]
  avg_b <- wd_filtered_ang$avg_b[i]
  avg_k <- wd_filtered_ang$avg_k[i]
  avg_wd <- wd_filtered_ang$avg_wd[i]

  predicted_h <- exp(avg_log_a) * (1 - exp(-avg_b * dbh_seq^avg_k))

  temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_h = predicted_h, avg_wd = avg_wd)
  prediction_df_ang <- rbind(prediction_df_ang, temp_df)
}

# # Plot DBH vs. Height (H) for Angiosperms with low and high WD
# ggplot(prediction_df_ang, aes(x = dbh, y = predicted_h, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
#   geom_line(alpha = 0.7) +
#   scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
#   labs(x = "DBH (cm)", y = "Height (m)", color = "WD Quantiles") +
#   theme_minimal() +
#   theme(
#     legend.title = element_text(size = 12),
#     plot.title = element_text(size = 14)
#   )

# Plot DBH vs. Height (H) for Angiosperms with low and high WD
p3 <- ggplot(prediction_df_ang, aes(x = dbh, y = predicted_h, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
  geom_line(alpha = 0.7) +
  scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
  labs(x = "DBH", y = "Height", color = "WD Quantiles") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    plot.title = element_text(size = 14),
    axis.text = element_blank()  # Remove axis tick labels
  )

p3

```

# Combine in 1 plot

```{r}
p1 <- p1 + theme(legend.position = "none")
p2 <- p2 + theme(legend.position = "none")
p3 <- p3 + theme(legend.position = "right")

p <- p1 + p2 + p3 + plot_layout(ncol = 3)
p

```

# Writing function

## p1
```{r}
library(ggplot2)
library(targets)
library(tidyverse)
library(patchwork)
```

```{r}
fit_wd_ang_cr_summary_pl_wd <- tar_read(fit_wd_ang_cr_summary_pl_wd)
fit_wd_ang_cr_summary_pl_wd
beta_ang_cr_pl  <- fit_wd_ang_cr_summary_pl_wd |> filter(grepl("beta", variable))
beta_ang_cr_pl

tallo_reduced_lr_df_ang_cr <- tar_read(tallo_reduced_lr_df_ang_cr)
tallo_reduced_lr_df_ang_cr

```

```{r}
generate_p1 <- function(fit_wd_ang_cr_summary_pl_wd, tallo_reduced_lr_df_ang_cr) {
  # Filter the beta parameters from the fit data
  beta_ang_cr_pl <- fit_wd_ang_cr_summary_pl_wd |> 
    filter(grepl("beta", variable))

  # Filter Angiosperm data from the tallo data
  ang_cr_df <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")

  # Create species ID mapping
  sp_id_mapping <- ang_cr_df |> 
    select(sp) |> 
    distinct() |> 
    mutate(sp_id = row_number())

  # Process beta parameters into log_a and b
  angiosperm_parameters <- beta_ang_cr_pl |>
    mutate(param_type = case_when(
      grepl("1,", variable) ~ "log_a",
      grepl("2,", variable) ~ "b"
    )) |>
    mutate(sp_id = rep(1:(nrow(beta_ang_cr_pl) / 2), each = 2))

  # Pivot the parameters into wide format
  angiosperm_params_wide <- angiosperm_parameters |>
    pivot_wider(
      names_from = param_type,
      values_from = mean,
      id_cols = sp_id
    )

  # Combine species parameters with the Angiosperm data
  angiosperm_params_with_sp <- angiosperm_params_wide |>
    left_join(sp_id_mapping, by = "sp_id")

  # Create the full dataset by joining the tree data and species parameters
  angiosperm_data_full <- ang_cr_df |>
    left_join(angiosperm_params_with_sp, by = "sp") |>
    select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b)

  # Summarize species data to calculate avg values
  species_summary <- angiosperm_data_full |>
    group_by(sp) |>
    summarize(
      avg_dbh = mean(dbh, na.rm = TRUE),
      avg_log_a = mean(log_a, na.rm = TRUE),
      avg_b = mean(b, na.rm = TRUE),
      avg_wd = mean(wd, na.rm = TRUE)
    )

  # Calculate the 25% and 75% quantiles of WD
  wd_quantiles <- quantile(species_summary$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)

  # Filter data for low and high WD species
  low_wd_species <- species_summary |>
    filter(abs(avg_wd - wd_quantiles[1]) == min(abs(avg_wd - wd_quantiles[1])))
  high_wd_species <- species_summary |>
    filter(abs(avg_wd - wd_quantiles[2]) == min(abs(avg_wd - wd_quantiles[2])))

  # Combine low and high WD data
  wd_filtered <- bind_rows(low_wd_species, high_wd_species)

  # Create DBH sequence
  dbh_seq <- seq(min(species_summary$avg_dbh), max(species_summary$avg_dbh), length.out = 100)

  # Generate predictions for low and high WD
  prediction_df <- data.frame()

  for (i in 1:nrow(wd_filtered)) {
    sp <- wd_filtered$sp[i]
    avg_log_a <- wd_filtered$avg_log_a[i]
    avg_b <- wd_filtered$avg_b[i]
    avg_wd <- wd_filtered$avg_wd[i]

    predicted_cr <- exp(avg_log_a) * (dbh_seq ^ avg_b)

    temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
    prediction_df <- rbind(prediction_df, temp_df)
  }

  # Plot
  p1 <- ggplot(prediction_df, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
    geom_line(alpha = 0.7) +
    scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
    labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +  # Removed units from labels
    theme_minimal() +
    theme(
      legend.title = element_text(size = 12),
      plot.title = element_text(size = 14),
      axis.text = element_blank()  # Remove axis tick labels
    )

  return(p1)
}

# Now call the function using the fit data and tallo data
p1 <- generate_p1(fit_wd_ang_cr_summary_pl_wd, tallo_reduced_lr_df_ang_cr)
print(p1)
```

## p2

```{r}
fit_wd_gym_cr_summary_gmm_wd <- tar_read(fit_wd_gym_cr_summary_gmm_wd)
fit_wd_gym_cr_summary_gmm_wd
beta_gym_cr_gmm  <- fit_wd_gym_cr_summary_gmm_wd |> filter(grepl("beta", variable))
beta_gym_cr_gmm

tallo_reduced_nlr_df_gym_cr <- tar_read(tallo_reduced_nlr_df_gym_cr)

generate_p2 <- function(fit_wd_gym_cr_summary_gmm_wd, tallo_reduced_nlr_df_gym_cr) {
  # Filter the beta parameters from the fit data
  beta_gym_cr_gmm <- fit_wd_gym_cr_summary_gmm_wd |> filter(grepl("beta", variable))

  # Filter Gymnosperm data from the tallo data
  gym_cr_df <- tallo_reduced_nlr_df_gym_cr |> filter(division == "Gymnosperm")

  # Create species ID mapping
  sp_id_mapping_gym <- gym_cr_df |> 
    select(sp) |> 
    distinct() |> 
    mutate(sp_id = row_number())

  # Process beta parameters into log_a, b, and k
  gymnosperm_parameters <- beta_gym_cr_gmm |>
    mutate(param_type = case_when(
      grepl("\\[.*,1\\]", variable) ~ "log_a",
      grepl("\\[.*,2\\]", variable) ~ "b",
      grepl("\\[.*,3\\]", variable) ~ "k"
    )) |>
    mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))

  # Pivot the parameters into wide format
  gymnosperm_params_wide <- gymnosperm_parameters |>
    pivot_wider(
      names_from = param_type,
      values_from = mean,
      id_cols = sp_id
    )

  # Combine species parameters with the Gymnosperm data
  gymnosperm_params_with_sp <- gymnosperm_params_wide |>
    left_join(sp_id_mapping_gym, by = "sp_id")

  # Create the full dataset by joining the tree data and species parameters
  gymnosperm_data_full <- gym_cr_df |>
    left_join(gymnosperm_params_with_sp, by = "sp") |>
    select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b, k)

  # Summarize species data to calculate avg values
  species_summary_gym <- gymnosperm_data_full |>
    group_by(sp) |>
    summarize(
      avg_dbh = mean(dbh, na.rm = TRUE),
      avg_log_a = mean(log_a, na.rm = TRUE),
      avg_b = mean(b, na.rm = TRUE),
      avg_k = mean(k, na.rm = TRUE),
      avg_wd = mean(wd, na.rm = TRUE)
    )

  # Calculate the 25% and 75% quantiles of WD
  wd_quantiles_gym <- quantile(species_summary_gym$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)

  # Filter data for low and high WD species
  low_wd_species_gym <- species_summary_gym |>
    filter(abs(avg_wd - wd_quantiles_gym[1]) == min(abs(avg_wd - wd_quantiles_gym[1])))
  high_wd_species_gym <- species_summary_gym |>
    filter(abs(avg_wd - wd_quantiles_gym[2]) == min(abs(avg_wd - wd_quantiles_gym[2])))

  # Combine low and high WD data
  wd_filtered_gym <- bind_rows(low_wd_species_gym, high_wd_species_gym)

  # Create DBH sequence
  dbh_seq <- seq(min(species_summary_gym$avg_dbh), max(species_summary_gym$avg_dbh), length.out = 100)

  # Generate predictions for low and high WD
  prediction_df_gym <- data.frame()

  for (i in 1:nrow(wd_filtered_gym)) {
    sp <- wd_filtered_gym$sp[i]
    avg_log_a <- wd_filtered_gym$avg_log_a[i]
    avg_b <- wd_filtered_gym$avg_b[i]
    avg_k <- wd_filtered_gym$avg_k[i]
    avg_wd <- wd_filtered_gym$avg_wd[i]

    predicted_cr <- (exp(avg_log_a) * dbh_seq^avg_b) / (avg_k + dbh_seq^avg_b)

    temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
    prediction_df_gym <- rbind(prediction_df_gym, temp_df)
  }

  # Plot
  p2 <- ggplot(prediction_df_gym, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
    geom_line(alpha = 0.7) +
    scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
    labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +  # Removed units from labels
    theme_minimal() +
    theme(
      legend.title = element_text(size = 12),
      plot.title = element_text(size = 14),
      axis.text = element_blank()  # Remove axis tick labels
    )

  return(p2)
}

# Now call the function using the fit data and tallo data
p2 <- generate_p2(fit_wd_gym_cr_summary_gmm_wd, tallo_reduced_nlr_df_gym_cr)
print(p2)
```

## p3
```{r}
fit_wd_ang_h_summary_weibull_wd <- tar_read(fit_wd_ang_h_summary_weibull_wd)
fit_wd_ang_h_summary_weibull_wd

beta_ang_h_wb  <- fit_wd_ang_h_summary_weibull_wd |> filter(grepl("beta", variable))
beta_ang_h_wb

tallo_reduced_nlr_df_ang_h <- tar_read(tallo_reduced_nlr_df_ang_h)
tallo_reduced_nlr_df_ang_h

generate_p3 <- function(fit_wd_ang_h_summary_weibull_wd, tallo_reduced_nlr_df_ang_h) {
  # Filter the beta parameters from the fit data
  beta_ang_h_wb <- fit_wd_ang_h_summary_weibull_wd |> filter(grepl("beta", variable))

  # Filter Angiosperm data from the tallo data
  ang_h_df <- tallo_reduced_nlr_df_ang_h |> filter(division == "Angiosperm")

  # Create species ID mapping
  sp_id_mapping_ang <- ang_h_df |>
    select(sp) |>
    distinct() |>
    mutate(sp_id = row_number())

  # Process beta parameters into log_a, b, and k
  beta_ang_h_weibull <- beta_ang_h_wb |>
    mutate(param_type = case_when(
      grepl("\\[.*,1\\]", variable) ~ "log_a",
      grepl("\\[.*,2\\]", variable) ~ "b",
      grepl("\\[.*,3\\]", variable) ~ "k"
    )) |>
    mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))

  # Pivot the parameters into wide format
  angiosperm_params_wide <- beta_ang_h_weibull |>
    pivot_wider(
      names_from = param_type,
      values_from = mean,
      id_cols = sp_id
    )

  # Combine species parameters with the Angiosperm data
  angiosperm_params_with_sp <- angiosperm_params_wide |>
    left_join(sp_id_mapping_ang, by = "sp_id")

  # Create the full dataset by joining the tree data and species parameters
  angiosperm_data_full <- ang_h_df |>
    left_join(angiosperm_params_with_sp, by = "sp") |>
    select(tree_id, sp, sp_id, dbh, h, wd, log_a, b, k)

  # Summarize species data to calculate avg values
  species_summary_ang <- angiosperm_data_full |>
    group_by(sp) |>
    summarize(
      avg_dbh = mean(dbh, na.rm = TRUE),
      avg_h = mean(h, na.rm = TRUE),
      avg_log_a = mean(log_a, na.rm = TRUE),
      avg_b = mean(b, na.rm = TRUE),
      avg_k = mean(k, na.rm = TRUE),
      avg_wd = mean(wd, na.rm = TRUE)
    )

  # Calculate the 25% and 75% quantiles of WD
  wd_quantiles_ang <- quantile(species_summary_ang$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)

  # Filter data for low and high WD species
  low_wd_species_ang <- species_summary_ang |>
    filter(abs(avg_wd - wd_quantiles_ang[1]) == min(abs(avg_wd - wd_quantiles_ang[1])))
  high_wd_species_ang <- species_summary_ang |>
    filter(abs(avg_wd - wd_quantiles_ang[2]) == min(abs(avg_wd - wd_quantiles_ang[2])))

  # Combine low and high WD data
  wd_filtered_ang <- bind_rows(low_wd_species_ang, high_wd_species_ang)

  # Create DBH sequence
  dbh_seq <- seq(min(species_summary_ang$avg_dbh), max(species_summary_ang$avg_dbh), length.out = 100)

  # Generate height predictions for low and high WD
  prediction_df_ang <- data.frame()

  for (i in 1:nrow(wd_filtered_ang)) {
    sp <- wd_filtered_ang$sp[i]
    avg_log_a <- wd_filtered_ang$avg_log_a[i]
    avg_b <- wd_filtered_ang$avg_b[i]
    avg_k <- wd_filtered_ang$avg_k[i]
    avg_wd <- wd_filtered_ang$avg_wd[i]

    predicted_h <- exp(avg_log_a) * (1 - exp(-avg_b * dbh_seq^avg_k))

    temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_h = predicted_h, avg_wd = avg_wd)
    prediction_df_ang <- rbind(prediction_df_ang, temp_df)
  }

  # Plot
  p3 <- ggplot(prediction_df_ang, aes(x = dbh, y = predicted_h, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
    geom_line(alpha = 0.7) +
    scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
    labs(x = "DBH", y = "Height", color = "WD Quantiles") +
    theme_minimal() +
    theme(
      legend.title = element_text(size = 12),
      plot.title = element_text(size = 14),
      axis.text = element_blank()  # Remove axis tick labels
    )

  return(p3)
}

p3 <- generate_p3(fit_wd_ang_h_summary_weibull_wd, tallo_reduced_nlr_df_ang_h)
print(p3)
```

# 3 function in 1
```{r}

generate_wd_ef <- function(fit_wd_ang_cr_summary_pl_wd, tallo_reduced_lr_df_ang_cr,
                           fit_wd_gym_cr_summary_gmm_wd, tallo_reduced_nlr_df_gym_cr,
                           fit_wd_ang_h_summary_weibull_wd, tallo_reduced_nlr_df_ang_h) {

  generate_p1 <- function() {
    beta_ang_cr_pl <- fit_wd_ang_cr_summary_pl_wd |> filter(grepl("beta", variable))
    ang_cr_df <- tallo_reduced_lr_df_ang_cr |> filter(division == "Angiosperm")
    sp_id_mapping <- ang_cr_df |> select(sp) |> distinct() |> mutate(sp_id = row_number())
    angiosperm_parameters <- beta_ang_cr_pl |>
      mutate(param_type = case_when(
        grepl("1,", variable) ~ "log_a",
        grepl("2,", variable) ~ "b"
      )) |>
      mutate(sp_id = rep(1:(nrow(beta_ang_cr_pl) / 2), each = 2))
    angiosperm_params_wide <- angiosperm_parameters |>
      pivot_wider(names_from = param_type, values_from = mean, id_cols = sp_id)
    angiosperm_params_with_sp <- angiosperm_params_wide |>
      left_join(sp_id_mapping, by = "sp_id")
    angiosperm_data_full <- ang_cr_df |>
      left_join(angiosperm_params_with_sp, by = "sp") |>
      select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b)
    species_summary <- angiosperm_data_full |>
      group_by(sp) |>
      summarize(avg_dbh = mean(dbh, na.rm = TRUE), avg_log_a = mean(log_a, na.rm = TRUE),
                avg_b = mean(b, na.rm = TRUE), avg_wd = mean(wd, na.rm = TRUE))
    wd_quantiles <- quantile(species_summary$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)
    low_wd_species <- species_summary |> filter(abs(avg_wd - wd_quantiles[1]) == min(abs(avg_wd - wd_quantiles[1])))
    high_wd_species <- species_summary |> filter(abs(avg_wd - wd_quantiles[2]) == min(abs(avg_wd - wd_quantiles[2])))
    wd_filtered <- bind_rows(low_wd_species, high_wd_species)
    dbh_seq <- seq(min(species_summary$avg_dbh), max(species_summary$avg_dbh), length.out = 100)
    prediction_df <- data.frame()
    for (i in 1:nrow(wd_filtered)) {
      sp <- wd_filtered$sp[i]
      avg_log_a <- wd_filtered$avg_log_a[i]
      avg_b <- wd_filtered$avg_b[i]
      avg_wd <- wd_filtered$avg_wd[i]
      predicted_cr <- exp(avg_log_a) * (dbh_seq ^ avg_b)
      temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
      prediction_df <- rbind(prediction_df, temp_df)
    }
    p1 <- ggplot(prediction_df, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
      geom_line(alpha = 0.7) +
      scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
      labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +
      theme_minimal() +
      theme(legend.title = element_text(size = 12), plot.title = element_text(size = 14), axis.text = element_blank())
    return(p1)
  }

  generate_p2 <- function() {
    beta_gym_cr_gmm <- fit_wd_gym_cr_summary_gmm_wd |> filter(grepl("beta", variable))
    gym_cr_df <- tallo_reduced_nlr_df_gym_cr |> filter(division == "Gymnosperm")
    sp_id_mapping_gym <- gym_cr_df |> select(sp) |> distinct() |> mutate(sp_id = row_number())
    gymnosperm_parameters <- beta_gym_cr_gmm |>
      mutate(param_type = case_when(
        grepl("\\[.*,1\\]", variable) ~ "log_a",
        grepl("\\[.*,2\\]", variable) ~ "b",
        grepl("\\[.*,3\\]", variable) ~ "k"
      )) |>
      mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))
    gymnosperm_params_wide <- gymnosperm_parameters |>
      pivot_wider(names_from = param_type, values_from = mean, id_cols = sp_id)
    gymnosperm_params_with_sp <- gymnosperm_params_wide |>
      left_join(sp_id_mapping_gym, by = "sp_id")
    gymnosperm_data_full <- gym_cr_df |>
      left_join(gymnosperm_params_with_sp, by = "sp") |>
      select(tree_id, sp, sp_id, dbh, cr, h, wd, log_a, b, k)
    species_summary_gym <- gymnosperm_data_full |>
      group_by(sp) |>
      summarize(avg_dbh = mean(dbh, na.rm = TRUE), avg_log_a = mean(log_a, na.rm = TRUE),
                avg_b = mean(b, na.rm = TRUE), avg_k = mean(k, na.rm = TRUE), avg_wd = mean(wd, na.rm = TRUE))
    wd_quantiles_gym <- quantile(species_summary_gym$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)
    low_wd_species_gym <- species_summary_gym |> filter(abs(avg_wd - wd_quantiles_gym[1]) == min(abs(avg_wd - wd_quantiles_gym[1])))
    high_wd_species_gym <- species_summary_gym |> filter(abs(avg_wd - wd_quantiles_gym[2]) == min(abs(avg_wd - wd_quantiles_gym[2])))
    wd_filtered_gym <- bind_rows(low_wd_species_gym, high_wd_species_gym)
    dbh_seq <- seq(min(species_summary_gym$avg_dbh), max(species_summary_gym$avg_dbh), length.out = 100)
    prediction_df_gym <- data.frame()
    for (i in 1:nrow(wd_filtered_gym)) {
      sp <- wd_filtered_gym$sp[i]
      avg_log_a <- wd_filtered_gym$avg_log_a[i]
      avg_b <- wd_filtered_gym$avg_b[i]
      avg_k <- wd_filtered_gym$avg_k[i]
      avg_wd <- wd_filtered_gym$avg_wd[i]
      predicted_cr <- (exp(avg_log_a) * dbh_seq^avg_b) / (avg_k + dbh_seq^avg_b)
      temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_cr = predicted_cr, avg_wd = avg_wd)
      prediction_df_gym <- rbind(prediction_df_gym, temp_df)
    }
    p2 <- ggplot(prediction_df_gym, aes(x = dbh, y = predicted_cr, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
      geom_line(alpha = 0.7) +
      scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
      labs(x = "DBH", y = "Crown Radius", color = "WD Quantiles") +
      theme_minimal() +
      theme(legend.title = element_text(size = 12), plot.title = element_text(size = 14), axis.text = element_blank())
    return(p2)
  }

  generate_p3 <- function() {
    beta_ang_h_wb <- fit_wd_ang_h_summary_weibull_wd |> filter(grepl("beta", variable))
    ang_h_df <- tallo_reduced_nlr_df_ang_h |> filter(division == "Angiosperm")
    sp_id_mapping_ang <- ang_h_df |> select(sp) |> distinct() |> mutate(sp_id = row_number())
    beta_ang_h_weibull <- beta_ang_h_wb |>
      mutate(param_type = case_when(
        grepl("\\[.*,1\\]", variable) ~ "log_a",
        grepl("\\[.*,2\\]", variable) ~ "b",
        grepl("\\[.*,3\\]", variable) ~ "k"
      )) |>
      mutate(sp_id = as.numeric(gsub("beta\\[|,.*", "", variable)))
    angiosperm_params_wide <- beta_ang_h_weibull |>
      pivot_wider(names_from = param_type, values_from = mean, id_cols = sp_id)
    angiosperm_params_with_sp <- angiosperm_params_wide |>
      left_join(sp_id_mapping_ang, by = "sp_id")
    angiosperm_data_full <- ang_h_df |>
      left_join(angiosperm_params_with_sp, by = "sp") |>
      select(tree_id, sp, sp_id, dbh, h, wd, log_a, b, k)
    species_summary_ang <- angiosperm_data_full |>
      group_by(sp) |>
      summarize(avg_dbh = mean(dbh, na.rm = TRUE), avg_h = mean(h, na.rm = TRUE),
                avg_log_a = mean(log_a, na.rm = TRUE), avg_b = mean(b, na.rm = TRUE),
                avg_k = mean(k, na.rm = TRUE), avg_wd = mean(wd, na.rm = TRUE))
    wd_quantiles_ang <- quantile(species_summary_ang$avg_wd, probs = c(0.25, 0.75), na.rm = TRUE)
    low_wd_species_ang <- species_summary_ang |>
      filter(abs(avg_wd - wd_quantiles_ang[1]) == min(abs(avg_wd - wd_quantiles_ang[1])))
    high_wd_species_ang <- species_summary_ang |>
      filter(abs(avg_wd - wd_quantiles_ang[2]) == min(abs(avg_wd - wd_quantiles_ang[2])))
    wd_filtered_ang <- bind_rows(low_wd_species_ang, high_wd_species_ang)
    dbh_seq <- seq(min(species_summary_ang$avg_dbh), max(species_summary_ang$avg_dbh), length.out = 100)
    prediction_df_ang <- data.frame()
    for (i in 1:nrow(wd_filtered_ang)) {
      sp <- wd_filtered_ang$sp[i]
      avg_log_a <- wd_filtered_ang$avg_log_a[i]
      avg_b <- wd_filtered_ang$avg_b[i]
      avg_k <- wd_filtered_ang$avg_k[i]
      avg_wd <- wd_filtered_ang$avg_wd[i]
      predicted_h <- exp(avg_log_a) * (1 - exp(-avg_b * dbh_seq^avg_k))
      temp_df <- data.frame(sp = sp, dbh = dbh_seq, predicted_h = predicted_h, avg_wd = avg_wd)
      prediction_df_ang <- rbind(prediction_df_ang, temp_df)
    }
    p3 <- ggplot(prediction_df_ang, aes(x = dbh, y = predicted_h, group = avg_wd, color = as.factor(round(avg_wd, 2)))) +
      geom_line(alpha = 0.7) +
      scale_color_manual(values = c("#f48d10", "#7f590e"), labels = c("Light WD species", "Dense WD species")) +
      labs(x = "DBH", y = "Height", color = "WD Quantiles") +
      theme_minimal() +
      theme(legend.title = element_text(size = 12), plot.title = element_text(size = 14), axis.text = element_blank())
    return(p3)
  }

  # Generate all three plots
  p1 <- generate_p1()
  p2 <- generate_p2()
  p3 <- generate_p3()

  # return(list(p1 = p1, p2 = p2, p3 = p3))
  p <- p1 + p2 + p3 + plot_layout(nrow = 1)

  return(p)
}

# Usage
p <- generate_wd_ef(fit_wd_ang_cr_summary_pl_wd, tallo_reduced_lr_df_ang_cr,
                        fit_wd_gym_cr_summary_gmm_wd, tallo_reduced_nlr_df_gym_cr,
                        fit_wd_ang_h_summary_weibull_wd, tallo_reduced_nlr_df_ang_h)

```

```{r}

```

```{r}

```