### 2025-03-27 sup/sup_old.qmd

---
fontsize: 12pt
geometry: margin=1in
link-citations: true
bibliography: "../ms/tree-allometry.bib"
csl: "../ms/journal-of-forestry-research.csl"
crossref:
  fig-title: Figure
  fig-prefix: Figure
  tbl-title: Table
  eq-prefix: Eq.
format:
  pdf:
    toc: false
    keep-tex: true
    latex-clean: false
    pdf-engine: xelatex
    include-in-header:
      text: |
        \usepackage[default]{sourcesanspro}
        \usepackage{sourcecodepro}
        \usepackage{lineno}
        \usepackage{setspace}
        \doublespacing
        \linenumbers
        \renewcommand{\thetable}{S\arabic{table}}
        \renewcommand{\thefigure}{S\arabic{figure}}
    prefer-html: true
  docx:
    toc: false
    include-in-header:
      text: |
        \usepackage[default]{sourcesanspro}
        \usepackage{lineno}
        \linespread{1.2}

---

*Supplement of*

**Saturating allometric relationships reveal how wood density shapes global tree architecture**

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r include=FALSE}
library(targets)
library(tidyverse)
library(here)
library(kableExtra)
library(targets)
library(patchwork)
library(MASS)
library(scales)
library(grid)
library(knitr)
library(gsubfn)
library(yaml)
library(flextable)
library(officer)

source("R/figs.R")
source("R/functions.R")
data_params <- read_yaml("data.yaml")
data_posterior <- read_yaml("posterior.yaml")
data_metrics <- read_yaml("agb_metrics_ang.yaml")
```

\newpage

This supplement includes the following materials:

# Section S1. Extra figures and tables

**Table S1.** List of sources for wood density data used in this study, as released in version 6 of the TRY database.
The table includes authors' names, dataset IDs, and the name of each dataset.

**Table S2.** Characteristics of sub-datasets were used to select the best predictive model of each allometric relationship for angiosperms and gymnosperms.

**Table S3.** Parameter estimates for all hierarchical models for tree height allometry.

**Table S4.** Parameter estimates for all hierarchical models for crown radius allometry.

**Table S5.** Parameter estimates for all hierarchical models for DBH allometry.

**Table S6.** Parameter estimates for predicting tree height (m) from DBH (cm) of 1290 species (see file TableS6_height_sp_estimates.xlsx).

**Table S7.** Parameter estimates for predicting crown radius (m) from DBH (cm) of 821 species (see file TableS7_crown_radius_sp_estimates.xlsx).

**Table S8.** Parameter estimates for predicting DBH (cm) from crown radius (m) and tree height (m) of 800 species (see file TableS8_dbh_sp_estimates.xlsx).

**Figure S1.** Function curves are used to describe allometric relationships of tree dimensions with DBH in log scale.

# Section S2. Stan code used to fit alternative allometric models

\newpage

# Section S1. Extra tables and figures

**Table S1**: List of sources for wood density data utilized in this study, as released in version 6 of the TRY database.
The table includes authors' names, dataset IDs, and the name of each dataset.

```{r echo=FALSE, results='asis'}
try_data <- process_try_data("data-raw/30953.txt")
try_data <- try_data |>
  mutate(across(where(is.character), ~iconv(.x, to = "UTF-8", sub = "")))

ft <- try_data |> flextable () |>
  align(j = colnames(try_data), align = "center", part = "all") |>
  fontsize(size = 7, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit()

if (knitr::is_latex_output()) {
  try_data |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}

```

\newpage

**Table S2**: Characteristics of sub-datasets used to select the best predictive model of each allometric relationship for angiosperms and gymnosperms.
The ranges represent the minimum and maximum values for the respective variables: DBH (cm), tree height (H, m), crown radius (CR, m), and wood density (WD, g cm$^{-3}$).

```{r echo=FALSE, results='asis'}
tbl_subdata <- tar_read(tbl_subdata)

line_rows <- c(5, 10, 15, 20, 25)

ft <- tbl_subdata |>
  flextable() |>
  set_header_labels(
    `Dependent variables` = "Dependent\nvariable",
    `Predictor variable` = "Predictor\nvariable",
    `Characteristics` = "Characteristic",
    `Angiosperm` = "Angiosperm",
    `Gymnosperm` = "Gymnosperm"
  ) |>
  # add_header_row(
  #   values = c("", "", "", "Division"),
  #   colwidths = c(1, 1, 1, 2)
  # ) |>
  align(align = "center", part = "all") |>
  merge_h(i = 1, part = "header") |>
  fontsize(size = 6, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit() |>
  border(i = line_rows, j = 1:ncol(tbl_subdata), border.bottom = fp_border())

if (knitr::is_latex_output()) {
  tbl_subdata |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}

```

\newpage
**Table S3**: Posterior estimates of the parameters of the hierarchical models for tree height allometry.
Table entries correspond to the median and 95\% CIs for the community-level parameters of each allometric function.

```{r echo=FALSE, results='asis'}
posterior_rounded_df <- tar_read(posterior_rounded_df)
df_h <- posterior_rounded_df |>
  filter(Dependent_variable == "Tree Height") |>
  dplyr::select(-Dependent_variable) |>
  mutate(
    Functional_form = factor(Functional_form, levels = c("Power-Law", "gMM", "Weibull")),
    Wood_density = factor(Wood_density, levels = c("Without", "With"))
  ) |>
  arrange(Division, Functional_form, Wood_density)

line_rows <- c(2, 5, 8, 11, 13, 16, 19)
special_line_rows <- c(11)

ft <- df_h |>
  flextable() |>
  set_header_labels(
    Division = "Division",
    Functional_form = "Functional\nform",
    Wood_density = "Wood\ndensity",
    Parameter = "Parameter",
    Intercept_CI = "γ₀ (Base line)",
    Slope_CI = "γ₁ (Exponent rate)",
    Tau_CI = "σ (Variation)"
  ) |>
  align(align = "center", part = "all") |>
  merge_v(j = 1, part = "body") |>
  fontsize(size = 6, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit() |>
  border(i = line_rows, j = 2:ncol(df_h), border.bottom = fp_border()) |>
  border(i = special_line_rows, j = 1:ncol(df_h), border.bottom = fp_border())

if (knitr::is_latex_output()) {
  df_h |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}

```

\newpage
**Table S4**: Posterior estimates of the parameters of the hierarchical models for crown radius allometry.
Table entries correspond to the median and 95\% CIs for the community-level parameters of each allometric function.

```{r echo=FALSE, results='asis'}
posterior_rounded_df <- tar_read(posterior_rounded_df)
df_cr <- posterior_rounded_df |>
  filter(Dependent_variable == "Crown Radius") |>
  dplyr::select(-Dependent_variable) |>
  mutate(
    Functional_form = factor(Functional_form, levels = c("Power-Law", "gMM", "Weibull")),
    Wood_density = factor(Wood_density, levels = c("Without", "With"))
  ) |>
  arrange(Division, Functional_form, Wood_density)

line_rows <- c(2, 4, 7, 10, 12, 15, 18)
special_line_rows <- c(10)

ft <- df_cr |>
  flextable() |>
  set_header_labels(
    Division = "Division",
    Functional_form = "Functional\nform",
    Wood_density = "Wood\ndensity",
    Parameter = "Parameter",
    Intercept_CI = "γ₀ (Base line)",
    Slope_CI = "γ₁ (Exponent rate)",
    Tau_CI = "σ (Variation)"
  ) |>
  align(align = "center", part = "all") |>
  merge_v(j = 1, part = "body") |>
  fontsize(size = 6, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit() |>
  border(i = line_rows, j = 2:ncol(df_cr), border.bottom = fp_border()) |>
  border(i = special_line_rows, j = 1:ncol(df_cr), border.bottom = fp_border())

if (knitr::is_latex_output()) {
  df_cr |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}

```

\newpage
**Table S5**: Posterior estimates of the parameters of the hierarchical models for DBH allometry.
Table entries correspond to the median and 95\% CIs for the community-level parameters of each allometric function.

```{r echo=FALSE, results='asis'}
posterior_rounded_df <- tar_read(posterior_rounded_df)

df_dbh <- posterior_rounded_df |>
  filter(Dependent_variable %in% c("DBH", "DBH1", "DBH2", "DBH3")) |>
  mutate(
    Predictor = case_when(
      Dependent_variable == "DBH1" ~ "CR × H",
      Dependent_variable == "DBH2" ~ "CR",
      Dependent_variable == "DBH3" ~ "H",
      Dependent_variable == "DBH" ~ "CR, H",
      TRUE ~ NA_character_
    )
  ) |>
  dplyr::select(-Dependent_variable)

df_dbh <- df_dbh |>
  mutate(Predictor = factor(Predictor, levels = c("CR, H", "CR × H", "CR", "H"))) |>
  dplyr::select(Division, Predictor, everything()) |>
  dplyr::arrange(Division,
          Predictor,
          factor(Wood_density, levels = c("Without", "With")),
          Parameter)

line_rows <- c(3, 6, 8, 10, 12, 15, 18, 20, 22)
special_line_rows <- c(12)

ft <- df_dbh |>
  flextable() |>

  set_header_labels(
    Division = "Division",
    Predictor = "Predictor\nvariable",
    Functional_form = "Functional\nform",
    Wood_density = "Wood\ndensity",
    Parameter = "Parameter",
    Intercept_CI = "γ₀ (Base line)",
    Slope_CI = "γ₁ (Exponent rate)",
    Tau_CI = "σ (Variation)"
  ) |>

  align(align = "center", part = "all") |>
  merge_v(j = 1, part = "body") |>
  fontsize(size = 6, part = "all") |>
  set_table_properties(layout = "autofit") |>
  autofit()

ft <- ft |>
  border(i = line_rows, j = 2:ncol(df_dbh), border.bottom = fp_border()) |>
  border(i = special_line_rows, j = 1:ncol(df_dbh), border.bottom = fp_border())

if (knitr::is_latex_output()) {
  df_dbh |> kable(align = "c", booktabs = TRUE) |>
    kable_styling(full_width = FALSE, position = "center", font_size = 10, latex_options = c("hold_position"))
} else {
    ft
}
```

\newpage

![
Allometric relationships between tree height and crown radius with DBH of angiosperm (a, c) and gymnosperm (b, d) based on three different models: power-law (orange), generalized Michaelis-Menten (green), and Weibull (blue).
The density of overlapping points is shown using a colour gradient ranging from black (low density) to yellow (high density).
Point density was calculated using a 2D kernel density estimation on a 300 $\times$ 300 grid, estimating the concentration of data points in the log-transformed space of DBH and tree height or crown radius.
All axes are on the log scale.
](../figs/h_cr_dbh_log.png){#fig-compare fig-pos="H"}

\newpage

# Section S2. Stan code used to fit alternative allometric models

## Power-law model

```{stan, class.source="small", file="stan/pl.stan", echo=TRUE, eval=FALSE, output.var="hoge", tidy=TRUE}
```

\newpage

## Power-law model with wood density

```{stan, size="small", file="stan/pl_wd.stan", echo=TRUE, eval=FALSE, output.var="hoge", tidy=TRUE}
```

\newpage

## Generalized Michaelis-Menten model

```{stan, size="small", file="stan/gmm.stan", echo=TRUE, eval=FALSE, output.var="hoge"}
```

\newpage

## Generalized Michaelis-Menten model with wood density

```{stan, size="small", file="stan/gmm_wd.stan", echo=TRUE, eval=FALSE, output.var="hoge"}
```

\newpage

## Weibull model

```{stan, size="small", file="stan/weibull.stan", echo=TRUE, eval=FALSE, output.var="hoge"}
```

\newpage

## Weibull model with wood density

```{stan, size="small", file="stan/weibull_wd.stan", echo=TRUE, eval=FALSE, output.var="hoge"}
```